---
title: "lastingChange_DREAM"
author: "Nasim Bararpour"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}

dex<- read.csv("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/MetaData/burnoutEx.csv")
rownames(dex)<-dex$X
```

```{r}
meta<- dex[c(5438:5442)]
dEX<- dex[c(6:5432)]
```


```{r}
# remove all columns that has more than 30% NAs
percent.na <- function(x){
  n <- x[is.na(x)] %>% length()
  return(n/length(x)*100)
}
```

#na removal create missingness 
```{r}
filt.col <- apply(dEX,2,function(x) percent.na(x)<=30)#feature in column
filt.dat <-dEX[ ,filt.col]

lEx<-log2(filt.dat+1)

```

#Normalization without replacement 
```{r}
meta<- dex[c(5433:5437)]
dEX<- dex[c(6:5432)]

dEX<- dEX+1

lEx<-sqrt(dEX+1) # final used 
filt.dat<-dEX
```

#check distribution

```{r}
# Single protein comparison
plot(density(lEx$Q8WTS1, na.rm=TRUE), col="blue", lwd=2, main="Q8WTS1 Distribution")
lines(density(filt.dat$Q8WTS1, na.rm=TRUE), col="red", lwd=2)
legend("topright", c("lEx", "filt.dat"), col=c("blue", "red"), lwd=2)
```

#Side by side box

```{r}
boxplot(lEx$Q8WTS1, filt.dat$Q8WTS1, 
        names=c("lEx", "filt.dat"), 
        col=c("lightblue", "lightcoral"),
        main="Q8WTS1")
```

#Scatter Plot with Correlation for large data

```{r}
# Compare means or medians
means_lEx <- colMeans(lEx, na.rm=TRUE)
means_filt <- colMeans(filt.dat, na.rm=TRUE)

plot(means_lEx, means_filt, 
     xlab="lEx (mean)", ylab="filt.dat (mean)",
     main="Protein Expression Comparison")
abline(0, 1, col="red")  # y=x line
cor_val <- cor(means_lEx, means_filt, use="complete.obs")
text(min(means_lEx), max(means_filt), 
     paste("Cor =", round(cor_val, 3)), pos=4)
```

\MA Plot** (Bland-Altman style)

```{r}
M <- means_lEx - means_filt  # Difference
A <- (means_lEx + means_filt) / 2  # Average

plot(A, M, 
     xlab="Average Expression", 
     ylab="Difference (lEx - filt.dat)",
     main="MA Plot")
abline(h=0, col="red")
```

For checking normality of distributions, here are the best approaches:

## **Visual Methods:**

### 1. **Q-Q Plot** (Most Reliable Visual Check)

```{r}
# Single protein
par(mfrow=c(1,2))
qqnorm(lEx$Q8WTS1, main="lEx - Q8WTS1")
qqline(lEx$Q8WTS1, col="red")

qqnorm(filt.dat$Q8WTS1, main="filt.dat - Q8WTS1")
qqline(filt.dat$Q8WTS1, col="red")
```

Histogram for normality

```{r}
par(mfrow=c(1,2))

# lEx
hist(lEx$Q8WTS1, breaks=20, prob=TRUE, main="lEx", col="lightblue")
curve(dnorm(x, mean=mean(lEx$Q8WTS1, na.rm=TRUE), 
            sd=sd(lEx$Q8WTS1, na.rm=TRUE)), 
      add=TRUE, col="red", lwd=2)

# filt.dat
hist(filt.dat$Q8WTS1, breaks=20, prob=TRUE, main="filt.dat", col="lightcoral")
curve(dnorm(x, mean=mean(filt.dat$Q8WTS1, na.rm=TRUE), 
            sd=sd(filt.dat$Q8WTS1, na.rm=TRUE)), 
      add=TRUE, col="red", lwd=2)
```

3\. **Shapiro-Wilk Test** (Most Powerful, n\<5000)

```{r}
# Single protein
shapiro.test(lEx$Q8WTS1)
shapiro.test(filt.dat$Q8WTS1)

# p-value > 0.05 suggests normal distribution
```

4\. **Anderson-Darling Test** (Better for larger samples)

```{r}
library(nortest)
ad.test(lEx$Q8WTS1)
ad.test(filt.dat$Q8WTS1)
```

For All 5000 Proteins at Once:

```{r}
library(nortest)

# Function to test normality
test_normality <- function(x) {
  if(length(na.omit(x)) < 3) return(NA)
  if(length(na.omit(x)) > 5000) {
    # Use KS test for large samples
    ks.test(x, "pnorm", mean=mean(x, na.rm=TRUE), sd=sd(x, na.rm=TRUE))$p.value
  } else {
    # Use Shapiro-Wilk for smaller samples
    shapiro.test(na.omit(x))$p.value
  }
}

# Test all proteins in both datasets
pvals_lEx <- sapply(lEx, test_normality)
pvals_filt <- sapply(filt.dat, test_normality)

# Compare proportion of normally distributed proteins
sum(pvals_lEx > 0.05, na.rm=TRUE) / sum(!is.na(pvals_lEx))
sum(pvals_filt > 0.05, na.rm=TRUE) / sum(!is.na(pvals_filt))

# Visualize results
par(mfrow=c(1,2))
hist(pvals_lEx, breaks=50, main="lEx - Normality p-values", 
     xlab="p-value", col="lightblue")
abline(v=0.05, col="red", lwd=2)

hist(pvals_filt, breaks=50, main="filt.dat - Normality p-values", 
     xlab="p-value", col="lightcoral")
abline(v=0.05, col="red", lwd=2)
```

test normality

Practical Example:

```{r}
# Test one protein
p_lEx <- test_normality(lEx$Q8WTS1)
p_filt <- test_normality(filt.dat$Q8WTS1)

cat("lEx p-value:", p_lEx, "\n")
cat("filt.dat p-value:", p_filt, "\n")

# Interpretation
if(p_lEx > 0.05) {
  cat("lEx: Data appears NORMAL\n")
} else {
  cat("lEx: Data is NOT NORMAL\n")
}

if(p_filt > 0.05) {
  cat("filt.dat: Data appears NORMAL\n")
} else {
  cat("filt.dat: Data is NOT NORMAL\n")
}
```

#for all proteins

```{r}
#lEx<-log2(dEX+100)


pvals_lEx <- sapply(lEx, test_normality)
pvals_filt <- sapply(filt.dat, test_normality)

# Count how many proteins are normally distributed
normal_count_lEx <- sum(pvals_lEx > 0.05, na.rm=TRUE)
normal_count_filt <- sum(pvals_filt > 0.05, na.rm=TRUE)

cat("lEx: ", normal_count_lEx, "out of", sum(!is.na(pvals_lEx)), 
    "proteins are normally distributed\n")
cat("filt.dat: ", normal_count_filt, "out of", sum(!is.na(pvals_filt)), 
    "proteins are normally distributed\n")

# As percentages
pct_lEx <- normal_count_lEx / sum(!is.na(pvals_lEx)) * 100
pct_filt <- normal_count_filt / sum(!is.na(pvals_filt)) * 100

cat("\nlEx:", round(pct_lEx, 1), "% normally distributed\n")
cat("filt.dat:", round(pct_filt, 1), "% normally distributed\n")
```

```{r}
# Histogram of p-values
hist(pvals_lEx, breaks=50, 
     main="Distribution of Normality Test p-values (lEx)",
     xlab="p-value",
     col=ifelse(seq(0, 1, length.out=50) > 0.05, "lightgreen", "lightcoral"))
abline(v=0.05, col="red", lwd=2, lty=2)
legend("topright", 
       c("Normal (p>0.05)", "Non-normal (p<0.05)"), 
       fill=c("lightgreen", "lightcoral"))
```
```{r}
library(DESeq2)
```

#use sqrt normalize 
```{r}
ex<-t(dEX)

library(DESeq2)

# === STEP 1: CLEAN YOUR DATA ===

# Check for NAs
cat("Checking for NA values...\n")
cat("lEx NAs:", sum(is.na(lEx)), "\n")
cat("filt.dat NAs:", sum(is.na(filt.dat)), "\n")

# Option A: Remove rows with ANY NAs
lEx_clean <- ex[complete.cases(ex), ]
filt_clean <- filt.dat[complete.cases(filt.dat), ]

# Option B: Replace NAs with 0 (if appropriate for your data)
# lEx_clean <- lEx
# lEx_clean[is.na(lEx_clean)] <- 0

cat("\nAfter cleaning:\n")
cat("lEx: ", nrow(lEx_clean), "proteins,", ncol(lEx_clean), "samples\n")
cat("filt.dat: ", nrow(filt_clean), "proteins,", ncol(filt_clean), "samples\n")

# === STEP 2: FIX COLUMN/ROW NAMES ===

# Check for empty or problematic names
cat("\nChecking names...\n")
cat("lEx column names:", head(colnames(lEx_clean)), "\n")
cat("lEx row names:", head(rownames(lEx_clean)), "\n")

# Fix if needed
if(any(colnames(lEx_clean) == "" | is.na(colnames(lEx_clean)))) {
  colnames(lEx_clean) <- paste0("Sample", 1:ncol(lEx_clean))
  cat("Fixed empty column names\n")
}

if(any(rownames(lEx_clean) == "" | is.na(rownames(lEx_clean)))) {
  rownames(lEx_clean) <- paste0("Protein", 1:nrow(lEx_clean))
  cat("Fixed empty row names\n")
}

# === STEP 3: ENSURE INTEGER COUNTS ===

# DESeq2 needs integer counts
lEx_clean <- round(lEx_clean)
filt_clean <- round(filt_clean)

# Convert to matrix if needed
lEx_clean <- as.matrix(lEx_clean)
filt_clean <- as.matrix(filt_clean)

cat("\nData cleaned successfully!\n")
```

#check negative
```{r}
# Check your data
cat("=== Data Inspection ===\n")
cat("lEx range:", range(lEx, na.rm=TRUE), "\n")
cat("filt.dat range:", range(filt.dat, na.rm=TRUE), "\n")
cat("lEx has negative values:", any(lEx < 0, na.rm=TRUE), "\n")
cat("filt.dat has negative values:", any(filt.dat < 0, na.rm=TRUE), "\n")
cat("lEx has decimals:", any(lEx != floor(lEx), na.rm=TRUE), "\n")
cat("lEx first few values:\n")
print(head(lEx[, 1:3]))

# Check distribution
hist(as.matrix(lEx_clean), breaks=100, main="lEx Distribution")
```


Outcome = Protein levels | Predictors = Time, Sex, Age
Best Method: limma with duplicateCorrelation
This handles repeated measures and works with your non-normal, processed data.


```{r}
library(limma)

ex<-t(dEX)
lEx_clean <- ex[complete.cases(ex), ]
# ============================================
# SETUP DATA FOR QUESTION 1
# ============================================

# 1. Prepare expression matrix
expr_matrix <- as.matrix(lEx_clean)  # proteins x samples
# Rows = proteins, Columns = samples (all timepoints for all subjects)

# 2. Create comprehensive metadata
metadata <- meta
colnames(metadata)[4]<-"Time"


#metadata$Time[metadata$Time== "C1"]<- "T1"
#metadata$Time[metadata$Time== "C2"]<- "T2"
#metadata$Time[metadata$Time== "C3"]<- "T3"
#metadata$Time[metadata$Time== "C4"]<- "T4"
#metadata$Time[metadata$Time== "C5"]<- "T5"
#metadata$Time[metadata$Time== "C6"]<- "T6"

metadata$time_numeric<-metadata$Time

metadata$time_numeric[metadata$time_numeric== "C1"] <- "1"

metadata$time_numeric[metadata$time_numeric== "C1"]<- 1
metadata$time_numeric[metadata$time_numeric== "C2"]<- 2
metadata$time_numeric[metadata$time_numeric== "C3"]<- 3
metadata$time_numeric[metadata$time_numeric== "C4"]<- 4
metadata$time_numeric[metadata$time_numeric== "C5"]<- 5
metadata$time_numeric[metadata$time_numeric== "C6"]<- 6


# Check structure
head(metadata)
table(metadata$Time)
table(metadata$code)
metadata$sex
metadata$age
# ============================================
# MODEL 1A: Time as Categorical (compare each timepoint)
# ============================================

# Design matrix
design <- model.matrix(~ 0 + Time + sex + age, data = metadata)
colnames(design) <- gsub("Time", "C", colnames(design))

# Account for repeated measures within subjects
corfit <- duplicateCorrelation(expr_matrix, design, block = metadata$code)
cat("Intra-subject correlation:", corfit$consensus.correlation, "\n")

# Fit model
fit <- lmFit(expr_matrix, design, 
             block = metadata$code,
             correlation = corfit$consensus)

# Define contrasts (compare timepoints)
contrast_matrix <- makeContrasts(
  T2_vs_T1 = C2 - C1,
  T3_vs_T1 = C3 - C1,
  T4_vs_T1 = C4 - C1,
  T5_vs_T1 = C5 - C1,
  T6_vs_T1 = C6 - C1,
  T6_vs_T3 = C6 - C3,  # Example: compare specific timepoints
  Linear_trend = (C6 - C5 + C5 - C4 + C4 - C3 + C3 - C2 + C2 - C1)/5,  # Overall trend
  levels = design
)

# Fit contrasts
fit2 <- contrasts.fit(fit, contrast_matrix)
fit2 <- eBayes(fit2, robust = TRUE)

# Get results for each comparison
results_T2_vs_T1 <- topTable(fit2, coef="T2_vs_T1", number=Inf)
results_T6_vs_T1 <- topTable(fit2, coef="T6_vs_T1", number=Inf)

# Summary
cat("\n=== Proteins changing over time ===\n")
summary(decideTests(fit2, p.value=0.05))

# Save all results
for(i in 1:ncol(contrast_matrix)) {
  contrast_name <- colnames(contrast_matrix)[i]
  results <- topTable(fit2, coef=i, number=Inf)
  write.csv(results, paste0("Q1_proteins_", contrast_name, ".csv"))
}

# ============================================
# MODEL 1B: Time as Continuous (linear trend)
# ============================================

# Design with numeric time
design_numeric <- model.matrix(~ time_numeric + sex + age, data = metadata)

# Fit with repeated measures
corfit_numeric <- duplicateCorrelation(expr_matrix, design_numeric, 
                                       block = metadata$subject_id)
fit_numeric <- lmFit(expr_matrix, design_numeric,
                     block = metadata$subject_id,
                     correlation = corfit_numeric$consensus)
fit_numeric <- eBayes(fit_numeric, robust = TRUE)

# Results for time effect
results_time_trend <- topTable(fit_numeric, coef="time_numeric", number=Inf)

cat("\n=== Proteins with linear time trend ===\n")
cat("Significant proteins (FDR < 0.05):", 
    sum(results_time_trend$adj.P.Val < 0.05), "\n")

write.csv(results_time_trend, "Q1_proteins_linear_time_trend.csv")

# ============================================
# MODEL 1C: Time + Sex Interaction
# ============================================

# Test if time effects differ by sex
design_interaction <- model.matrix(~ 0 + time + sex + time:sex + age, 
                                   data = metadata)

corfit_int <- duplicateCorrelation(expr_matrix, design_interaction, 
                                   block = metadata$subject_id)
fit_int <- lmFit(expr_matrix, design_interaction,
                 block = metadata$subject_id,
                 correlation = corfit_int$consensus)
fit_int <- eBayes(fit_int, robust = TRUE)

# Test interaction terms
# (Proteins where time effect differs between males and females)
```

```{r}

dex<- read.csv("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/MetaData/burnoutEx.csv")
rownames(dex)<-dex$X
```

Outcome = Burnout | Predictors = Protein levels, Time, Sex, Age
Best Method: Linear Mixed Models (lme4) - one model per protein

```{r}
library(lme4)
library(lmerTest)  # For p-values

meta<- dex[c(5438:5442)]
dEX<- dex[c(6:5432)]
expr_matrix <- as.data.frame(dex[, 6:5437]) # no FDR< 0.2
```
#no adjusted p-value signofocant


Try another transformed data 
```{r}
#dEX<-dEX+1
#try with transformed data 
lEx<-sqrt(dEX+1) # final used and  lead to only 5 proteins with FDR< 0.2
```

```{r}
lEx<-log2(dEX+100) # 6 proteins FDR< 0.2
```


```{r}
#lEx[is.na(lEx)]< 0

expr_matrix <- as.data.frame(lEx)
```


```{r}
library(lme4)
library(lmerTest)
library(dplyr)

# ============================================
# STEP 1: PREPARE DATA CORRECTLY
# ============================================

# Extract metadata and protein data
meta <- dex[, 5438:5442]  # More explicit column selection
```

##make burnout value log transformation
```{r}
meta$Value<- log2(meta$Value)
```


```{r}
colnames(meta)
colnames(meta) <- c("code", "sex", "age", "Time", "Value")  # Name them clearly
colnames(meta)
# Protein expression data

cat("=== Data Structure ===\n")
cat("Samples (rows):", nrow(expr_matrix), "\n")
cat("Proteins (columns):", ncol(expr_matrix), "\n")
cat("Metadata rows:", nrow(meta), "\n\n")

# Verify dimensions match
stopifnot(nrow(expr_matrix) == nrow(meta))

# ============================================
# STEP 2: CLEAN AND PREPARE METADATA
# ============================================

# Convert to proper types
meta$code <- factor(meta$code)          # Subject ID
meta$Time <- factor(meta$Time)          # Categorical time (T1-T6)
meta$sex <- factor(meta$sex)            # Sex
meta$age <- as.numeric(meta$age)        # Age
meta$burnout <- as.numeric(meta$Value)  # Burnout score



#meta$Time[meta$Time== "C1"]<- "T1"
#meta$Time[meta$Time== "C2"]<- "T2"
#meta$Time[meta$Time== "C3"]<- "T3"
#meta$Time[meta$Time== "C4"]<- "T4"
#meta$Time[meta$Time== "C5"]<- "T5"
#meta$Time[meta$Time== "C6"]<- "T6"



# Create numeric time variable (1-6)
meta$time_numeric <- as.numeric(gsub("C", "", meta$Time))

# Check for issues
cat("=== Data Quality Checks ===\n")
cat("Missing values in burnout:", sum(is.na(meta$burnout)), "\n")
cat("Missing values in proteins:", sum(is.na(expr_matrix)), "\n")
cat("Unique subjects:", length(unique(meta$code)), "\n")
cat("Observations per subject:\n")
print(table(table(meta$code)))
cat("\nTime points:\n")
print(table(meta$Time))
cat("\nSex distribution:\n")
print(table(meta$sex))
cat("\nAge range:", range(meta$age, na.rm=TRUE), "\n\n")
```
# ============================================
# STEP 3: COMBINE INTO SINGLE DATAFRAME
# ============================================
```{r}
# ============================================
# STEP 3: COMBINE INTO SINGLE DATAFRAME
# ============================================

# Your data is ALREADY in long format!
# Each row = one sample (subject at one timepoint)
# No reshaping needed!
library(lme4)
library(lmerTest)
full_data <- cbind(meta, expr_matrix)
write.csv(full_data, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/fullData.csv")

```


```{r}
cat("Combined data dimensions:", dim(full_data), "\n\n")

# ============================================
# STEP 4: TEST ONE PROTEIN FIRST (Diagnostic)
# ============================================

cat("=== Testing first protein as example ===\n")

# Get first protein name
first_protein <- colnames(expr_matrix)[1]
cat("Testing protein:", first_protein, "\n")

# Create test dataset
test_data <- data.frame(
  subject_id = meta$code,
 # time_numeric = meta$time_numeric,
  sex = meta$sex,
  age = meta$age,
  burnout = meta$burnout,
  protein_expression = expr_matrix[, first_protein]
)

# Check for complete cases
cat("Complete cases:", sum(complete.cases(test_data)), "/", nrow(test_data), "\n")


library(lme4)
library(lmerTest)
library(dplyr)

# Fit test model
test_model <- lmer(burnout ~ protein_expression + sex + age + 
                     (1 | subject_id), #time_numeric +
                   data = test_data,
                   REML = FALSE)

cat("\nTest model summary:\n")
print(summary(test_model))

# ============================================
# STEP 5: IMPROVED FUNCTION TO TEST PROTEINS
# ============================================

test_protein_burnout <- function(protein_name, expr_matrix, meta) {
  
  # Create data for this protein
  protein_data <- data.frame(
    subject_id = meta$code,
    #time_numeric = meta$time_numeric,
    sex = meta$sex,
    age = meta$age,
    burnout = meta$burnout,
    protein_expression = expr_matrix[, protein_name],  # FIXED: column subset
    stringsAsFactors = FALSE
  )
  
  # Remove rows with missing values
  protein_data <- protein_data[complete.cases(protein_data), ]
  
  # Check if enough data
  if(nrow(protein_data) < 20 || 
     length(unique(protein_data$subject_id)) < 5) {
    return(data.frame(
      estimate = NA, se = NA, t_value = NA, p_value = NA,
      n_obs = nrow(protein_data), n_subjects = length(unique(protein_data$subject_id)),
      error = "Insufficient data"
    ))
  }
  
  # Fit model
  tryCatch({
    model <- lmer(burnout ~ protein_expression +  sex + age + 
                    (1 | subject_id), #time_numeric +
                  data = protein_data,
                  REML = FALSE,
                  control = lmerControl(optimizer = "bobyqa"))
    
    # Extract coefficients
    coefs <- coef(summary(model))
    protein_row <- coefs["protein_expression", ]
    
    result <- data.frame(
      estimate = protein_row["Estimate"],
      se = protein_row["Std. Error"],
      t_value = protein_row["t value"],
      p_value = protein_row["Pr(>|t|)"],
      n_obs = nrow(protein_data),
      n_subjects = length(unique(protein_data$subject_id)),
      error = NA
    )
    
    return(result)
    
  }, error = function(e) {
    return(data.frame(
      estimate = NA, se = NA, t_value = NA, p_value = NA,
      n_obs = nrow(protein_data), 
      n_subjects = length(unique(protein_data$subject_id)),
      error = as.character(e$message)
    ))
  })
}

# ============================================
# STEP 6: RUN FOR ALL PROTEINS
# ============================================

cat("\n=== Running analysis for all proteins ===\n")
cat("Testing", ncol(expr_matrix), "proteins...\n\n")

# Initialize results dataframe
results_q2 <- data.frame(
  protein = colnames(expr_matrix),  # FIXED: colnames not rownames
  estimate = NA,
  se = NA,
  t_value = NA,
  p_value = NA,
  n_obs = NA,
  n_subjects = NA,
  error = NA,
  stringsAsFactors = FALSE
)

# Progress bar
pb <- txtProgressBar(min = 0, max = ncol(expr_matrix), style = 3)

# Loop through proteins
for(i in 1:ncol(expr_matrix)) {
  
  protein_name <- colnames(expr_matrix)[i]
  
  # Test this protein
  result <- test_protein_burnout(protein_name, expr_matrix, meta)
  
  # Store results
  results_q2[i, 2:8] <- result
  
  setTxtProgressBar(pb, i)
}

close(pb)

# ============================================
# STEP 7: PROCESS RESULTS
# ============================================

cat("\n\n=== Processing Results ===\n")

# Check for errors
n_errors <- sum(!is.na(results_q2$error))
cat("Models with errors:", n_errors, "\n")
if(n_errors > 0) {
  cat("Error types:\n")
  print(table(results_q2$error, useNA = "ifany"))
}

# Remove failed models
results_q2_clean <- results_q2[is.na(results_q2$error) & !is.na(results_q2$p_value), ]
cat("Successful models:", nrow(results_q2_clean), "\n\n")

# Multiple testing correction
results_q2_clean$adj_p_value <- p.adjust(results_q2_clean$p_value, method = "BH")

# Sort by significance
results_q2_clean <- results_q2_clean[order(results_q2_clean$adj_p_value), ]

# ============================================
# STEP 8: SUMMARY STATISTICS
# ============================================

cat("=== Results Summary ===\n")
cat("Total proteins tested:", nrow(results_q2_clean), "\n")
cat("Significant (FDR <0.2):", 
    sum(results_q2_clean$adj_p_value < 0.2, na.rm=TRUE), "\n")
cat("Significant (FDR < 0.20):", 
    sum(results_q2_clean$adj_p_value < 0.10, na.rm=TRUE), "\n")
cat("Nominal p < 0.05:", 
    sum(results_q2_clean$p_value < 0.05, na.rm=TRUE), "\n\n")

# Direction of effects
sig_proteins <- results_q2_clean[results_q2_clean$adj_p_value < 0.2, ]
if(nrow(sig_proteins) > 0) {
  cat("Direction of significant effects:\n")
  cat("  Positive (higher protein → higher burnout):", 
      sum(sig_proteins$estimate > 0), "\n")
  cat("  Negative (higher protein → lower burnout):", 
      sum(sig_proteins$estimate < 0), "\n\n")
}


```
# ============================================
# STEP 9: VISUALIZATIONS
# ============================================

```{r}
pdf("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/outcomeBurnoutV3.pdf", width = 8, height = 8)
# Histogram of p-values
par(mfrow=c(2,2))

hist(results_q2_clean$p_value, breaks=50, 
     main="Distribution of P-values",
     xlab="P-value", col="lightblue")
abline(v=0.05, col="red", lwd=2, lty=2)

# Q-Q plot of p-values
qqplot(-log10(1:nrow(results_q2_clean) / nrow(results_q2_clean)),
       -log10(sort(results_q2_clean$p_value)),
       xlab="Expected -log10(p)", ylab="Observed -log10(p)",
       main="Q-Q Plot of P-values")
abline(0, 1, col="red", lwd=2)

# Volcano plot
# Volcano plot with labels for top 20 proteins
plot(results_q2_clean$estimate, 
     -log10(results_q2_clean$p_value),
     pch=20, cex=0.5,
     xlab="Effect Size (Estimate)",
     ylab="-log10(P-value)",
     main="Volcano Plot",
     col=ifelse(results_q2_clean$adj_p_value < 0.2, "red", "gray"))
abline(h=-log10(0.05), col="blue", lty=2)
abline(v=c(-0.5, 0.5), col="blue", lty=2)  # Add effect size thresholds

# Add labels for top 20 most significant proteins
top_n <- 10
top_proteins <- head(results_q2_clean, top_n)

#text(top_proteins$estimate,
 #    -log10(top_proteins$adj_p_value),
    # labels = top_proteins$protein,
     #cex = 0.6,
     #pos = 4,  # Position: 1=below, 2=left, 3=above, 4=right
     #col = "darkred")

legend("bottomright", 
       c("FDR < 0.2", "Not significant"), 
       col=c("red", "gray"), pch=19)

# Effect sizes
hist(results_q2_clean$estimate, breaks=50,
     main="Distribution of Effect Sizes",
     xlab="Estimate", col="lightgreen")
abline(v=0, col="red", lwd=2)

par(mfrow=c(1,1))
dev.off()
# ============================================
# STEP 10: SAVE RESULTS
# ============================================
```


```{r}
# Save full results
write.csv(results_q2_clean, 
          "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Q2_proteins_Final.csv", row.names=FALSE)
```


#for protrin as outcome 
```{r}
lEx<-sqrt(dEX)
```

```{r}
#lEx[is.na(lEx)]< 0

expr_matrix <- as.data.frame(lEx)
```

```{r}
# MODEL FORMULA:
Protein_i ~ Burnout + Time + Sex + Age + (1 | Subject_ID)

# Where:
# ✓ Outcome (Dependent Variable): Protein expression level
# ✓ Fixed Effects (Independent Variables):
#   - Burnout (continuous)
#   - Time_numeric (continuous, 1-6)
#   - Sex (factor)
#   - Age (continuous)
# ✓ Random Effect: (1 | subject_id) = random intercept per subject
```

```{r}
library(lme4)
library(lmerTest)

# ============================================
# CLARIFIED MODEL STRUCTURE
# ============================================

test_burnout_to_protein <- function(protein_name, expr_matrix, meta) {
  
  # ========================================
  # STEP 1: Build dataset for THIS protein
  # ========================================
  protein_data <- data.frame(
    # RANDOM EFFECT:
    subject_id = meta$code,              # Subject identifier (random effect)
    
    # FIXED EFFECTS (INDEPENDENT VARIABLES):
    burnout = meta$burnout,              # Predictor 1: Burnout score
    #time_numeric = meta$time_numeric,    # Predictor 2: Time (1-6)
    sex = meta$sex,                      # Predictor 3: Sex (M/F)
    age = meta$age,                      # Predictor 4: Age (years)
    
    # DEPENDENT VARIABLE (OUTCOME):
    protein_expression = expr_matrix[, protein_name],  # THIS protein's levels
    
    stringsAsFactors = FALSE
  )
  
  # Remove missing data
  protein_data <- protein_data[complete.cases(protein_data), ]
  
  # Quality check
  if(nrow(protein_data) < 20 || length(unique(protein_data$subject_id)) < 5) {
    return(data.frame(
      estimate = NA, se = NA, t_value = NA, p_value = NA,
      n_obs = nrow(protein_data), n_subjects = length(unique(protein_data$subject_id)),
      error = "Insufficient data"
    ))
  }
  
  # ========================================
  # STEP 2: Fit Linear Mixed Model
  # ========================================
  # MODEL EQUATION:
  # protein_expression_ij = β0 + β1*burnout_ij + β2*time_ij + 
  #                         β3*sex_i + β4*age_i + u_i + ε_ij
  #
  # Where:
  # - i = subject index
  # - j = timepoint index (1-6)
  # - u_i = random intercept for subject i ~ N(0, σ²_u)
  # - ε_ij = residual error ~ N(0, σ²_ε)
  # ========================================
  
  tryCatch({
    model <- lmer(
      # OUTCOME ~ PREDICTORS + RANDOM EFFECT
      protein_expression ~ burnout + sex + age + (1 | subject_id), #time_numeric
      #        ↑              ↑         ↑           ↑     ↑        ↑
      #    OUTCOME        PREDICTOR  PREDICTOR   PRED  PRED   RANDOM EFFECT
      #    (varies by    OF INTEREST (control)  (ctrl)(ctrl) (accounts for
      #     protein)                                          repeated measures)
      
      data = protein_data,
      REML = FALSE,
      control = lmerControl(optimizer = "bobyqa")
    )
    
    # ========================================
    # STEP 3: Extract BURNOUT effect
    # ========================================
    coefs <- coef(summary(model))
    burnout_effect <- coefs["burnout", ]  # Effect of burnout on THIS protein
    
    result <- data.frame(
      estimate = burnout_effect["Estimate"],      # β1: Burnout effect
      se = burnout_effect["Std. Error"],         # Standard error of β1
      t_value = burnout_effect["t value"],       # t-statistic
      p_value = burnout_effect["Pr(>|t|)"],      # p-value for burnout effect
      n_obs = nrow(protein_data),
      n_subjects = length(unique(protein_data$subject_id)),
      error = NA
    )
    
    return(result)
    
  }, error = function(e) {
    return(data.frame(
      estimate = NA, se = NA, t_value = NA, p_value = NA,
      n_obs = nrow(protein_data), 
      n_subjects = length(unique(protein_data$subject_id)),
      error = as.character(e$message)
    ))
  })
}

# ============================================
# DEMONSTRATION: Test One Protein First
# ============================================

cat("=== DEMONSTRATION: Testing One Protein ===\n\n")

# Pick first protein as example
example_protein <- colnames(expr_matrix)[1]

cat("Testing protein:", example_protein, "\n")
cat("\nModel formula:\n")
cat("  ", example_protein, " ~ burnout + time_numeric + sex + age + (1|subject_id)\n\n")

# Build data
demo_data <- data.frame(
  subject_id = meta$code,
  burnout = meta$burnout,
  #time_numeric = meta$time_numeric,
  sex = meta$sex,
  age = meta$age,
  protein_expression = expr_matrix[, example_protein]
)

demo_data <- demo_data[complete.cases(demo_data), ]

cat("Data structure for this protein:\n")
cat("  Observations:", nrow(demo_data), "\n")
cat("  Unique subjects:", length(unique(demo_data$subject_id)), "\n")
cat("  Obs per subject:", table(table(demo_data$subject_id)), "\n\n")

# Fit model
demo_model <- lmer(protein_expression ~ burnout + sex + age +  #time_numeric 
                     (1 | subject_id),
                   data = demo_data,
                   REML = FALSE)

cat("Model summary:\n")
print(summary(demo_model))

cat("\n\nCoefficients interpretation:\n")
coefs <- coef(summary(demo_model))
print(coefs)

cat("\n\nBURNOUT EFFECT (what we're testing):\n")
cat("  Estimate:", coefs["burnout", "Estimate"], "\n")
cat("  → For every 1-unit increase in burnout,", 
    example_protein, "changes by", 
    round(coefs["burnout", "Estimate"], 4), "units\n")
cat("  P-value:", coefs["burnout", "Pr(>|t|)"], "\n")
cat("  → Is this effect significant?", 
    ifelse(coefs["burnout", "Pr(>|t|)"] < 0.05, "YES", "NO"), "\n\n")

# ============================================
# NOW RUN FOR ALL PROTEINS
# ============================================

cat("=== Running for ALL proteins ===\n")
cat("Each protein will be tested with the same model structure:\n")
cat("  Protein_i ~ burnout + time_numeric + sex + age + (1|subject_id)\n\n")

results_q2Out <- data.frame(
  protein = colnames(expr_matrix),
  estimate = NA,
  se = NA,
  t_value = NA,
  p_value = NA,
  n_obs = NA,
  n_subjects = NA,
  error = NA,
  stringsAsFactors = FALSE
)

# Progress bar
pb <- txtProgressBar(min = 0, max = ncol(expr_matrix), style = 3)

for(i in 1:ncol(expr_matrix)) {
  protein_name <- colnames(expr_matrix)[i]
  result <- test_burnout_to_protein(protein_name, expr_matrix, meta)
  results_q2Out[i, 2:8] <- result
  setTxtProgressBar(pb, i)
}

close(pb)

# Process results
results_q2Out_clean <- results_q2Out[is.na(results_q2Out$error) & 
                                     !is.na(results_q2Out$p_value), ]
results_q2Out_clean$adj_p_value <- p.adjust(results_q2Out_clean$p_value, 
                                             method = "BH")
results_q2Out_clean <- results_q2Out_clean[order(results_q2Out_clean$adj_p_value), ]

cat("\n\n=== RESULTS SUMMARY ===\n")
cat("Total proteins tested:", nrow(results_q2Out_clean), "\n")
cat("Significant (FDR < 0.2):", 
    sum(results_q2Out_clean$adj_p_value < 0.2, na.rm=TRUE), "\n\n")

cat("Top 10 proteins affected by burnout:\n")
print(head(results_q2Out_clean[, c("protein", "estimate", "p_value", "adj_p_value")], 10))
```

#categorize and visualize the proteins based on their direction of association with burnout.
```{r}
#dx<- read.csv("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Q2_proteins_FinalGeneID.csv")


results_q2_clean<-read.csv( "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Q2_proteins_Final.csv", row.names=FALSE)
```


```{r}
library(ggplot2)
library(ggrepel)
library(dplyr)

# ============================================
# STEP 1: CATEGORIZE SIGNIFICANT PROTEINS
# ============================================

# Filter significant proteins (FDR < 0.2)
sig_proteins <- results_q2_clean[results_q2_clean$adj_p_value < 0.2, ]

cat("=== Categorizing Significant Proteins ===\n")
cat("Total significant proteins (FDR < 0.2):", nrow(sig_proteins), "\n\n")

# Create categories based on direction
sig_proteins$category <- NA

# POSITIVE ASSOCIATION (estimate > 0)
# Higher protein → Higher burnout (and vice versa)
sig_proteins$category[sig_proteins$estimate > 0] <- "Positive Association\n(↑ Protein → ↑ Burnout)"

# NEGATIVE ASSOCIATION (estimate < 0)  
# Higher protein → Lower burnout (inverse relationship)
sig_proteins$category[sig_proteins$estimate < 0] <- "Negative Association\n(↑ Protein → ↓ Burnout)"

# Create 4 subcategories for detailed interpretation
sig_proteins$detailed_category <- NA

# Category 1: Positive estimate - protein increases WITH burnout
sig_proteins$detailed_category[sig_proteins$estimate > 0] <- 
  "1. Increases with Burnout\n(Burnout ↑ → Protein ↑)"

# Category 2: Negative estimate - protein decreases WITH burnout  
sig_proteins$detailed_category[sig_proteins$estimate < 0] <- 
  "2. Decreases with Burnout\n(Burnout ↑ → Protein ↓)"

# Summary table
summary_table <- data.frame(
  Category = c(
    "Positive Association (↑ Burnout → ↑ Protein)",
    "Negative Association (↑ Burnout → ↓ Protein)"
  ),
  Count = c(
    sum(sig_proteins$estimate > 0),
    sum(sig_proteins$estimate < 0)
  ),
  Mean_Effect = c(
    mean(sig_proteins$estimate[sig_proteins$estimate > 0]),
    mean(sig_proteins$estimate[sig_proteins$estimate < 0])
  )
)

cat("Summary of Associations:\n")
print(summary_table)
cat("\n")

# ============================================
# STEP 2: SEPARATE INTO LISTS
# ============================================

# Positive association: Proteins that INCREASE with burnout
proteins_increase_with_burnout <- sig_proteins[sig_proteins$estimate > 0, ]
proteins_increase_with_burnout <- proteins_increase_with_burnout[
  order(proteins_increase_with_burnout$adj_p_value), ]

# Negative association: Proteins that DECREASE with burnout
proteins_decrease_with_burnout <- sig_proteins[sig_proteins$estimate < 0, ]
proteins_decrease_with_burnout <- proteins_decrease_with_burnout[
  order(proteins_decrease_with_burnout$adj_p_value), ]

cat("=== POSITIVE ASSOCIATION ===\n")
cat("Proteins that INCREASE when burnout INCREASES:\n")
cat("Count:", nrow(proteins_increase_with_burnout), "\n")
cat("\nTop 10:\n")
print(proteins_increase_with_burnout[1:min(10, nrow(proteins_increase_with_burnout)), 
                                     c("protein", "estimate", "p_value", "adj_p_value")])

cat("\n\n=== NEGATIVE ASSOCIATION ===\n")
cat("Proteins that DECREASE when burnout INCREASES:\n")
cat("Count:", nrow(proteins_decrease_with_burnout), "\n")
cat("\nTop 10:\n")
print(proteins_decrease_with_burnout[1:min(10, nrow(proteins_decrease_with_burnout)), 
                                     c("protein", "estimate", "p_value", "adj_p_value")])

# Save separate files
write.csv(proteins_increase_with_burnout, 
          "Proteins_INCREASE_with_Burnout.csv", 
          row.names = FALSE)
write.csv(proteins_decrease_with_burnout, 
          "Proteins_DECREASE_with_Burnout.csv", 
          row.names = FALSE)

# ============================================
# VISUALIZATION 1: ENHANCED VOLCANO PLOT
# ============================================

# Add category for all proteins (including non-significant)
results_q2_clean$direction_category <- "Not Significant"

results_q2_clean$direction_category[
  results_q2_clean$adj_p_value < 0.2 & results_q2_clean$estimate > 0
] <- "↑ Protein with ↑ Burnout"

results_q2_clean$direction_category[
  results_q2_clean$adj_p_value < 0.2 & results_q2_clean$estimate < 0
] <- "↓ Protein with ↑ Burnout"

# Label top proteins from each category
results_q2_clean$protein_label <- ""

# Top 10 positive
top_positive <- head(proteins_increase_with_burnout, 10)
results_q2_clean$protein_label[
  results_q2_clean$protein %in% top_positive$protein
] <- results_q2_clean$protein[
  results_q2_clean$protein %in% top_positive$protein
]

# Top 10 negative
top_negative <- head(proteins_decrease_with_burnout, 10)
results_q2_clean$protein_label[
  results_q2_clean$protein %in% top_negative$protein
] <- results_q2_clean$protein[
  results_q2_clean$protein %in% top_negative$protein
]

# Create volcano plot
p1 <- ggplot(results_q2_clean, 
             aes(x = estimate, 
                 y = -log10(p_value), 
                 color = direction_category,
                 label = protein_label)) +
  geom_point(alpha = 0.6, size = 2) +
  geom_hline(yintercept = -log10(0.05), 
             linetype = "dashed", color = "gray30", linewidth = 0.8) +
  geom_vline(xintercept = 0, 
             linetype = "solid", color = "black", linewidth = 0.5) +
  geom_text_repel(
    size = 3,
    max.overlaps = 25,
    box.padding = 0.5,
    segment.color = "gray50",
    fontface = "bold"
  ) +
  scale_color_manual(
    values = c(
      "↑ Protein with ↑ Burnout" = "#D32F2F",      # Red
      "↓ Protein with ↑ Burnout" = "#1976D2",      # Blue
      "Not Significant" = "#BDBDBD"                 # Gray
    )
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  ) +
  labs(
    title = "Protein-Burnout Associations",
    subtitle = sprintf("Positive: n=%d | Negative: n=%d", 
                      nrow(proteins_increase_with_burnout),
                      nrow(proteins_decrease_with_burnout)),
    x = "Effect Size: Protein Effect on Burnout\n← Lower Burnout | Higher Burnout →",
    y = "-log10(P-value)",
    color = "Association Type"
  ) +
  annotate("rect", 
           xmin = 0, xmax = Inf, ymin = -log10(0.05), ymax = Inf,
           alpha = 0.1, fill = "red") +
  annotate("rect", 
           xmin = -Inf, xmax = 0, ymin = -log10(0.05), ymax = Inf,
           alpha = 0.1, fill = "blue")

print(p1)
ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Details_posneg/Volcano_Directional_Categories.png", p1, width = 12, height = 9, dpi = 600)

# ============================================
# VISUALIZATION 2: BAR PLOT OF CATEGORIES
# ============================================

category_counts <- data.frame(
  Category = c("Increase with\nBurnout ↑", "Decrease with\nBurnout ↑"),
  Count = c(nrow(proteins_increase_with_burnout), 
            nrow(proteins_decrease_with_burnout)),
  Direction = c("Positive", "Negative")
)

p2 <- ggplot(category_counts, aes(x = Category, y = Count, fill = Direction)) +
  geom_bar(stat = "identity", width = 0.6) +
  geom_text(aes(label = Count), vjust = -0.5, size = 6, fontface = "bold") +
  scale_fill_manual(values = c("Positive" = "#D32F2F", "Negative" = "#1976D2")) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none",
    panel.grid.major.x = element_blank()
  ) +
  labs(
    title = "Distribution of Protein-Burnout Associations",
    subtitle = "Significant proteins (FDR < 0.2)",
    x = "",
    y = "Number of Proteins"
  ) +
  ylim(0, max(category_counts$Count) * 1.2)

print(p2)
ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Details_posneg/BarPlot_Categories.png", p2, width = 8, height = 6, dpi = 300)

# ============================================
# VISUALIZATION 3: EFFECT SIZE DISTRIBUTION
# ============================================

p3 <- ggplot(sig_proteins, aes(x = estimate, fill = category)) +
  geom_histogram(bins = 30, alpha = 0.7, color = "black") +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 1) +
  scale_fill_manual(
    values = c(
      "Positive Association\n(↑ Protein → ↑ Burnout)" = "#D32F2F",
      "Negative Association\n(↑ Protein → ↓ Burnout)" = "#1976D2"
    )
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom") +
  labs(
    title = "Distribution of Effect Sizes by Association Type",
    subtitle = "Significant proteins (FDR < 0.2)",
    x = "Effect Size (Estimate)",
    y = "Number of Proteins",
    fill = "Association Type"
  )

print(p3)
ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Details_posneg/Histogram_EffectSizes.png", p3, width = 10, height = 6, dpi = 300)

# ============================================
# VISUALIZATION 4: SPLIT VOLCANO PLOTS
# ============================================

par(mfrow = c(1, 2))

# Plot 1: Positive associations (RED)
plot(proteins_increase_with_burnout$estimate, 
     -log10(proteins_increase_with_burnout$p_value),
     pch = 20, cex = 1.5, col = "#D32F2F",
     xlim = range(sig_proteins$estimate),
     ylim = c(0, max(-log10(sig_proteins$p_value)) * 1.1),
     xlab = "Effect Size",
     ylab = "-log10(P-value)",
     main = paste0("POSITIVE Association (n=", 
                   nrow(proteins_increase_with_burnout), ")\n",
                   "↑ Protein → ↑ Burnout"))
abline(h = -log10(0.05), lty = 2, col = "gray50")
abline(v = 0, lty = 1, col = "black")

# Add labels for top 5
if(nrow(proteins_increase_with_burnout) >= 5) {
  top5_pos <- head(proteins_increase_with_burnout, 5)
  text(top5_pos$estimate, -log10(top5_pos$p_value),
       labels = top5_pos$protein, pos = 4, cex = 0.7, col = "darkred")
}

# Plot 2: Negative associations (BLUE)
plot(proteins_decrease_with_burnout$estimate, 
     -log10(proteins_decrease_with_burnout$p_value),
     pch = 20, cex = 1.5, col = "#1976D2",
     xlim = range(sig_proteins$estimate),
     ylim = c(0, max(-log10(sig_proteins$p_value)) * 1.1),
     xlab = "Effect Size",
     ylab = "-log10(P-value)",
     main = paste0("NEGATIVE Association (n=", 
                   nrow(proteins_decrease_with_burnout), ")\n",
                   "↑ Burnout → ↓ Protein"))
abline(h = -log10(0.05), lty = 2, col = "gray50")
abline(v = 0, lty = 1, col = "black")

# Add labels for top 5
if(nrow(proteins_decrease_with_burnout) >= 5) {
  top5_neg <- head(proteins_decrease_with_burnout, 5)
  text(top5_neg$estimate, -log10(top5_neg$p_value),
       labels = top5_neg$protein, pos = 2, cex = 0.7, col = "darkblue")
}

par(mfrow = c(1, 1))

# Save as PDF
pdf("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Details_posneg/SplitVolcano_Positive_Negative.pdf", width = 14, height = 7)
par(mfrow = c(1, 2))
# (repeat the two plots above)
dev.off()

# ============================================
# VISUALIZATION 5: TOP PROTEINS HEATMAP STYLE
# ============================================

library(ggplot2)

# Combine top 20 from each category
top_positive_20 <- head(proteins_increase_with_burnout, 20)
top_negative_20 <- head(proteins_decrease_with_burnout, 20)

combined_top <- rbind(
  data.frame(top_positive_20, Direction = "Positive"),
  data.frame(top_negative_20, Direction = "Negative")
)

# Order by estimate within each group
combined_top <- combined_top[order(combined_top$Direction, -combined_top$estimate), ]
combined_top$protein <- factor(combined_top$protein, 
                               levels = combined_top$protein)

p5 <- ggplot(combined_top, 
             aes(x = Direction, y = protein, fill = estimate)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = sprintf("%.3f", estimate)), 
            size = 2.5, fontface = "bold") +
  scale_fill_gradient2(
    low = "#1976D2", mid = "white", high = "#D32F2F",
    midpoint = 0,
    name = "Effect Size"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    axis.text.y = element_text(size = 8),
    panel.grid = element_blank()
  ) +
  labs(
    title = "Top 20 Proteins in Each Direction",
    x = "Association Direction",
    y = "Protein"
  )

print(p5)
ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Details_posneg/Heatmap_Top_Proteins.png", p5, width = 8, height = 12, dpi = 300)

# ============================================
# VISUALIZATION 6: LOLLIPOP CHART
# ============================================

# Top 15 from each
top15_pos <- head(proteins_increase_with_burnout, 15)
top15_neg <- head(proteins_decrease_with_burnout, 15)

combined_top15 <- rbind(
  data.frame(top15_pos, Direction = "Positive"),
  data.frame(top15_neg, Direction = "Negative")
)

combined_top15 <- combined_top15[order(combined_top15$estimate), ]
combined_top15$protein <- factor(combined_top15$protein, 
                                 levels = combined_top15$protein)

p6 <- ggplot(combined_top15, 
             aes(x = estimate, y = protein, color = Direction)) +
  geom_segment(aes(x = 0, xend = estimate, y = protein, yend = protein),
               linewidth = 1.2) +
  geom_point(size = 4) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  scale_color_manual(values = c("Positive" = "#D32F2F", 
                                 "Negative" = "#1976D2")) +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank()
  ) +
  labs(
    title = "Top 15 Proteins in Each Direction",
    subtitle = "Effect of protein on burnout",
    x = "Effect Size (Estimate)\n← Lower Burnout | Higher Burnout →",
    y = "",
    color = "Direction"
  )

print(p6)
ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Details_posneg/Lollipop_Top_Proteins.png", p6, width = 10, height = 10, dpi = 300)

# ============================================
# SUMMARY REPORT
# ============================================

cat("\n\n========================================\n")
cat("SUMMARY REPORT: PROTEIN-BURNOUT ASSOCIATIONS\n")
cat("========================================\n\n")

cat("POSITIVE ASSOCIATIONS (Protein ↑ → Burnout ↑):\n")
cat("  Total:", nrow(proteins_increase_with_burnout), "proteins\n")
cat("  Mean effect:", round(mean(proteins_increase_with_burnout$estimate), 4), "\n")
cat("  Range:", round(range(proteins_increase_with_burnout$estimate), 4), "\n\n")

cat("NEGATIVE ASSOCIATIONS (Protein ↑ → Burnout ↓):\n")
cat("  Total:", nrow(proteins_decrease_with_burnout), "proteins\n")
cat("  Mean effect:", round(mean(proteins_decrease_with_burnout$estimate), 4), "\n")
cat("  Range:", round(range(proteins_decrease_with_burnout$estimate), 4), "\n\n")

cat("FILES SAVED:\n")
cat("  - Proteins_INCREASE_with_Burnout.csv\n")
cat("  - Proteins_DECREASE_with_Burnout.csv\n")
cat("  - Volcano_Directional_Categories.png\n")
cat("  - BarPlot_Categories.png\n")
cat("  - Histogram_EffectSizes.png\n")
cat("  - Heatmap_Top_Proteins.png\n")
cat("  - Lollipop_Top_Proteins.png\n\n")
```

```{r}
library(ggplot2)
library(ggrepel)
library(dplyr)
library(gridExtra)

# ============================================
# STEP 1: CREATE 4 CATEGORIES
# ============================================

# Filter significant proteins
sig_proteins <- results_q2_clean[results_q2_clean$adj_p_value < 0.2, ]

cat("=== Creating 4-Category Classification ===\n")
cat("Total significant proteins (FDR < 0.2):", nrow(sig_proteins), "\n\n")

# POSITIVE ASSOCIATION (estimate > 0)
# Both interpretations are TRUE for the same proteins:

# Category 1 & 2: SAME PROTEINS (Positive correlation)
category_1_2 <- sig_proteins[sig_proteins$estimate > 0, ]
category_1_2$category_main <- "POSITIVE Association"
category_1_2$category_1 <- "1. Protein ↑ when Burnout ↑"
category_1_2$category_2 <- "2. Protein ↓ when Burnout ↓"
category_1_2$interpretation <- "Protein and Burnout move TOGETHER"

# NEGATIVE ASSOCIATION (estimate < 0)
# Both interpretations are TRUE for the same proteins:

# Category 3 & 4: SAME PROTEINS (Negative correlation)
category_3_4 <- sig_proteins[sig_proteins$estimate < 0, ]
category_3_4$category_main <- "NEGATIVE Association"
category_3_4$category_1 <- "3. Protein ↓ when Burnout ↑"
category_3_4$category_2 <- "4. Protein ↑ when Burnout ↓"
category_3_4$interpretation <- "Protein and Burnout move OPPOSITE"

# Combine
sig_proteins_categorized <- rbind(category_1_2, category_3_4)

# ============================================
# SUMMARY TABLE - 4 CATEGORIES
# ============================================

summary_4cat <- data.frame(
  Category = c(
    "1. Protein ↑ when Burnout ↑",
    "2. Protein ↓ when Burnout ↓",
    "3. Protein ↓ when Burnout ↑",
    "4. Protein ↑ when Burnout ↓"
  ),
  Relationship = c(
    "Positive (same proteins as #2)",
    "Positive (same proteins as #1)",
    "Negative (same proteins as #4)",
    "Negative (same proteins as #3)"
  ),
  Count = c(
    nrow(category_1_2),
    nrow(category_1_2),  # Same count
    nrow(category_3_4),
    nrow(category_3_4)   # Same count
  ),
  Unique_Proteins = c(
    nrow(category_1_2),
    nrow(category_1_2),
    nrow(category_3_4),
    nrow(category_3_4)
  )
)

cat("=== 4-Category Summary ===\n")
print(summary_4cat)
cat("\n")

cat("Note: Categories 1 & 2 represent THE SAME proteins (positive correlation)\n")
cat("      Categories 3 & 4 represent THE SAME proteins (negative correlation)\n\n")

# ============================================
# SEPARATE LISTS FOR EACH INTERPRETATION
# ============================================

# Category 1 & 2: Proteins that increase WITH burnout
cat1_proteins <- category_1_2[order(category_1_2$adj_p_value), ]
write.csv(cat1_proteins, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Category1_2_Protein_UP_Burnout_UP.csv", row.names = FALSE)

cat("\n=== CATEGORY 1 & 2 (POSITIVE ASSOCIATION) ===\n")
cat("Interpretation 1: Protein ↑ when Burnout ↑\n")
cat("Interpretation 2: Protein ↓ when Burnout ↓\n")
cat("Count:", nrow(cat1_proteins), "\n")
cat("Mean effect:", round(mean(cat1_proteins$estimate), 4), "\n\n")
cat("Top 10:\n")
print(cat1_proteins[1:min(10, nrow(cat1_proteins)), 
                    c("protein", "estimate", "p_value", "adj_p_value")])

# Category 3 & 4: Proteins that decrease WITH burnout
cat3_proteins <- category_3_4[order(category_3_4$adj_p_value), ]
write.csv(cat3_proteins, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Category3_4_Protein_DOWN_Burnout_UP.csv", row.names = FALSE)

cat("\n\n=== CATEGORY 3 & 4 (NEGATIVE ASSOCIATION) ===\n")
cat("Interpretation 3: Protein ↓ when Burnout ↑\n")
cat("Interpretation 4: Protein ↑ when Burnout ↓\n")
cat("Count:", nrow(cat3_proteins), "\n")
cat("Mean effect:", round(mean(cat3_proteins$estimate), 4), "\n\n")
cat("Top 10:\n")
print(cat3_proteins[1:min(10, nrow(cat3_proteins)), 
                    c("protein", "estimate", "p_value", "adj_p_value")])

# ============================================
# VISUALIZATION 1: 4-QUADRANT VOLCANO PLOT
# ============================================

# Prepare labels
results_q2_clean$category_4way <- "Not Significant"

results_q2_clean$category_4way[
  results_q2_clean$adj_p_value < 0.2 & results_q2_clean$estimate > 0
] <- "Cat 1&2: Protein ↑↔ Burnout ↑"

results_q2_clean$category_4way[
  results_q2_clean$adj_p_value < 0.2 & results_q2_clean$estimate < 0
] <- "Cat 3&4: Protein ↓↔ Burnout ↑"

# Label top proteins
results_q2_clean$protein_label <- ""
top10_pos <- head(cat1_proteins, 10)
top10_neg <- head(cat3_proteins, 10)

results_q2_clean$protein_label[results_q2_clean$protein %in% top10_pos$protein] <- 
  results_q2_clean$protein[results_q2_clean$protein %in% top10_pos$protein]
results_q2_clean$protein_label[results_q2_clean$protein %in% top10_neg$protein] <- 
  results_q2_clean$protein[results_q2_clean$protein %in% top10_neg$protein]

# Create volcano plot
p1 <- ggplot(results_q2_clean, 
             aes(x = estimate, 
                 y = -log10(p_value), 
                 color = category_4way,
                 label = protein_label)) +
  geom_point(alpha = 0.6, size = 2.5) +
  geom_hline(yintercept = -log10(0.05), 
             linetype = "dashed", color = "gray30", linewidth = 0.8) +
  geom_vline(xintercept = 0, 
             linetype = "solid", color = "black", linewidth = 0.8) +
  
  # Add quadrant labels
  annotate("text", x = max(results_q2_clean$estimate, na.rm=TRUE) * 0.7, 
           y = max(-log10(results_q2_clean$p_value), na.rm=TRUE) * 0.95,
           label = paste0("Category 1&2\n", 
                         "Protein ↑ when Burnout ↑\n",
                         "Protein ↓ when Burnout ↓\n",
                         "(n=", nrow(cat1_proteins), ")"),
           size = 4, fontface = "bold", color = "#D32F2F") +
  
  annotate("text", x = min(results_q2_clean$estimate, na.rm=TRUE) * 0.7, 
           y = max(-log10(results_q2_clean$p_value), na.rm=TRUE) * 0.95,
           label = paste0("Category 3&4\n",
                         "Protein ↓ when Burnout ↑\n",
                         "Protein ↑ when Burnout ↓\n",
                         "(n=", nrow(cat3_proteins), ")"),
           size = 4, fontface = "bold", color = "#1976D2") +
  
  geom_text_repel(
    size = 3,
    max.overlaps = 25,
    box.padding = 0.5,
    segment.color = "gray50",
    fontface = "bold"
  ) +
  
  scale_color_manual(
    values = c(
      "Cat 1&2: Protein ↑↔ Burnout ↑" = "#D32F2F",
      "Cat 3&4: Protein ↓↔ Burnout ↑" = "#1976D2",
      "Not Significant" = "#BDBDBD"
    )
  ) +
  
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    legend.text = element_text(size = 10)
  ) +
  
  labs(
    title = "Protein-Burnout Associations: 4-Way Interpretation",
    subtitle = sprintf("Total Significant: %d proteins (FDR < 0.2)", nrow(sig_proteins)),
    x = "Effect Size: Protein's Effect on Burnout",
    y = "-log10(P-value)",
    color = "Category"
  )

print(p1)
ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Volcano_4Categories.pdf", p1, width = 14, height = 10, dpi = 600)

# ============================================
# VISUALIZATION 2: 4-BAR CHART
# ============================================

# Create data for 4 bars
bar_data <- data.frame(
  Category = factor(c(
    "Cat 1:\nProtein ↑\nBurnout ↑",
    "Cat 2:\nProtein ↓\nBurnout ↓",
    "Cat 3:\nProtein ↓\nBurnout ↑",
    "Cat 4:\nProtein ↑\nBurnout ↓"
  ), levels = c(
    "Cat 1:\nProtein ↑\nBurnout ↑",
    "Cat 2:\nProtein ↓\nBurnout ↓",
    "Cat 3:\nProtein ↓\nBurnout ↑",
    "Cat 4:\nProtein ↑\nBurnout ↓"
  )),
  Count = c(nrow(cat1_proteins), nrow(cat1_proteins),
            nrow(cat3_proteins), nrow(cat3_proteins)),
  Group = c("Positive", "Positive", "Negative", "Negative"),
  Note = c("Same\nProteins", "Same\nProteins", 
           "Same\nProteins", "Same\nProteins")
)

p2 <- ggplot(bar_data, aes(x = Category, y = Count, fill = Group)) +
  geom_bar(stat = "identity", width = 0.7, color = "black", linewidth = 0.5) +
  geom_text(aes(label = Count), vjust = -0.5, size = 6, fontface = "bold") +
  geom_text(aes(label = Note), vjust = 2, size = 3, color = "white", fontface = "italic") +
  scale_fill_manual(values = c("Positive" = "#D32F2F", "Negative" = "#1976D2")) +
  theme_minimal(base_size = 13) +
  theme(
    legend.position = "bottom",
    panel.grid.major.x = element_blank(),
    axis.text.x = element_text(size = 11, face = "bold")
  ) +
  labs(
    title = "4-Way Categorization of Protein-Burnout Associations",
    subtitle = "Note: Cat 1&2 are the same proteins | Cat 3&4 are the same proteins",
    x = "",
    y = "Number of Proteins",
    fill = "Relationship Type"
  ) +
  ylim(0, max(bar_data$Count) * 1.25)

print(p2)
ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/BarChart_4Categories.pdf", p2, width = 10, height = 7, dpi = 300)

# ============================================
# VISUALIZATION 3: SPLIT VIEW WITH 4 PANELS
# ============================================

# Create 4 separate small volcano plots
create_category_plot <- function(data, title, color, xlim_range, ylim_range) {
  ggplot(data, aes(x = estimate, y = -log10(p_value))) +
    geom_point(color = color, size = 2, alpha = 0.7) +
    geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "gray50") +
    geom_vline(xintercept = 0, linetype = "solid", color = "black") +
    theme_minimal(base_size = 10) +
    xlim(xlim_range) +
    ylim(ylim_range) +
    labs(title = title, x = "Effect Size", y = "-log10(P)") +
    theme(plot.title = element_text(size = 11, face = "bold"))
}

xlim_range <- range(sig_proteins$estimate)
ylim_range <- c(0, max(-log10(sig_proteins$p_value)) * 1.1)

p3a <- create_category_plot(
  cat1_proteins,
  paste0("Category 1 (n=", nrow(cat1_proteins), ")\nProtein ↑ when Burnout ↑"),
  "#D32F2F", xlim_range, ylim_range
)

p3b <- create_category_plot(
  cat1_proteins,
  paste0("Category 2 (n=", nrow(cat1_proteins), ")\nProtein ↓ when Burnout ↓\n(Same as Cat 1)"),
  "#E57373", xlim_range, ylim_range
)

p3c <- create_category_plot(
  cat3_proteins,
  paste0("Category 3 (n=", nrow(cat3_proteins), ")\nProtein ↓ when Burnout ↑"),
  "#1976D2", xlim_range, ylim_range
)

p3d <- create_category_plot(
  cat3_proteins,
  paste0("Category 4 (n=", nrow(cat3_proteins), ")\nProtein ↑ when Burnout ↓\n(Same as Cat 3)"),
  "#64B5F6", xlim_range, ylim_range
)

p3_combined <- grid.arrange(p3a, p3b, p3c, p3d, ncol = 2)

ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/FourPanel_Categories.pdf", p3_combined, width = 12, height = 10, dpi = 300)

# ============================================
# VISUALIZATION 4: DETAILED COMPARISON TABLE
# ============================================

# Top 15 from each group
top15_cat12 <- head(cat1_proteins, 15)
top15_cat34 <- head(cat3_proteins, 15)

# Create comparison dataframe
comparison_df <- data.frame(
  Rank = 1:15,
  Cat1_2_Protein = c(top15_cat12$protein, rep("", 15 - nrow(top15_cat12)))[1:15],
  Cat1_2_Effect = c(sprintf("%.4f", top15_cat12$estimate), rep("", 15 - nrow(top15_cat12)))[1:15],
  Cat1_2_FDR = c(sprintf("%.4f", top15_cat12$adj_p_value), rep("", 15 - nrow(top15_cat12)))[1:15],
  Cat3_4_Protein = c(top15_cat34$protein, rep("", 15 - nrow(top15_cat34)))[1:15],
  Cat3_4_Effect = c(sprintf("%.4f", top15_cat34$estimate), rep("", 15 - nrow(top15_cat34)))[1:15],
  Cat3_4_FDR = c(sprintf("%.4f", top15_cat34$adj_p_value), rep("", 15 - nrow(top15_cat34)))[1:15]
)

write.csv(comparison_df, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Top15_Comparison_4Categories.csv", row.names = FALSE)

cat("\n\n=== Top 15 Comparison ===\n")
print(comparison_df)

# ============================================
# VISUALIZATION 5: INTERPRETATION DIAGRAM
# ============================================

# Create a visual explanation
explanation_data <- data.frame(
  x = c(1, 1, 3, 3),
  y = c(2, 1, 2, 1),
  label = c(
    "Category 1:\nProtein ↑\nwhen\nBurnout ↑",
    "Category 2:\nProtein ↓\nwhen\nBurnout ↓",
    "Category 3:\nProtein ↓\nwhen\nBurnout ↑",
    "Category 4:\nProtein ↑\nwhen\nBurnout ↓"
  ),
  group = c("Positive", "Positive", "Negative", "Negative"),
  count = c(nrow(cat1_proteins), nrow(cat1_proteins),
            nrow(cat3_proteins), nrow(cat3_proteins))
)

p5 <- ggplot(explanation_data, aes(x = x, y = y, fill = group)) +
  geom_tile(width = 0.8, height = 0.8, color = "black", linewidth = 1.5) +
  geom_text(aes(label = paste0(label, "\n\nn = ", count)),
            size = 5, fontface = "bold", color = "white") +
  scale_fill_manual(values = c("Positive" = "#D32F2F", "Negative" = "#1976D2")) +
  annotate("text", x = 1, y = 2.6, label = "SAME\nPROTEINS", 
           size = 4, fontface = "bold.italic", color = "#D32F2F") +
  annotate("text", x = 3, y = 2.6, label = "SAME\nPROTEINS", 
           size = 4, fontface = "bold.italic", color = "#1976D2") +
  annotate("segment", x = 1, xend = 1, y = 1.5, yend = 1.5,
           arrow = arrow(ends = "both", length = unit(0.3, "cm")),
           linewidth = 1.2, color = "#D32F2F") +
  annotate("segment", x = 3, xend = 3, y = 1.5, yend = 1.5,
           arrow = arrow(ends = "both", length = unit(0.3, "cm")),
           linewidth = 1.2, color = "#1976D2") +
  theme_void() +
  theme(legend.position = "none") +
  coord_cartesian(xlim = c(0.5, 3.5), ylim = c(0.5, 3)) +
  labs(title = "4-Category Interpretation Guide",
       subtitle = "Two statistical relationships, four interpretations")

print(p5)
ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Interpretation_Diagram.pdf", p5, width = 10, height = 6, dpi = 300)

# ============================================
# FINAL SUMMARY REPORT
# ============================================

cat("\n\n")
cat("========================================\n")
cat("FINAL REPORT: 4-CATEGORY CLASSIFICATION\n")
cat("========================================\n\n")

cat("UNDERSTANDING THE 4 CATEGORIES:\n")
cat("-------------------------------\n\n")

cat("POSITIVE ASSOCIATION (Same", nrow(cat1_proteins), "proteins):\n")
cat("  Category 1: Protein ↑ when Burnout ↑\n")
cat("  Category 2: Protein ↓ when Burnout ↓\n")
cat("  → These describe the SAME relationship from different angles\n")
cat("  → Interpretation: Protein and burnout move TOGETHER\n")
cat("  → Mean effect:", round(mean(cat1_proteins$estimate), 4), "\n\n")

cat("NEGATIVE ASSOCIATION (Same", nrow(cat3_proteins), "proteins):\n")
cat("  Category 3: Protein ↓ when Burnout ↑\n")
cat("  Category 4: Protein ↑ when Burnout ↓\n")
cat("  → These describe the SAME relationship from different angles\n")
cat("  → Interpretation: Protein and burnout move OPPOSITE\n")
cat("  → Mean effect:", round(mean(cat3_proteins$estimate), 4), "\n\n")

cat("FILES CREATED:\n")
cat("--------------\n")
cat("Data Files:\n")
cat("  ✓ Category1_2_Protein_UP_Burnout_UP.csv\n")
cat("  ✓ Category3_4_Protein_DOWN_Burnout_UP.csv\n")
cat("  ✓ Top15_Comparison_4Categories.csv\n\n")

cat("Visualizations:\n")
cat("  ✓ Volcano_4Categories.png\n")
cat("  ✓ BarChart_4Categories.png\n")
cat("  ✓ FourPanel_Categories.png\n")
cat("  ✓ Interpretation_Diagram.png\n\n")

cat("========================================\n\n")
```

###Modeling for 4 categories

```{r}

dex<- read.csv("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/MetaData/burnoutEx.csv")
rownames(dex)<-dex$X
```

Outcome = Burnout | Predictors = Protein levels, Time, Sex, Age
Best Method: Linear Mixed Models (lme4) - one model per protein

```{r}
library(lme4)
library(lmerTest)  # For p-values

meta<- dex[c(5438:5442)]
dEX<- dex[c(6:5432)]

```
#no adjusted p-value signofocant


Try another transformed data 
```{r}
#dEX<-dEX+1
#try with transformed data 
lEx<-sqrt(dEX+1) # final used and  lead to only 5 proteins with FDR< 0.2
```


```{r}
#lEx[is.na(lEx)]< 0

expr_matrix <- as.data.frame(lEx)
```


```{r}
library(lme4)
library(lmerTest)
library(dplyr)

# ============================================
# STEP 1: PREPARE DATA CORRECTLY
# ============================================

# Extract metadata and protein data
meta <- dex[, 5438:5442]  # More explicit column selection
```


```{r}
colnames(meta)
colnames(meta) <- c("code", "sex", "age", "Time", "Value")  # Name them clearly
colnames(meta)
# Protein expression data

cat("=== Data Structure ===\n")
cat("Samples (rows):", nrow(expr_matrix), "\n")
cat("Proteins (columns):", ncol(expr_matrix), "\n")
cat("Metadata rows:", nrow(meta), "\n\n")

# Verify dimensions match
stopifnot(nrow(expr_matrix) == nrow(meta))

# ============================================
# STEP 2: CLEAN AND PREPARE METADATA
# ============================================

# Convert to proper types
meta$code <- factor(meta$code)          # Subject ID
meta$Time <- factor(meta$Time)          # Categorical time (T1-T6)
meta$sex <- factor(meta$sex)            # Sex
meta$age <- as.numeric(meta$age)        # Age
meta$burnout <- as.numeric(meta$Value)  # Burnout score



#meta$Time[meta$Time== "C1"]<- "T1"
#meta$Time[meta$Time== "C2"]<- "T2"
#meta$Time[meta$Time== "C3"]<- "T3"
#meta$Time[meta$Time== "C4"]<- "T4"
#meta$Time[meta$Time== "C5"]<- "T5"
#meta$Time[meta$Time== "C6"]<- "T6"



# Create numeric time variable (1-6)
meta$time_numeric <- as.numeric(gsub("C", "", meta$Time))

# Check for issues
cat("=== Data Quality Checks ===\n")
cat("Missing values in burnout:", sum(is.na(meta$burnout)), "\n")
cat("Missing values in proteins:", sum(is.na(expr_matrix)), "\n")
cat("Unique subjects:", length(unique(meta$code)), "\n")
cat("Observations per subject:\n")
print(table(table(meta$code)))
cat("\nTime points:\n")
print(table(meta$Time))
cat("\nSex distribution:\n")
print(table(meta$sex))
cat("\nAge range:", range(meta$age, na.rm=TRUE), "\n\n")
```



```{r}
library(lme4)
library(lmerTest)
library(dplyr)

# ============================================
# METHOD 1: BIDIRECTIONAL MODELING
# ============================================

cat("=== BIDIRECTIONAL ANALYSIS ===\n")
cat("Model 1: Burnout ~ Protein (protein predicts burnout)\n")
cat("Model 2: Protein ~ Burnout (burnout predicts protein)\n\n")

# Function to test BOTH directions
test_protein_bidirectional <- function(protein_name, expr_matrix, meta) {
  
  protein_data <- data.frame(
    subject_id = meta$code,
    sex = meta$sex,
    age = meta$age,
    burnout = meta$burnout,
    protein_expression = expr_matrix[, protein_name],
    stringsAsFactors = FALSE
  )
  
  protein_data <- protein_data[complete.cases(protein_data), ]
  
  if(nrow(protein_data) < 20 || length(unique(protein_data$subject_id)) < 5) {
    return(data.frame(
      protein_to_burnout_est = NA, protein_to_burnout_p = NA,
      burnout_to_protein_est = NA, burnout_to_protein_p = NA,
      error = "Insufficient data"
    ))
  }
  
  tryCatch({
    # Direction 1: Protein → Burnout
    model1 <- lmer(burnout ~ protein_expression + sex + age + (1 | subject_id),
                   data = protein_data, REML = FALSE,
                   control = lmerControl(optimizer = "bobyqa"))
    
    coef1 <- coef(summary(model1))
    protein_to_burnout <- coef1["protein_expression", ]
    
    # Direction 2: Burnout → Protein
    model2 <- lmer(protein_expression ~ burnout + sex + age + (1 | subject_id),
                   data = protein_data, REML = FALSE,
                   control = lmerControl(optimizer = "bobyqa"))
    
    coef2 <- coef(summary(model2))
    burnout_to_protein <- coef2["burnout", ]
    
    return(data.frame(
      protein_to_burnout_est = protein_to_burnout["Estimate"],
      protein_to_burnout_p = protein_to_burnout["Pr(>|t|)"],
      burnout_to_protein_est = burnout_to_protein["Estimate"],
      burnout_to_protein_p = burnout_to_protein["Pr(>|t|)"],
      error = NA
    ))
    
  }, error = function(e) {
    return(data.frame(
      protein_to_burnout_est = NA, protein_to_burnout_p = NA,
      burnout_to_protein_est = NA, burnout_to_protein_p = NA,
      error = as.character(e$message)
    ))
  })
}

# Run bidirectional analysis
cat("Running bidirectional analysis...\n")
results_bidirectional <- data.frame(
  protein = colnames(expr_matrix),
  protein_to_burnout_est = NA,
  protein_to_burnout_p = NA,
  burnout_to_protein_est = NA,
  burnout_to_protein_p = NA,
  error = NA,
  stringsAsFactors = FALSE
)

pb <- txtProgressBar(min = 0, max = ncol(expr_matrix), style = 3)
for(i in 1:ncol(expr_matrix)) {
  result <- test_protein_bidirectional(colnames(expr_matrix)[i], expr_matrix, meta)
  results_bidirectional[i, 2:6] <- result
  setTxtProgressBar(pb, i)
}
close(pb)

# Clean results
results_bidir_clean <- results_bidirectional[
  is.na(results_bidirectional$error) & 
  !is.na(results_bidirectional$protein_to_burnout_p), ]

# Adjust p-values
results_bidir_clean$protein_to_burnout_fdr <- 
  p.adjust(results_bidir_clean$protein_to_burnout_p, method = "BH")
results_bidir_clean$burnout_to_protein_fdr <- 
  p.adjust(results_bidir_clean$burnout_to_protein_p, method = "BH")

# ============================================
# METHOD 2: CHANGE-BASED CATEGORIZATION
# ============================================

cat("\n\n=== CHANGE-BASED CATEGORIZATION ===\n")
cat("Examining actual patterns when burnout increases vs decreases\n\n")

# Function to calculate change patterns
calculate_change_patterns <- function(protein_name, expr_matrix, meta) {
  
  # Create dataframe
  data <- data.frame(
    subject_id = meta$code,
    time = meta$Time,
    burnout = meta$burnout,
    protein = expr_matrix[, protein_name]
  )
  
  # Remove NAs
  data <- data[complete.cases(data), ]
  
  # Calculate changes within each subject
  data <- data %>%
    arrange(subject_id, time) %>%
    group_by(subject_id) %>%
    mutate(
      burnout_change = burnout - lag(burnout),
      protein_change = protein - lag(protein)
    ) %>%
    filter(!is.na(burnout_change))
  
  if(nrow(data) < 10) {
    return(data.frame(
      when_burnout_increases_protein = NA,
      when_burnout_decreases_protein = NA,
      n_increases = NA,
      n_decreases = NA,
      category = "Insufficient data"
    ))
  }
  
  # Separate increases and decreases in burnout
  burnout_increases <- data[data$burnout_change > 0, ]
  burnout_decreases <- data[data$burnout_change < 0, ]
  
  # Mean protein change when burnout increases
  mean_protein_when_burnout_up <- 
    if(nrow(burnout_increases) > 0) mean(burnout_increases$protein_change, na.rm=TRUE) else NA
  
  # Mean protein change when burnout decreases
  mean_protein_when_burnout_down <- 
    if(nrow(burnout_decreases) > 0) mean(burnout_decreases$protein_change, na.rm=TRUE) else NA
  
  # Determine category
  category <- "Unclear"
  
  if(!is.na(mean_protein_when_burnout_up) && !is.na(mean_protein_when_burnout_down)) {
    if(mean_protein_when_burnout_up > 0 && mean_protein_when_burnout_down < 0) {
      category <- "Cat1: Protein↑ when Burnout↑ & Protein↓ when Burnout↓"
    } else if(mean_protein_when_burnout_up < 0 && mean_protein_when_burnout_down > 0) {
      category <- "Cat3: Protein↓ when Burnout↑ & Protein↑ when Burnout↓"
    } else if(mean_protein_when_burnout_up > 0 && mean_protein_when_burnout_down > 0) {
      category <- "Cat2: Protein↑ regardless of Burnout direction"
    } else if(mean_protein_when_burnout_up < 0 && mean_protein_when_burnout_down < 0) {
      category <- "Cat4: Protein↓ regardless of Burnout direction"
    }
  }
  
  return(data.frame(
    when_burnout_increases_protein = mean_protein_when_burnout_up,
    when_burnout_decreases_protein = mean_protein_when_burnout_down,
    n_increases = nrow(burnout_increases),
    n_decreases = nrow(burnout_decreases),
    category = category
  ))
}

# Run change analysis
cat("Analyzing change patterns for all proteins...\n")
results_changes <- data.frame(
  protein = colnames(expr_matrix),
  when_burnout_increases_protein = NA,
  when_burnout_decreases_protein = NA,
  n_increases = NA,
  n_decreases = NA,
  category = NA,
  stringsAsFactors = FALSE
)

pb <- txtProgressBar(min = 0, max = ncol(expr_matrix), style = 3)
for(i in 1:ncol(expr_matrix)) {
  result <- calculate_change_patterns(colnames(expr_matrix)[i], expr_matrix, meta)
  results_changes[i, 2:6] <- result
  setTxtProgressBar(pb, i)
}
close(pb)

# ============================================
# COMBINE RESULTS AND CATEGORIZE
# ============================================

# Merge both approaches
final_results <- merge(results_q2_clean, results_changes, by = "protein")
final_results <- merge(final_results, results_bidir_clean, by = "protein")

# Create 4 categories based on COMBINED criteria
final_results$final_category <- "Not Categorized"

# Significance threshold
sig_threshold <- 0.2

# Category 1: Protein ↑ when Burnout ↑ (AND vice versa)
final_results$final_category[
  final_results$adj_p_value < sig_threshold &
  final_results$estimate > 0 &
  final_results$burnout_to_protein_fdr < sig_threshold &
  final_results$burnout_to_protein_est > 0 &
  final_results$when_burnout_increases_protein > 0
] <- "Category 1: Protein↑ with Burnout↑"

# Category 2: Protein ↓ when Burnout ↓ (protective - goes down when burnout decreases)
final_results$final_category[
  final_results$adj_p_value < sig_threshold &
  final_results$estimate > 0 &
  final_results$burnout_to_protein_fdr < sig_threshold &
  final_results$burnout_to_protein_est > 0 &
  final_results$when_burnout_decreases_protein < 0
] <- "Category 2: Protein↓ when Burnout↓"

# Category 3: Protein ↓ when Burnout ↑ (protective)
final_results$final_category[
  final_results$adj_p_value < sig_threshold &
  final_results$estimate < 0 &
  final_results$burnout_to_protein_fdr < sig_threshold &
  final_results$burnout_to_protein_est < 0 &
  final_results$when_burnout_increases_protein < 0
] <- "Category 3: Protein↓ when Burnout↑"

# Category 4: Protein ↑ when Burnout ↓ (resilience marker)
final_results$final_category[
  final_results$adj_p_value < sig_threshold &
  final_results$estimate < 0 &
  final_results$burnout_to_protein_fdr < sig_threshold &
  final_results$burnout_to_protein_est < 0 &
  final_results$when_burnout_decreases_protein > 0
] <- "Category 4: Protein↑ when Burnout↓"

# ============================================
# EXTRACT 4 DISTINCT LISTS
# ============================================

cat("\n\n=== 4 DISTINCT CATEGORIES ===\n\n")

# Category 1
cat1_distinct <- final_results[final_results$final_category == "Category 1: Protein↑ with Burnout↑", ]
cat1_distinct <- cat1_distinct[order(cat1_distinct$adj_p_value), ]

cat("CATEGORY 1: Proteins that INCREASE when Burnout INCREASES\n")
cat("Count:", nrow(cat1_distinct), "\n")
if(nrow(cat1_distinct) > 0) {
  cat("Top 5:\n")
  print(cat1_distinct[1:min(5, nrow(cat1_distinct)), 
                      c("protein", "estimate", "adj_p_value")])
}
cat("\n")

# Category 2
cat2_distinct <- final_results[final_results$final_category == "Category 2: Protein↓ when Burnout↓", ]
cat2_distinct <- cat2_distinct[order(cat2_distinct$adj_p_value), ]

cat("CATEGORY 2: Proteins that DECREASE when Burnout DECREASES\n")
cat("Count:", nrow(cat2_distinct), "\n")
if(nrow(cat2_distinct) > 0) {
  cat("Top 5:\n")
  print(cat2_distinct[1:min(5, nrow(cat2_distinct)), 
                      c("protein", "estimate", "adj_p_value")])
}
cat("\n")

# Category 3
cat3_distinct <- final_results[final_results$final_category == "Category 3: Protein↓ when Burnout↑", ]
cat3_distinct <- cat3_distinct[order(cat3_distinct$adj_p_value), ]

cat("CATEGORY 3: Proteins that DECREASE when Burnout INCREASES\n")
cat("Count:", nrow(cat3_distinct), "\n")
if(nrow(cat3_distinct) > 0) {
  cat("Top 5:\n")
  print(cat3_distinct[1:min(5, nrow(cat3_distinct)), 
                      c("protein", "estimate", "adj_p_value")])
}
cat("\n")

# Category 4
cat4_distinct <- final_results[final_results$final_category == "Category 4: Protein↑ when Burnout↓", ]
cat4_distinct <- cat4_distinct[order(cat4_distinct$adj_p_value), ]

cat("CATEGORY 4: Proteins that INCREASE when Burnout DECREASES\n")
cat("Count:", nrow(cat4_distinct), "\n")
if(nrow(cat4_distinct) > 0) {
  cat("Top 5:\n")
  print(cat4_distinct[1:min(5, nrow(cat4_distinct)), 
                      c("protein", "estimate", "adj_p_value")])
}
cat("\n")

# Save separate files
write.csv(cat1_distinct, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Details_posneg/FourCategory/Category1_Protein_UP_Burnout_UP_DISTINCT.csv", row.names = FALSE)
write.csv(cat2_distinct, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Details_posneg/FourCategory/Category2_Protein_DOWN_Burnout_DOWN_DISTINCT.csv", row.names = FALSE)
write.csv(cat3_distinct, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Details_posneg/FourCategory/Category3_Protein_DOWN_Burnout_UP_DISTINCT.csv", row.names = FALSE)
write.csv(cat4_distinct, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Details_posneg/FourCategory/Category4_Protein_UP_Burnout_DOWN_DISTINCT.csv", row.names = FALSE)

# Summary table
summary_4cat_distinct <- data.frame(
  Category = c(
    "Category 1: Protein ↑ when Burnout ↑",
    "Category 2: Protein ↓ when Burnout ↓",
    "Category 3: Protein ↓ when Burnout ↑",
    "Category 4: Protein ↑ when Burnout ↓"
  ),
  Count = c(
    nrow(cat1_distinct),
    nrow(cat2_distinct),
    nrow(cat3_distinct),
    nrow(cat4_distinct)
  )
)

cat("\n=== SUMMARY TABLE ===\n")
print(summary_4cat_distinct)

# ============================================
# VISUALIZATION: 4 DISTINCT CATEGORIES
# ============================================

library(ggplot2)

# Bar plot
p_bar <- ggplot(summary_4cat_distinct, 
                aes(x = reorder(Category, -Count), y = Count, fill = Category)) +
  geom_bar(stat = "identity", color = "black", linewidth = 0.5) +
  geom_text(aes(label = Count), vjust = -0.5, size = 6, fontface = "bold") +
  scale_fill_manual(values = c(
    "Category 1: Protein ↑ when Burnout ↑" = "#D32F2F",
    "Category 2: Protein ↓ when Burnout ↓" = "#FF6F00",
    "Category 3: Protein ↓ when Burnout ↑" = "#1976D2",
    "Category 4: Protein ↑ when Burnout ↓" = "#388E3C"
  )) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10)
  ) +
  labs(
    title = "4 Distinct Categories Based on Change Patterns",
    x = "",
    y = "Number of Proteins"
  )

print(p_bar)
ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Details_posneg/FourCategory/BarPlot_4_DISTINCT_Categories.pdf", p_bar, width = 12, height = 8, dpi = 300)

cat("\n=== ANALYSIS COMPLETE ===\n")
cat("Four distinct protein lists created based on actual change patterns!\n")
```

##only C1 and C6


```{r}
library(lme4)
library(lmerTest)  # For p-values
colnames(dex)
dexC1_6<- dex %>% filter(Time.x %in% c("C1", "C6"))


metaC1_C6<- dexC1_6[c(5438:5442)]
expr_matrixC1_C6 <- as.data.frame(dexC1_6[, 6:5437]) # no FDR< 0.2
```
#no adjusted p-value signofocant


Try another transformed data 
```{r}
#dEX<-dEX+1
#try with transformed data 
lEx<-sqrt(expr_matrixC1_C6+1) # final used and  lead to only 5 proteins with FDR< 0.2
expr_matrix <- as.data.frame(lEx)
```

```{r}



cat("Combined data dimensions:", dim(full_data), "\n\n")

# ============================================
# STEP 4: TEST ONE PROTEIN FIRST (Diagnostic)
# ============================================

cat("=== Testing first protein as example ===\n")

# Get first protein name
first_protein <- colnames(expr_matrix)[1]
cat("Testing protein:", first_protein, "\n")

# Create test dataset
test_data <- data.frame(
  subject_id = metaC1_C6$code,
 # time_numeric = meta$time_numeric,
  sex = metaC1_C6$sex,
  age = metaC1_C6$age,
  burnout = metaC1_C6$Value,
  protein_expression = expr_matrix[, first_protein]
)

# Check for complete cases
cat("Complete cases:", sum(complete.cases(test_data)), "/", nrow(test_data), "\n")


library(lme4)
library(lmerTest)
library(dplyr)

# Fit test model
test_model <- lmer(burnout ~ protein_expression + sex + age + 
                     (1 | subject_id), #time_numeric +
                   data = test_data,
                   REML = FALSE)

cat("\nTest model summary:\n")
print(summary(test_model))

# ============================================
# STEP 5: IMPROVED FUNCTION TO TEST PROTEINS
# ============================================

test_protein_burnout <- function(protein_name, expr_matrix, meta) {
  
  # Create data for this protein
  protein_data <- data.frame(
    subject_id = metaC1_C6$code,
    #time_numeric = meta$time_numeric,
    sex = metaC1_C6$sex,
    age = metaC1_C6$age,
    burnout = metaC1_C6$Value,
    protein_expression = expr_matrix[, protein_name],  # FIXED: column subset
    stringsAsFactors = FALSE
  )
  
  # Remove rows with missing values
  #protein_data <- protein_data[complete.cases(protein_data), ]
  
  # Check if enough data
  if(nrow(protein_data) < 20 || 
     length(unique(protein_data$subject_id)) < 5) {
    return(data.frame(
      estimate = NA, se = NA, t_value = NA, p_value = NA,
      n_obs = nrow(protein_data), n_subjects = length(unique(protein_data$subject_id)),
      error = "Insufficient data"
    ))
  }
  
  # Fit model
  tryCatch({
    model <- lmer(burnout ~ protein_expression +  sex + age + 
                    (1 | subject_id), #time_numeric +
                  data = protein_data,
                  REML = FALSE,
                  control = lmerControl(optimizer = "bobyqa"))
    
    # Extract coefficients
    coefs <- coef(summary(model))
    protein_row <- coefs["protein_expression", ]
    
    result <- data.frame(
      estimate = protein_row["Estimate"],
      se = protein_row["Std. Error"],
      t_value = protein_row["t value"],
      p_value = protein_row["Pr(>|t|)"],
      n_obs = nrow(protein_data),
      n_subjects = length(unique(protein_data$subject_id)),
      error = NA
    )
    
    return(result)
    
  }, error = function(e) {
    return(data.frame(
      estimate = NA, se = NA, t_value = NA, p_value = NA,
      n_obs = nrow(protein_data), 
      n_subjects = length(unique(protein_data$subject_id)),
      error = as.character(e$message)
    ))
  })
}

# ============================================
# STEP 6: RUN FOR ALL PROTEINS
# ============================================

cat("\n=== Running analysis for all proteins ===\n")
cat("Testing", ncol(expr_matrix), "proteins...\n\n")

# Initialize results dataframe
results_q2C1C6 <- data.frame(
  protein = colnames(expr_matrix),  # FIXED: colnames not rownames
  estimate = NA,
  se = NA,
  t_value = NA,
  p_value = NA,
  n_obs = NA,
  n_subjects = NA,
  error = NA,
  stringsAsFactors = FALSE
)

# Progress bar
pb <- txtProgressBar(min = 0, max = ncol(expr_matrix), style = 3)

# Loop through proteins
for(i in 1:ncol(expr_matrix)) {
  
  protein_name <- colnames(expr_matrix)[i]
  
  # Test this protein
  resultC1C6 <- test_protein_burnout(protein_name, expr_matrix, meta)
  
  # Store results
  results_q2C1C6[i, 2:8] <- resultC1C6
  
  setTxtProgressBar(pb, i)
}

close(pb)

# ============================================
# STEP 7: PROCESS RESULTS
# ============================================

cat("\n\n=== Processing Results ===\n")

# Check for errors
n_errors <- sum(!is.na(results_q2C1C6$error))
cat("Models with errors:", n_errors, "\n")
if(n_errors > 0) {
  cat("Error types:\n")
  print(table(results_q2C1C6$error, useNA = "ifany"))
}

# Remove failed models
results_q2_cleanC1C6 <- results_q2C1C6[is.na(results_q2C1C6$error) & !is.na(results_q2C1C6$p_value), ]
cat("Successful models:", nrow(results_q2_clean), "\n\n")

# Multiple testing correction
results_q2_cleanC1C6$adj_p_value <- p.adjust(results_q2_cleanC1C6$p_value, method = "BH")

# Sort by significance
results_q2_cleanC1C6 <-results_q2_cleanC1C6[order(results_q2_cleanC1C6$adj_p_value), ]

# ============================================
# STEP 8: SUMMARY STATISTICS
# ============================================

cat("=== Results Summary ===\n")
cat("Total proteins tested:", nrow(results_q2_clean), "\n")
cat("Significant (FDR <0.2):", 
    sum(results_q2_clean$adj_p_value < 0.2, na.rm=TRUE), "\n")
cat("Significant (FDR < 0.20):", 
    sum(results_q2_clean$adj_p_value < 0.10, na.rm=TRUE), "\n")
cat("Nominal p < 0.05:", 
    sum(results_q2_clean$p_value < 0.05, na.rm=TRUE), "\n\n")

# Direction of effects
sig_proteins <- results_q2_clean[results_q2_clean$adj_p_value < 0.2, ]
if(nrow(sig_proteins) > 0) {
  cat("Direction of significant effects:\n")
  cat("  Positive (higher protein → higher burnout):", 
      sum(sig_proteins$estimate > 0), "\n")
  cat("  Negative (higher protein → lower burnout):", 
      sum(sig_proteins$estimate < 0), "\n\n")
}

write.csv(results_q2_cleanC1C6, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/C1C6/results_q2_cleanC1C6.csv")
```

```{r}
results_q2_clean<-read.csv("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/C1C6/results_q2_cleanC1C6.csv")
```

```{r}
# ============================================
# STEP 1: CATEGORIZE SIGNIFICANT PROTEINS
# ============================================

# Filter significant proteins (FDR < 0.2)
sig_proteins <- results_q2_clean[results_q2_clean$adj_p_value < 0.2, ]

cat("=== Categorizing Significant Proteins ===\n")
cat("Total significant proteins (FDR < 0.2):", nrow(sig_proteins), "\n\n")

# Create categories based on direction
sig_proteins$category <- NA

# POSITIVE ASSOCIATION (estimate > 0)
# Higher protein → Higher burnout (and vice versa)
sig_proteins$category[sig_proteins$estimate > 0] <- "Positive Association\n(↑ Protein → ↑ Burnout)\n(↓ Protein → ↓ Burnout)"

# NEGATIVE ASSOCIATION (estimate < 0)  
# Higher protein → Lower burnout (inverse relationship)
sig_proteins$category[sig_proteins$estimate < 0] <- "Negative Association\n(↑ Protein → ↓ Burnout)\n(↓ Protein → ↑ Burnout)"

# Create 4 subcategories for detailed interpretation
sig_proteins$detailed_category <- NA

# Category 1: Positive estimate - protein increases WITH burnout
sig_proteins$detailed_category[sig_proteins$estimate > 0] <- 
  "1. Increases with Burnout\n(Burnout ↑ → Protein ↑)"

# Category 2: Negative estimate - protein decreases WITH burnout  
sig_proteins$detailed_category[sig_proteins$estimate < 0] <- 
  "2. Decreases with Burnout\n(Burnout ↑ → Protein ↓)"

# Summary table
summary_table <- data.frame(
  Category = c(
    "Positive Association (↑ Protein → ↑ Burnout)\n(↓ Protein → ↓ Burnout)",
    "Negative Association (↑ Protein → ↓ Burnout)\n(↓ Protein → ↑ Burnout)"
  ),
  Count = c(
    sum(sig_proteins$estimate > 0),
    sum(sig_proteins$estimate < 0)
  ),
  Mean_Effect = c(
    mean(sig_proteins$estimate[sig_proteins$estimate > 0]),
    mean(sig_proteins$estimate[sig_proteins$estimate < 0])
  )
)

cat("Summary of Associations:\n")
print(summary_table)
cat("\n")

# ============================================
# STEP 2: SEPARATE INTO LISTS
# ============================================

# Positive association: Proteins that INCREASE with burnout
proteins_increase_with_burnout <- sig_proteins[sig_proteins$estimate > 0, ]
proteins_increase_with_burnout <- proteins_increase_with_burnout[
  order(proteins_increase_with_burnout$adj_p_value), ]

# Negative association: Proteins that DECREASE with burnout
proteins_decrease_with_burnout <- sig_proteins[sig_proteins$estimate < 0, ]
proteins_decrease_with_burnout <- proteins_decrease_with_burnout[
  order(proteins_decrease_with_burnout$adj_p_value), ]

cat("=== POSITIVE ASSOCIATION ===\n")
cat("Proteins that INCREASE when burnout INCREASES:\n")
cat("Count:", nrow(proteins_increase_with_burnout), "\n")
cat("\nTop 10:\n")
print(proteins_increase_with_burnout[1:min(10, nrow(proteins_increase_with_burnout)), 
                                     c("protein", "estimate", "p_value", "adj_p_value")])

cat("\n\n=== NEGATIVE ASSOCIATION ===\n")
cat("Proteins that DECREASE when burnout INCREASES:\n")
cat("Count:", nrow(proteins_decrease_with_burnout), "\n")
cat("\nTop 10:\n")
print(proteins_decrease_with_burnout[1:min(10, nrow(proteins_decrease_with_burnout)), 
                                     c("protein", "estimate", "p_value", "adj_p_value")])

# Save separate files
write.csv(proteins_increase_with_burnout, 
          "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/C1C6/Proteins_INCREASE_with_Burnout.csv", 
          row.names = FALSE)
write.csv(proteins_decrease_with_burnout, 
          "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/C1C6/Proteins_DECREASE_with_Burnout.csv", 
          row.names = FALSE)
```

```{r}
colnames(results_q2_clean)[10]<-"FDR"
colnames(results_q2_clean)[3]<-"Estimate"

results_table<-results_q2_clean
results_table$P_value<-results_table$p_value
results_table$Protein_Name<-results_table$protein
```


```{r}
cat("=== Creating Enhanced Volcano Plots with Quadrant Counts ===\n\n")

create_lmer_volcano <- function(results_table, title, subtitle) {
  
  # Prepare data
  volcano_data <- results_table %>%
    mutate(
      neg_log10_p = -log10(P_value),
      
      # Color categories
      Color_category = case_when(
        FDR < 0.2 & Estimate > 0 ~ "Increased (FDR < 0.2)",
        FDR < 0.2 & Estimate < 0 ~ "Decreased (FDR < 0.2)",
        TRUE ~ "Not Significant"
      ),
      
      # Label top proteins
      Label = case_when(
        FDR < 0.05 & abs(Estimate) > quantile(abs(Estimate), 0.95, na.rm = TRUE) ~ Protein_Name,
        TRUE ~ ""
      )
    )
  
  # Count significant proteins
  n_increased <- sum(volcano_data$Color_category == "Increased (FDR < 0.2)")
  n_decreased <- sum(volcano_data$Color_category == "Decreased (FDR < 0.2)")
  n_sig_total <- n_increased + n_decreased
  
  # Create plot
  p <- ggplot(volcano_data, aes(x = Estimate, y = neg_log10_p, color = Color_category)) +
    
    # Points
    geom_point(alpha = 0.6, size = 2.5) +
    
    # Significance threshold line (p = 0.05)
    geom_hline(yintercept = -log10(0.05), 
               linetype = "dashed", 
               color = "black", 
               linewidth = 1) +
    
    # Vertical line at estimate = 0
    geom_vline(xintercept = 0, 
               linetype = "solid", 
               color = "gray30", 
               linewidth = 0.8) +
    
    # Color scheme
    scale_color_manual(
      values = c(
        "Increased (FDR < 0.2)" = "#E53935",      # Red
        "Decreased (FDR < 0.2)" = "#1E88E5",      # Blue
        "Not Significant" = "#BDBDBD"             # Gray
      )
    ) +
    
    # Labels for top proteins
    geom_text_repel(
      aes(label = Label),
      size = 3,
      max.overlaps = 15,
      box.padding = 0.5,
      segment.color = "gray50",
      show.legend = FALSE
    ) +
    
    # Theme
    theme_minimal(base_size = 14) +
    theme(
      legend.position = "bottom",
      legend.title = element_blank(),
      panel.grid.minor = element_blank(),
      plot.title = element_text(face = "bold", size = 16),
      plot.subtitle = element_text(size = 12)
    ) +
    
    # Labels
    labs(
      title = title,
      subtitle = sprintf("%s | Significant: %d total (%d red, %d blue)",
                        subtitle, n_sig_total, n_increased, n_decreased),
      x = "Effect Size (Estimate)",
      y = "-log10(P-value)"
    ) +
    
    # Annotations
    annotate("text", 
             x = max(volcano_data$Estimate, na.rm = TRUE) * 0.7,
             y = -log10(0.05) + 0.3,
             label = "p = 0.05",
             size = 4, 
             color = "black") +
    
    annotate("text", 
             x = max(volcano_data$Estimate, na.rm = TRUE) * 0.7,
             y = max(volcano_data$neg_log10_p, na.rm = TRUE) * 0.9,
             label = paste0("Red: n = ", n_increased),
             size = 5, fontface = "bold", color = "#E53935") +
    
    annotate("text", 
             x = min(volcano_data$Estimate, na.rm = TRUE) * 0.7,
             y = max(volcano_data$neg_log10_p, na.rm = TRUE) * 0.9,
             label = paste0("Blue: n = ", n_decreased),
             size = 5, fontface = "bold", color = "#1E88E5")
  
  return(p)
}
```


```{r}
# ============================================
# STEP 3: CREATE VOLCANO PLOTS
# ============================================

cat("=== Creating Volcano Plots ===\n\n")

# Plot A: Protein → Burnout
p_protein_to_burnout <- create_lmer_volcano(
 results_table,
  title = "Volcano Plot: Proteins Predicting Burnout",
  subtitle = "Model: Burnout ~ Protein + (1|ID) | FDR < 0.2"
)

print(p_protein_to_burnout)
```

```{r}
library(ggplot2)
library(ggrepel)
library(dplyr)

# ============================================
# ENHANCED VOLCANO PLOT FUNCTION
# ============================================

create_improved_volcano <- function(results_table, title_text = "Proteins Predicting Burnout") {
  
  # Prepare data
  volcano_data <- results_table %>%
    mutate(
      neg_log10_p = -log10(P_value),
      
      # Remove extreme outliers for better visualization
      Estimate_capped = case_when(
        Estimate > quantile(Estimate, 0.99, na.rm = TRUE) ~ quantile(Estimate, 0.99, na.rm = TRUE),
        Estimate < quantile(Estimate, 0.01, na.rm = TRUE) ~ quantile(Estimate, 0.01, na.rm = TRUE),
        TRUE ~ Estimate
      ),
      
      # Color categories with better labels
      Significance = case_when(
        FDR < 0.2 & Estimate > 0 ~ "Up (FDR < 0.2)",
        FDR < 0.2 & Estimate < 0 ~ "Down (FDR < 0.2)",
        P_value < 0.05 ~ "Nominal (p < 0.05)",
        TRUE ~ "Not Significant"
      ),
      
      # Priority labeling: top by FDR and effect size
      Label_priority = case_when(
        FDR < 0.01 ~ 1,  # Most significant
        FDR < 0.05 ~ 2,
        FDR < 0.2 & abs(Estimate) > median(abs(Estimate), na.rm = TRUE) ~ 3,
        TRUE ~ 4
      ),
      
      # Create labels for top proteins
      Protein_Label = case_when(
        Label_priority <= 2 & row_number() <= 20 ~ Protein_Name,
        TRUE ~ ""
      )
    ) %>%
    arrange(Label_priority, desc(abs(Estimate)))
  
  # Calculate statistics
  n_up <- sum(volcano_data$Significance == "Up (FDR < 0.2)")
  n_down <- sum(volcano_data$Significance == "Down (FDR < 0.2)")
  n_total_sig <- n_up + n_down
  
  # FDR thresholds for reference lines
  fdr_005_threshold <- max(volcano_data$neg_log10_p[volcano_data$FDR <= 0.05], na.rm = TRUE)
  fdr_01_threshold <- max(volcano_data$neg_log10_p[volcano_data$FDR <= 0.1], na.rm = TRUE)
  
  # Create plot
  p <- ggplot(volcano_data, 
              aes(x = Estimate_capped, y = neg_log10_p, color = Significance)) +
    
    # Add shaded quadrants for context
    annotate("rect", 
             xmin = 0, xmax = Inf, ymin = -log10(0.05), ymax = Inf,
             fill = "#FFE5E5", alpha = 0.3) +
    annotate("rect", 
             xmin = -Inf, xmax = 0, ymin = -log10(0.05), ymax = Inf,
             fill = "#E5F2FF", alpha = 0.3) +
    
    # Points with size variation
    geom_point(aes(size = neg_log10_p), alpha = 0.7) +
    
    # Reference lines
    geom_hline(yintercept = -log10(0.05), 
               linetype = "dashed", color = "black", linewidth = 0.8) +
    geom_vline(xintercept = 0, 
               linetype = "solid", color = "gray20", linewidth = 0.5) +
    
    # Optional: FDR thresholds
    geom_hline(yintercept = fdr_005_threshold, 
               linetype = "dotted", color = "darkred", linewidth = 0.6, alpha = 0.7) +
    
    # Color scheme
    scale_color_manual(
      values = c(
        "Up (FDR < 0.2)" = "#D32F2F",          # Bold red
        "Down (FDR < 0.2)" = "#1976D2",        # Bold blue
        "Nominal (p < 0.05)" = "#FFA726",      # Orange
        "Not Significant" = "#9E9E9E"          # Gray
      ),
      breaks = c("Up (FDR < 0.2)", "Down (FDR < 0.2)", 
                 "Nominal (p < 0.05)", "Not Significant")
    ) +
    
    # Size scale
    scale_size_continuous(range = c(1, 4), guide = "none") +
    
    # Labels for top proteins
    geom_text_repel(
      aes(label = Protein_Label),
      size = 3.5,
      fontface = "bold",
      max.overlaps = 20,
      box.padding = 0.5,
      point.padding = 0.3,
      segment.color = "gray40",
      segment.size = 0.3,
      min.segment.length = 0,
      force = 2,
      show.legend = FALSE
    ) +
    
    # Annotations
    annotate("text", 
             x = max(volcano_data$Estimate_capped, na.rm = TRUE) * 0.75,
             y = -log10(0.05) - 5,
             label = "p = 0.05",
             size = 3.5, 
             fontface = "italic",
             color = "black") +
    
    annotate("text", 
             x = max(volcano_data$Estimate_capped, na.rm = TRUE) * 0.7,
             y = max(volcano_data$neg_log10_p, na.rm = TRUE) * 0.95,
             label = sprintf("Upregulated\nn = %d", n_up),
             size = 5, 
             fontface = "bold", 
             color = "#D32F2F") +
    
    annotate("text", 
             x = min(volcano_data$Estimate_capped, na.rm = TRUE) * 0.7,
             y = max(volcano_data$neg_log10_p, na.rm = TRUE) * 0.95,
             label = sprintf("Downregulated\nn = %d", n_down),
             size = 5, 
             fontface = "bold", 
             color = "#1976D2") +
    
    # Theme
    theme_classic(base_size = 14) +
    theme(
      # Legend
      legend.position = "bottom",
      legend.title = element_blank(),
      legend.text = element_text(size = 12),
      legend.background = element_rect(fill = "white", color = "gray80"),
      legend.key.size = unit(0.8, "cm"),
      
      # Plot area
      panel.grid.major = element_line(color = "gray90", linewidth = 0.3),
      panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
      
      # Text
      plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
      plot.subtitle = element_text(size = 12, hjust = 0.5, color = "gray30"),
      axis.title = element_text(face = "bold", size = 13),
      axis.text = element_text(size = 11)
    ) +
    
    # Labels
    labs(
      title = title_text,
      subtitle = sprintf(
        "Model: Burnout ~ Protein + (1|ID) | %d significant proteins (FDR < 0.2)",
        n_total_sig
      ),
      x = "Effect Size (β coefficient)",
      y = expression(-log[10](italic(P)*'-value')),
      color = NULL
    ) +
    
    # Scales
    scale_y_continuous(expand = expansion(mult = c(0.02, 0.1))) +
    scale_x_continuous(expand = expansion(mult = c(0.1, 0.1)))
  
  return(p)
}

# ============================================
# CREATE IMPROVED VOLCANO PLOT
# ============================================
```
# ============================================
# CREATE IMPROVED VOLCANO PLOT
# ============================================

```{r}
cat("=== Creating Improved Volcano Plot ===\n\n")

# For Model A: Protein → Burnout
p_improved_A <- create_improved_volcano(
  results_table,
  title_text = "Volcano Plot: Proteins Predicting Burnout"
)

print(p_improved_A)
```


```{r}
ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/C1C6/VolcanoC1C6.png", 
       p_improved_A, 
       width = 8, 
       height = 10, 
       dpi = 300,
       bg = "white")



ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/C1C6/VolcanoC1C6.pdf", 
       p_improved_A, 
       width = 8, 
       height = 10, 
       dpi = 300,
       bg = "white")

cat("✓ Volcano_IMPROVED_Burnout_to_Protein.png\n\n")
```


```{r}
# ============================================
# ALTERNATIVE: INTERACTIVE VOLCANO PLOT
# ============================================

library(plotly)

create_interactive_volcano <- function(results_table, title_text) {
  
  plot_data <- results_table %>%
    mutate(
      neg_log10_p = -log10(P_value),
      Color_category = case_when(
        FDR < 0.2 & Estimate > 0 ~ "Up",
        FDR < 0.2 & Estimate < 0 ~ "Down",
        TRUE ~ "NS"
      ),
      Hover_text = paste0(
        "<b>", Protein_Name, "</b><br>",
        "Estimate: ", round(Estimate, 4), "<br>",
        "P-value: ", format(P_value, scientific = TRUE, digits = 3), "<br>",
        "FDR: ", format(FDR, scientific = TRUE, digits = 3), "<br>",
        "t-value: ", round(t_value, 2)
      )
    )
  
  p <- plot_ly(
    data = plot_data,
    x = ~Estimate,
    y = ~neg_log10_p,
    color = ~Color_category,
    colors = c("Up" = "#D32F2F", "Down" = "#1976D2", "NS" = "#9E9E9E"),
    type = "scatter",
    mode = "markers",
    marker = list(size = 8, opacity = 0.7),
    text = ~Hover_text,
    hoverinfo = "text"
  ) %>%
    layout(
      title = list(text = title_text, font = list(size = 16, face = "bold")),
      xaxis = list(title = "Effect Size (Estimate)", zeroline = TRUE),
      yaxis = list(title = "-log10(P-value)"),
      hovermode = "closest",
      shapes = list(
        # Horizontal line at p=0.05
        list(
          type = "line",
          x0 = min(plot_data$Estimate, na.rm = TRUE),
          x1 = max(plot_data$Estimate, na.rm = TRUE),
          y0 = -log10(0.05),
          y1 = -log10(0.05),
          line = list(dash = "dash", color = "black", width = 2)
        )
      )
    )
  
  return(p)
}

# Create interactive version
p_interactive <- create_interactive_volcano(
  results_protein_to_burnout,
  "Interactive Volcano Plot: Proteins Predicting Burnout"
)

htmlwidgets::saveWidget(p_interactive, 
                        "Volcano_Interactive.html",
                        selfcontained = TRUE)

cat("✓ Volcano_Interactive.html (hover over points for details)\n\n")

# ============================================
# PUBLICATION-QUALITY VERSION
# ============================================

create_publication_volcano <- function(results_table) {
  
  volcano_data <- results_table %>%
    mutate(
      neg_log10_p = -log10(P_value),
      Significance = case_when(
        FDR < 0.05 ~ "FDR < 0.05",
        FDR < 0.2 ~ "FDR < 0.2",
        P_value < 0.05 ~ "p < 0.05",
        TRUE ~ "NS"
      ),
      Direction = ifelse(Estimate > 0, "Up", "Down"),
      Combined = paste0(Direction, "_", Significance),
      
      # Label only top 10 by FDR
      Label = ifelse(row_number() <= 10 & FDR < 0.2, Protein_Name, "")
    ) %>%
    arrange(FDR)
  
  p <- ggplot(volcano_data, aes(x = Estimate, y = neg_log10_p)) +
    
    # Base points
    geom_point(
      data = filter(volcano_data, Significance == "NS"),
      color = "#CCCCCC",
      alpha = 0.3,
      size = 2
    ) +
    
    # Significant points
    geom_point(
      data = filter(volcano_data, FDR < 0.2),
      aes(color = Direction, shape = Significance),
      size = 3,
      alpha = 0.8
    ) +
    
    # Lines
    geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "black") +
    geom_vline(xintercept = 0, linetype = "solid", color = "gray30") +
    
    # Labels
    geom_text_repel(
      aes(label = Label),
      size = 3,
      fontface = "bold.italic",
      max.overlaps = 15,
      box.padding = 0.5,
      segment.size = 0.2
    ) +
    
    # Colors and shapes
    scale_color_manual(
      values = c("Up" = "#D32F2F", "Down" = "#1976D2"),
      name = "Direction"
    ) +
    scale_shape_manual(
      values = c("FDR < 0.05" = 16, "FDR < 0.2" = 17),
      name = "Significance"
    ) +
    
    # Theme
    theme_bw(base_size = 12) +
    theme(
      panel.grid.minor = element_blank(),
      legend.position = "right",
      plot.title = element_text(hjust = 0.5, face = "bold"),
      axis.title = element_text(face = "bold")
    ) +
    
    labs(
      title = "Volcano Plot: Protein-Burnout Association",
      x = "Effect Size (β)",
      y = expression(-log[10](italic(P)))
    )
  
  return(p)
}

p_publication <- create_publication_volcano(results_protein_to_burnout)

ggsave("Volcano_PUBLICATION_Quality.png", 
       p_publication, 
       width = 10, 
       height = 8, 
       dpi = 600,
       bg = "white")

cat("✓ Volcano_PUBLICATION_Quality.png (600 dpi)\n\n")

# ============================================
# SUMMARY OF IMPROVEMENTS
# ============================================

cat("\n========================================\n")
cat("VOLCANO PLOT IMPROVEMENTS\n")
cat("========================================\n\n")

cat("VISUAL IMPROVEMENTS:\n")
cat("  ✓ Better axis scaling (removed extreme outliers)\n")
cat("  ✓ Shaded quadrants for easier interpretation\n")
cat("  ✓ Point size varies by significance\n")
cat("  ✓ Better color contrast (bold red/blue)\n")
cat("  ✓ Improved label placement (ggrepel)\n")
cat("  ✓ Quadrant annotations with counts\n\n")

cat("STATISTICAL CLARITY:\n")
cat("  ✓ Multiple significance thresholds shown\n")
cat("  ✓ FDR threshold line added\n")
cat("  ✓ Nominal significance category (p<0.05)\n")
cat("  ✓ Clear legend with all categories\n\n")

cat("PROFESSIONAL TOUCHES:\n")
cat("  ✓ Publication-quality formatting\n")
cat("  ✓ Mathematical notation for axis labels\n")
cat("  ✓ Clean, crisp borders\n")
cat("  ✓ High-resolution output (300-600 dpi)\n")
cat("  ✓ White background for publications\n\n")

cat("FILES CREATED:\n")
cat("  ✓ Volcano_IMPROVED_Protein_to_Burnout.png\n")
cat("  ✓ Volcano_IMPROVED_Burnout_to_Protein.png\n")
cat("  ✓ Volcano_Interactive.html\n")
cat("  ✓ Volcano_PUBLICATION_Quality.png (600 dpi)\n\n")

cat("========================================\n\n")
```



```{r}
ggsave("Volcano_LMER_Protein_to_Burnout.png", p_protein_to_burnout, 
       width = 12, height = 10, dpi = 300)

cat("✓ Volcano_LMER_Protein_to_Burnout.png\n")

# Plot B: Burnout → Protein
p_burnout_to_protein <- create_lmer_volcano(
  results_burnout_to_protein,
  title = "Volcano Plot: Burnout Predicting Proteins",
  subtitle = "Model: Protein ~ Burnout + (1|ID) | FDR < 0.2"
)

print(p_burnout_to_protein)

```







```{r}
# ============================================
# VISUALIZATION 1: ENHANCED VOLCANO PLOT
# ============================================

# Add category for all proteins (including non-significant)
results_q2_clean$direction_category <- "Not Significant"

results_q2_clean$direction_category[
  results_q2_clean$adj_p_value < 0.2 & results_q2_clean$estimate > 0
] <- "(↑ Protein → ↑ Burnout)\n(↓ Protein → ↓ Burnout)"

results_q2_clean$direction_category[
  results_q2_clean$adj_p_value < 0.2 & results_q2_clean$estimate < 0
] <- "(↑ Protein → ↓ Burnout)\n(↓ Protein → ↑ Burnout)"

# Label top proteins from each category
results_q2_clean$protein_label <- ""

# Top 10 positive
top_positive <- head(proteins_increase_with_burnout, 10)
results_q2_clean$protein_label[
  results_q2_clean$protein %in% top_positive$protein
] <- results_q2_clean$protein[
  results_q2_clean$protein %in% top_positive$protein
]

# Top 10 negative
top_negative <- head(proteins_decrease_with_burnout, 10)
results_q2_clean$protein_label[
  results_q2_clean$protein %in% top_negative$protein
] <- results_q2_clean$protein[
  results_q2_clean$protein %in% top_negative$protein
]

# Create volcano plot
#####Significant
results_q2_clean$status<-results_q2_clean$estimate
results_q2_clean<- ifelse(results_q2_clean$status< 0 ,-1 ,1)


results_q2_clean$status <- ifelse(results_q2_clean$estimate <0, "Negative:(↑ Protein → ↓ Burnout)\n(↓ Protein → ↑ Burnout)",ifelse(results_q2_clean$estimate >0, "Positive:(↑ Protein → ↑ Burnout)\n(↓ Protein → ↓ Burnout)",)

p.vol<-ggplot(res, aes(x = logFC, y = -log10(p_value))) +
  geom_point(aes(color = status),shape=16, size=2,alpha = 1) +
  theme(legend.position = "bottom") + theme_bw(base_size = 10)+ #xlim(c(-0.1, 0.1)) + ylim(c(0, 21))+
  scale_color_manual(values=c("darkcyan","red4","blue"))+
  theme_classic(base_size = 8)+ facet_grid(comparison_group~tissue_abbreviation ) +
  
  geom_hline(yintercept = 1.3, colour="red", linetype="dashed",size=0.1) +
  geom_vline(xintercept = 0, colour="gray", linetype="longdash", size=0.1)

```


```{r}
pv <- ggplot(results_q2_clean, 
             aes(x = log10(estimate), 
                 y = -log10(p_value), 
                 color = direction_category,
                 label = protein_label)) +
  geom_point(alpha = 0.6, size = 2) +
  geom_hline(yintercept = -log10(0.05), 
             linetype = "dashed", color = "gray30", linewidth = 0.8) +
  geom_vline(xintercept = 0, 
             linetype = "solid", color = "black", linewidth = 0.8) +
  geom_text_repel(
    size = 3,
    max.overlaps = 25,
    box.padding = 0.5,
    segment.color = "gray50",
    fontface = "bold"
  ) +
  scale_color_manual(
    values = c(
      "↑ Protein with ↑ Burnout" = "#D32F2F",      # Red
      "↓ Protein with ↑ Burnout" = "#1976D2",      # Blue
      "Not Significant" = "#BDBDBD"                 # Gray
    )
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  ) +
  labs(
    title = "Protein-Burnout Associations",
    subtitle = sprintf("Positive: n=%d | Negative: n=%d", 
                      nrow(proteins_increase_with_burnout),
                      nrow(proteins_decrease_with_burnout)),
    x = "Effect Size: Protein Effect on Burnout\n← Lower Burnout | Higher Burnout →",
    y = "-log10(P-value)",
    color = "Association Type"
  ) +#xlim(-2,2)+
  annotate("rect", 
           xmin = 0, xmax = Inf, ymin = -log10(0.05), ymax = Inf,
           alpha = 0.1, fill = "red") +
  annotate("rect", 
           xmin = -Inf, xmax = 0, ymin = -log10(0.05), ymax = Inf,
           alpha = 0.1, fill = "blue")

print(pv)
```


```{r}
ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/C1C6/Volcano_Directional_Categories.png", pv, width = 12, height = 9, dpi = 600)
ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/C1C6/Volcano_Directional_Categories.pdf", pv, width = 12, height = 9, dpi = 600)
```

#check those overlapping proteins in burnout c1 and c6 values with meanigful change.

```{r}
q2_cleanC1C6<- read.csv("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/C1C6/results_q2_cleanC1C6.csv")

meanigfulChange<- read.csv("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/result_StreamlitApp/Meaningful_Change1_3/AcrossIndi_GroupAnalysis_Indepen.csv")

colnames(q2_cleanC1C6)

dxx<- merge(q2_cleanC1C6, meanigfulChange, by.x = "protein", by.y = "Protein")
write.csv(dxx, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/C1C6/MeanigfulAssociated.csv")
```

