---
title: "lastingChange_traj"
author: "Nasim Bararpour"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Category Data
```{r}
full_data<- read.csv("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/fullData.csv")


dx, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Q2_proteins_FinalGeneID.csv")

colnames(full_data)
```
# Long fromat data 
```{r}
library(dplyr)
library(tidyr)
 
# ============================================
# STEP 1: PREPARE DATA WITH ALL COLUMNS
# ============================================

cat("=== Preparing aggregated dataset ===\n")

# Assuming your data structure is:
# dex columns: [other columns] | code (ID) | Time | Value (burnout) | sex | age | Protein1 | Protein2 | ... | Protein5400

# Extract components
meta_cols <- full_data[, c("code", "sex", "age", "Time", "Value","burnout", "time_numeric" )]  # Adjust column names as needed
protein_cols <- Ex  # Proteins columns use EX

# Combine
full_data <- cbind(meta_cols, protein_cols)

write.csv(full_data, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/full_dataExOriginal.csv")


# Rename for clarity
colnames(full_data)[c(1,4,6)] <- c("ID", "Timepoint", "Burnout")


cat("Original data dimensions:", dim(full_data), "\n")
cat("Columns: ID, Timepoint, Burnout, sex, age, +", ncol(protein_cols), "proteins\n\n")
```

# ============================================
# STEP 2: AGGREGATE BY ID AND TIMEPOINT
# ============================================

```{r}
# ============================================
# SIMPLEST AGGREGATION APPROACH
# ============================================

# Specify your columns explicitly
data_to_aggregate <- data.frame(
  ID = full_data$ID,
  Timepoint = full_data$Timepoint,
  Burnout = full_data$Burnout,
  sex = full_data$sex,
  age = full_data$age,
  Ex  # All protein columns
)

# Aggregate using aggregate() function
library(dplyr)

aggregated <- data_to_aggregate %>%
  group_by(ID, Timepoint) %>%
  summarise(
    Burnout = mean(Burnout, na.rm = TRUE),
    sex = first(sex),
    age = mean(age, na.rm = TRUE),
    across(where(is.numeric) & !any_of(c("Burnout", "age")), ~mean(.x, na.rm = TRUE)),
    .groups = 'drop'
  )

write.csv(aggregated, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Aggregated_Simple.csv", row.names = FALSE)
cat("Saved: Aggregated_Simple.csv\n")
cat("Dimensions:", dim(aggregated), "\n")

aggregated_data<-aggregated
```


# ============================================
# STEP 3: VERIFY AGGREGATION
# ============================================

```{r}
cat("=== Verification ===\n")

# Check observations per ID
obs_per_id <- table(table(aggregated_data$ID))
cat("Observations per ID:\n")
print(obs_per_id)
cat("\n")

# Check timepoints
cat("Timepoints present:\n")
print(table(aggregated_data$Timepoint))
cat("\n")

# Sample of aggregated data
cat("Sample of aggregated data (first 5 rows, first 8 columns):\n")
print(head(aggregated_data[, 1:min(8, ncol(aggregated_data))], 5))
cat("\n")

# ============================================
# STEP 4: SAVE AGGREGATED DATASET
# ============================================

# Save complete aggregated data
write.csv(aggregated_data, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Aggregated_Data_by_ID_Timepoint.csv", row.names = FALSE)
cat("Saved: Aggregated_Data_by_ID_Timepoint.csv\n\n")
```


# ============================================
# STEP 5: CREATE SUMMARY STATISTICS
# ============================================
```{r}
cat("=== Summary Statistics ===\n")

# Summary by timepoint
summary_by_timepoint <- aggregated_data %>%
  group_by(Timepoint) %>%
  summarise(
    n_subjects = n(),
    mean_burnout = mean(Burnout, na.rm = TRUE),
    sd_burnout = sd(Burnout, na.rm = TRUE),
    min_burnout = min(Burnout, na.rm = TRUE),
    max_burnout = max(Burnout, na.rm = TRUE)
  )

cat("Summary by Timepoint:\n")
print(summary_by_timepoint)
cat("\n")

write.csv(summary_by_timepoint, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Summary_by_Timepoint.csv", row.names = FALSE)

# Summary by ID
summary_by_id <- aggregated_data %>%
  group_by(ID) %>%
  summarise(
    n_timepoints = n(),
    mean_burnout = mean(Burnout, na.rm = TRUE),
    sd_burnout = sd(Burnout, na.rm = TRUE),
    burnout_change = last(Burnout) - first(Burnout),  # T6 - T1
    sex = first(sex),
    age = first(age)
  )

cat("Summary by ID (first 10):\n")
print(head(summary_by_id, 10))
cat("\n")

write.csv(summary_by_id, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Summary_by_ID.csv", row.names = FALSE)

# ============================================
# STEP 6: CREATE WIDE FORMAT (ONE ROW PER ID)
# ============================================

cat("=== Creating wide format (one row per ID) ===\n")

library(tidyverse)

# Pivot to wide format: one row per ID, columns for each timepoint
wide_data <- aggregated_data %>%
  pivot_wider(
    id_cols = c(ID, sex, age),
    names_from = Timepoint,
    values_from = c(Burnout, all_of(colnames(protein_cols))),
    names_sep = "_"
  )

cat("Wide format dimensions:", dim(wide_data), "\n")
cat("Columns: ID, sex, age, + Burnout_T1-T6 + each protein at T1-T6\n\n")

write.csv(wide_data, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Wide_Format_by_ID.csv", row.names = FALSE)
cat("Saved: Wide_Format_by_ID.csv\n\n")

# ============================================
# STEP 7: EXTRACT SPECIFIC FORMATS FOR ANALYSIS
# ============================================

# Format 1: Long format with only ID, Timepoint, Burnout, Proteins
long_format_simple <- aggregated_data %>%
  select(ID, Timepoint, Burnout, all_of(colnames(protein_cols)))

write.csv(long_format_simple, "Long_Format_Simple.csv", row.names = FALSE)
cat("Saved: Long_Format_Simple.csv\n")
cat("  Format: ID | Timepoint | Burnout | Protein1 | ... | Protein5400\n\n")

# Format 2: Matrix format for proteins only (for correlation analysis)
protein_matrix <- as.matrix(aggregated_data[, colnames(protein_cols)])
rownames(protein_matrix) <- paste(aggregated_data$ID, 
                                  aggregated_data$Timepoint, 
                                  sep = "_")
write.csv(protein_matrix, "Protein_Matrix_Averaged.csv", row.names = TRUE)
cat("Saved: Protein_Matrix_Averaged.csv\n")
cat("  Format: Rows = ID_Timepoint, Columns = Proteins\n\n")

# ============================================
# STEP 8: VISUALIZE AGGREGATION RESULTS
# ============================================

library(ggplot2)

# Plot 1: Burnout across timepoints
p1 <- ggplot(aggregated_data, aes(x = Timepoint, y = Burnout)) +
  geom_boxplot(fill = "lightblue", alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.3, size = 0.8) +
  stat_summary(fun = mean, geom = "point", 
               shape = 23, size = 3, fill = "red") +
  theme_minimal() +
  labs(
    title = "Burnout Distribution Across Timepoints (Aggregated Data)",
    subtitle = "Red diamonds = mean",
    x = "Timepoint",
    y = "Burnout Score"
  )

print(p1)
ggsave("Burnout_by_Timepoint_Aggregated.png", p1, width = 10, height = 6, dpi = 300)

# Plot 2: Individual trajectories
p2 <- ggplot(aggregated_data, aes(x = Timepoint, y = Burnout, group = ID)) +
  geom_line(alpha = 0.3, color = "gray50") +
  geom_point(alpha = 0.4, size = 1) +
  stat_summary(aes(group = 1), fun = mean, geom = "line", 
               color = "red", linewidth = 1.5) +
  stat_summary(aes(group = 1), fun = mean, geom = "point", 
               color = "red", size = 3) +
  theme_minimal() +
  labs(
    title = "Individual Burnout Trajectories (Aggregated Data)",
    subtitle = "Red line = mean trajectory",
    x = "Timepoint",
    y = "Burnout Score"
  )

print(p2)
ggsave("Individual_Trajectories_Aggregated.png", p2, width = 10, height = 6, dpi = 300)

# Plot 3: Sample size at each timepoint
sample_size_plot <- ggplot(summary_by_timepoint, 
                           aes(x = Timepoint, y = n_subjects)) +
  geom_bar(stat = "identity", fill = "steelblue", color = "black") +
  geom_text(aes(label = n_subjects), vjust = -0.5, size = 5, fontface = "bold") +
  theme_minimal() +
  labs(
    title = "Sample Size at Each Timepoint",
    x = "Timepoint",
    y = "Number of Subjects"
  ) +
  ylim(0, max(summary_by_timepoint$n_subjects) * 1.15)

print(sample_size_plot)
ggsave("Sample_Size_by_Timepoint.png", sample_size_plot, 
       width = 8, height = 6, dpi = 300)

# ============================================
# STEP 9: CHECK FOR DUPLICATES/ISSUES
# ============================================

cat("=== Quality Checks ===\n")

# Check for duplicate ID-Timepoint combinations
duplicates <- aggregated_data %>%
  group_by(ID, Timepoint) %>%
  filter(n() > 1)

if(nrow(duplicates) > 0) {
  cat("WARNING: Found", nrow(duplicates), "duplicate ID-Timepoint combinations!\n")
  print(head(duplicates))
} else {
  cat("✓ No duplicate ID-Timepoint combinations\n")
}

# Check for missing values in key columns
missing_burnout <- sum(is.na(aggregated_data$Burnout))
missing_proteins <- sum(is.na(aggregated_data[, colnames(protein_cols)]))

cat("Missing values:\n")
cat("  Burnout:", missing_burnout, "\n")
cat("  Proteins:", missing_proteins, "total (", 
    round(missing_proteins / (nrow(aggregated_data) * ncol(protein_cols)) * 100, 2), 
    "%)\n\n")

# ============================================
# STEP 10: CREATE METADATA FILE
# ============================================

metadata_aggregated <- aggregated_data %>%
  select(ID, Timepoint, Burnout, sex, age, n_obs)

write.csv(metadata_aggregated, "Metadata_Aggregated.csv", row.names = FALSE)
cat("Saved: Metadata_Aggregated.csv\n")

# ============================================
# SUMMARY REPORT
# ============================================

cat("\n========================================\n")
cat("AGGREGATION COMPLETE\n")
cat("========================================\n\n")

cat("FILES CREATED:\n")
cat("1. Aggregated_Data_by_ID_Timepoint.csv\n")
cat("   - Long format: ID | Timepoint | Burnout | Proteins (averaged)\n")
cat("   - Dimensions:", dim(aggregated_data), "\n\n")

cat("2. Wide_Format_by_ID.csv\n")
cat("   - Wide format: One row per ID, columns for each timepoint\n")
cat("   - Dimensions:", dim(wide_data), "\n\n")

cat("3. Long_Format_Simple.csv\n")
cat("   - Simplified: ID | Timepoint | Burnout | Proteins only\n\n")

cat("4. Protein_Matrix_Averaged.csv\n")
cat("   - Matrix format: Rows = ID_Timepoint, Columns = Proteins\n\n")

cat("5. Summary files:\n")
cat("   - Summary_by_Timepoint.csv\n")
cat("   - Summary_by_ID.csv\n")
cat("   - Metadata_Aggregated.csv\n\n")

cat("6. Visualizations:\n")
cat("   - Burnout_by_Timepoint_Aggregated.png\n")
cat("   - Individual_Trajectories_Aggregated.png\n")
cat("   - Sample_Size_by_Timepoint.png\n\n")

cat("========================================\n\n")
```

```{r}
pcomb<- p1|p2|sample_size_plot
print(pcomb)
ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/pcombTraj.png", pcomb, 
       width = 10, height = 5, dpi = 300)

ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/pcombTraj.pdf", pcomb, 
       width = 10, height = 5, dpi = 300)
```

4-category analysis on aggregated data
# ============================================
# STEP 1: CALCULATE WITHIN-SUBJECT CHANGES
# ============================================
```{r}
library(dplyr)
library(ggplot2)
library(ggrepel)
library(lme4)
library(lmerTest)
library(tidyr)

# ============================================
# STEP 1: CALCULATE WITHIN-SUBJECT CHANGES
# ============================================

cat("=== Calculating Within-Subject Changes ===\n\n")

# Assuming 'result' is your aggregated data from previous step
# If you named it differently, adjust accordingly
aggregated_data<-read.csv("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Aggregated_Data_by_ID_Timepoint.csv")
```

```{r}
# Calculate mean for each timepoint
means_by_timepoint <- aggregated_data %>%
  group_by(Timepoint) %>%
  summarise(
    n_subjects = n(),
    mean_burnout = mean(Burnout, na.rm = TRUE),
    across(all_of(protein_cols), ~mean(.x, na.rm = TRUE))
  ) %>%
  arrange(Timepoint)

cat("Mean values calculated for each timepoint:\n")
print(means_by_timepoint[, 1:8])  # Show first few columns
cat("\n")

# Save means
write.csv(means_by_timepoint, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Mean_Values_by_Timepoint.csv", row.names = FALSE)
```


# ============================================
# STEP 2: ANALYZE BURNOUT TRAJECTORY
# ============================================

```{r}
library(tidyverse)
means_by_timepoint<- read.csv("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Olink_Cat4_Cat2/Mean_Values_by_Timepoint.csv")


cat("=== Analyzing Burnout Trajectory ===\n\n")

# Get burnout trajectory
burnout_trajectory <- means_by_timepoint %>%
  select(Timepoint, mean_burnout) %>%
  arrange(Timepoint)

cat("Burnout trajectory (T1 to T6):\n")
print(burnout_trajectory)
cat("\n")

# Calculate burnout change from T1 to T6
burnout_t1 <- burnout_trajectory$mean_burnout[burnout_trajectory$Timepoint == "C1"]
burnout_t3 <- burnout_trajectory$mean_burnout[burnout_trajectory$Timepoint == "C3"]
burnout_t6 <- burnout_trajectory$mean_burnout[burnout_trajectory$Timepoint == "C6"]

burnout_overall_change <-burnout_t6 - burnout_t1

cat("Burnout T1:", round(burnout_t1, 3), "\n")
cat("Burnout T3:", round(burnout_t3, 3), "\n")
cat("Burnout T6:", round(burnout_t6, 3), "\n")
cat("Overall change (T6-T1):", round(burnout_overall_change, 3), "\n")
cat("Direction:", ifelse(burnout_overall_change > 0, "INCREASING", "DECREASING"), "\n\n")
```


# ============================================
# STEP 3: ANALYZE PROTEIN TRAJECTORIES
# ============================================
```{r}
# ============================================
# STEP 3: ANALYZE PROTEIN TRAJECTORIES
# ============================================

cat("=== Analyzing Protein Trajectories ===\n\n")

# Function to categorize one protein based on trajectory
categorize_protein_trajectory <- function(protein_name, means_by_timepoint, burnout_trajectory) {
  
  # Get protein trajectory
  protein_values <- means_by_timepoint[[protein_name]]
  timepoints <- means_by_timepoint$Timepoint
  
  # Get T1 and T6 values
  protein_t1 <- protein_values[timepoints == "C1"]
 # protein_t2 <- protein_values[timepoints == "C2"]
 # protein_t3 <- protein_values[timepoints == "C3"]
  #protein_t4 <- protein_values[timepoints == "C4"]
 # protein_t5 <- protein_values[timepoints == "C5"]
  protein_t6 <- protein_values[timepoints == "C6"]
  
  # Calculate change if it is lasting 
  protein_change <- protein_t6 - protein_t1
  
  # Calculate correlation with burnout
  cor_result <- cor.test(burnout_trajectory$mean_burnout, protein_values)
  correlation <- cor_result$estimate
  cor_pvalue <- cor_result$p.value
  
  # Linear trend test
  time_numeric <- 1:length(timepoints)
  lm_protein <- lm(protein_values ~ time_numeric)
  protein_slope <- coef(lm_protein)[2]
  protein_trend_p <- summary(lm_protein)$coefficients[2, 4]
  
  lm_burnout <- lm(burnout_trajectory$mean_burnout ~ time_numeric)
  burnout_slope <- coef(lm_burnout)[2]
  
  # Determine category based on trajectories
  category <- "No clear pattern"
  
  # Check if changes are significant (using arbitrary but reasonable thresholds)
  protein_change_pct <- abs(protein_change) / abs(protein_t1) * 100
  
  if(protein_trend_p < 0.05) {  # Protein has significant trend
    
    if(protein_slope > 0 && burnout_slope > 0) {
      # Both increasing
      category <- "Cat1: Protein ↑ as Burnout ↑ (T1→T6)"
      
    } else if(protein_slope < 0 && burnout_slope < 0) {
      # Both decreasing
      category <- "Cat2: Protein ↓ as Burnout ↓ (T1→T6)"
      
    } else if(protein_slope < 0 && burnout_slope > 0) {
      # Protein decreasing while burnout increasing
      category <- "Cat3: Protein ↓ as Burnout ↑ (T1→T6)"
      
    } else if(protein_slope > 0 && burnout_slope < 0) {
      # Protein increasing while burnout decreasing
      category <- "Cat4: Protein ↑ as Burnout ↓ (T1→T6)"
    }
  }
  
  return(data.frame(
    protein = protein_name,
    protein_T1 = protein_t1,
    #protein_T3 = protein_t3,
    protein_T6 = protein_t6,
    protein_change = protein_change,
    protein_change_pct = protein_change_pct,
    protein_slope = protein_slope,
    protein_trend_p = protein_trend_p,
    correlation_with_burnout = correlation,
    cor_pvalue = cor_pvalue,
    category = category,
    stringsAsFactors = FALSE
  ))
}

# Analyze all proteins
cat("Analyzing trajectories for", length(protein_cols), "proteins...\n")

trajectory_results <- data.frame()

pb <- txtProgressBar(min = 0, max = length(protein_cols), style = 3)

for(i in 1:length(protein_cols)) {
  result_i <- categorize_protein_trajectory(protein_cols[i], 
                                             means_by_timepoint, 
                                             burnout_trajectory)
  trajectory_results <- rbind(trajectory_results, result_i)
  setTxtProgressBar(pb, i)
}

close(pb)

# Multiple testing correction
trajectory_results$fdr_trend <- p.adjust(trajectory_results$protein_trend_p, method = "BH")
trajectory_results$fdr_correlation <- p.adjust(trajectory_results$cor_pvalue, method = "BH")

cat("\n\n")
unique(trajectory_results$category)
colnames(trajectory_results)
write.csv(trajectory_results, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Olink_Cat4_Cat2/trajectory_results.csv")
```

# ============================================
# STEP 4: EXTRACT 4 CATEGORIES
# ============================================
```{r}
trajectory_results<- read.csv("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Olink_Cat4_Cat2/trajectory_results.csv")


cat("=== Extracting 4 Categories ===\n\n")

# Category 1: Protein increases as burnout increases
cat1_traj <- trajectory_results[
  trajectory_results$category == "Cat1: Protein ↑ as Burnout ↑ (T1→T6)" &
  trajectory_results$cor_pvalue < 0.05, ]
cat1_traj <- cat1_traj[order(-cat1_traj$protein_slope), ]

cat("CATEGORY 1: Proteins INCREASE as Burnout INCREASES (T1→T6)\n")
cat("Count:", nrow(cat1_traj), "\n")
if(nrow(cat1_traj) > 0) {
  cat("Top 10:\n")
  print(cat1_traj[1:min(10, nrow(cat1_traj)), 
                  c("protein", "protein_change", "protein_slope", "correlation_with_burnout", "protein_trend_p")])
}
cat("\n")

# Category 2: Protein decreases as burnout decreases
cat2_traj <- trajectory_results[
  trajectory_results$category == "Cat2: Protein ↓ as Burnout ↓ (T1→T6)" &
  trajectory_results$cor_pvalue < 0.05, ]
cat2_traj <- cat2_traj[order(cat2_traj$protein_slope), ]

cat("CATEGORY 2: Proteins DECREASE as Burnout DECREASES (T1→T6)\n")
cat("Count:", nrow(cat2_traj), "\n")
if(nrow(cat2_traj) > 0) {
  cat("Top 10:\n")
  print(cat2_traj[1:min(10, nrow(cat2_traj)), 
                  c("protein", "protein_change", "protein_slope", "correlation_with_burnout", "protein_trend_p")])
}
cat("\n")

# Category 3: Protein decreases as burnout increases
cat3_traj <- trajectory_results[
  trajectory_results$category == "Cat3: Protein ↓ as Burnout ↑ (T1→T6)" &
  trajectory_results$cor_pvalue < 0.05, ]
cat3_traj <- cat3_traj[order(cat3_traj$protein_slope), ]

cat("CATEGORY 3: Proteins DECREASE as Burnout INCREASES (T1→T6)\n")
cat("Count:", nrow(cat3_traj), "\n")
if(nrow(cat3_traj) > 0) {
  cat("Top 10:\n")
  print(cat3_traj[1:min(10, nrow(cat3_traj)), 
                  c("protein", "protein_change", "protein_slope", "correlation_with_burnout", "protein_trend_p")])
}
cat("\n")

# Category 4: Protein increases as burnout decreases
cat4_traj <- trajectory_results[
  trajectory_results$category == "Cat4: Protein ↑ as Burnout ↓ (T1→T6)" &
  trajectory_results$cor_pvalue < 0.05, ]
cat4_traj <- cat4_traj[order(-cat4_traj$protein_slope), ]

cat("CATEGORY 4: Proteins INCREASE as Burnout DECREASES (T1→T6)\n")
cat("Count:", nrow(cat4_traj), "\n")
if(nrow(cat4_traj) > 0) {
  cat("Top 10:\n")
  print(cat4_traj[1:min(10, nrow(cat4_traj)), 
                  c("protein", "protein_change", "protein_slope", "correlation_with_burnout", "protein_trend_p")])
}
cat("\n")
```

# ============================================
# STEP 5: SAVE RESULTS
# ============================================

```{r}
# ============================================
# STEP 5: SAVE RESULTS
# ============================================

write.csv(trajectory_results, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Trajectory_Analysis_All_Proteins.csv", row.names = FALSE)
write.csv(cat1_traj, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/Cat1_Trajectory_Protein_UP_Burnout_UP.csv", row.names = FALSE)
write.csv(cat2_traj, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/Cat2_Trajectory_Protein_DOWN_Burnout_DOWN.csv", row.names = FALSE)
write.csv(cat3_traj, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/Cat3_Trajectory_Protein_DOWN_Burnout_UP.csv", row.names = FALSE)
write.csv(cat4_traj, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/Cat4_Trajectory_Protein_UP_Burnout_DOWN.csv", row.names = FALSE)

cat("Files saved!\n\n")
```

# ============================================
# VISUALIZATION 1: BURNOUT TRAJECTORY
# ============================================
```{r}
# ============================================
# VISUALIZATION 1: BURNOUT TRAJECTORY
# ============================================

p_burnout <- ggplot(burnout_trajectory, aes(x = Timepoint, y = mean_burnout, group = 1)) +
  geom_line(linewidth = 1.5, color = "black") + #D32F2F
  geom_point(size = 4, color = "black") + #D32F2F
  geom_text(aes(label = round(mean_burnout, 2)), vjust = -1, size = 4, fontface = "bold") +
  theme_minimal(base_size = 14) +
  labs(
    title = "Burnout Trajectory Across Timepoints",
    subtitle = paste0("Overall change: ", round(burnout_overall_change, 3)),
    x = "Timepoint",
    y = "Mean Burnout Score"
  ) +
  theme(panel.grid.minor = element_blank())

print(p_burnout)
ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Burnout_Trajectory_T1_to_T6.png", p_burnout, width = 10, height = 6, dpi = 300)
```

# ============================================
# VISUALIZATION 2: BAR CHART OF CATEGORIES
# ============================================

```{r}
# ============================================
# VISUALIZATION 2: BAR CHART OF CATEGORIES
# ============================================

summary_traj <- data.frame(
  Category = c(
    "Cat 1:\nProtein ↑\nBurnout ↑",
    "Cat 2:\nProtein ↓\nBurnout ↓",
    "Cat 3:\nProtein ↓\nBurnout ↑",
    "Cat 4:\nProtein ↑\nBurnout ↓"
  ),
  Count = c(nrow(cat1_traj), nrow(cat2_traj), nrow(cat3_traj), nrow(cat4_traj)),
  Type = c("Parallel", "Parallel", "Opposite", "Opposite")
)

p_bar <- ggplot(summary_traj, aes(x = Category, y = Count, fill = Type)) +
  geom_bar(stat = "identity", color = "black", linewidth = 1) +
  geom_text(aes(label = Count), vjust = -0.5, size = 8, fontface = "bold") +
  scale_fill_manual(values = c("Parallel" = "#E53935", "Opposite" = "#1E88E5")) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(size = 12, face = "bold"),
    panel.grid.major.x = element_blank()
  ) +
  labs(
    title = "Protein Trajectories Relative to Burnout (T1→T6)",
    subtitle = "Based on average values at each timepoint",
    x = "",
    y = "Number of Proteins",
    fill = "Trajectory Type"
  ) +
  ylim(0, max(summary_traj$Count) * 1.2)

print(p_bar)

pTraj<- (p_burnout/p_bar)
print(pTraj)

ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/BarChart_4Categories_Trajectories.png", p_bar, width = 10, height = 7, dpi = 300)


ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/BarChart_4Categories_Trajectories.pdf", pTraj, width = 5, height = 7, dpi = 300)

```

# ============================================
# VISUALIZATION 3: EXAMPLE TRAJECTORIES
# ============================================

```{r}
# ============================================
# VISUALIZATION 3: EXAMPLE TRAJECTORIES
# ============================================
means_data<-means_by_timepoint
burnout_traj<-burnout_trajectory
# Function to plot trajectory for one protein
plot_trajectory <- function(protein_name, means_data, burnout_trajectory, category_name) {
  
  # Prepare data
  plot_data <- data.frame(
    Timepoint = means_data$Timepoint,
    Protein = means_data[[protein_name]],
    Burnout = burnout_trajectory$mean_burnout
  )
  
  # Normalize to 0-1 scale for comparison
  plot_data$Protein_norm <- (plot_data$Protein - min(plot_data$Protein)) / 
                             (max(plot_data$Protein) - min(plot_data$Protein))
  plot_data$Burnout_norm <- (plot_data$Burnout - min(plot_data$Burnout)) / 
                             (max(plot_data$Burnout) - min(plot_data$Burnout))
  
  # Reshape for plotting
  plot_long <- plot_data %>%
    select(Timepoint, Protein_norm, Burnout_norm) %>%
    pivot_longer(cols = c(Protein_norm, Burnout_norm),
                 names_to = "Variable",
                 values_to = "Value")
  
  p <- ggplot(plot_long, aes(x = Timepoint, y = Value, 
                              color = Variable, group = Variable)) +
    geom_line(linewidth = 1.2) +
    geom_point(size = 3) +
    scale_color_manual(
      values = c("Protein_norm" = "#2196F3", "Burnout_norm" = "#F44336"),
      labels = c("Burnout", "Protein")
    ) +
    theme_minimal(base_size = 12) +
    labs(
      title = paste0(category_name, ": ", protein_name),
      subtitle = "Normalized trajectories (0-1 scale)",
      x = "Timepoint",
      y = "Normalized Value",
      color = ""
    ) +
    theme(legend.position = "bottom")
  
  return(p)
}

# Plot examples from each category
example_plots <- list()

if(nrow(cat1_traj) > 0) {
  example_plots[[1]] <- plot_trajectory(cat1_traj$protein[1], means_by_timepoint, 
                                         burnout_trajectory, "Category 1")
}

if(nrow(cat2_traj) > 0) {
  example_plots[[2]] <- plot_trajectory(cat2_traj$protein[1], means_by_timepoint, 
                                         burnout_trajectory, "Category 2")
}

if(nrow(cat3_traj) > 0) {
  example_plots[[3]] <- plot_trajectory(cat3_traj$protein[1], means_by_timepoint, 
                                         burnout_trajectory, "Category 3")
}

if(nrow(cat4_traj) > 0) {
  example_plots[[4]] <- plot_trajectory(cat4_traj$protein[1], means_by_timepoint, 
                                         burnout_trajectory, "Category 4")
}

# Only combine if we have plots - FIXED
if(length(example_plots) > 0) {
  # Calculate grid layout
  n_plots <- length(example_plots)
  ncol <- 2
  nrow <- ceiling(n_plots / ncol)
  
  # Save individually first
  for(i in 1:length(example_plots)) {
    ggsave(paste0("Example_Trajectory_", names(example_plots)[i], ".png"), 
           example_plots[[i]], width = 7, height = 5, dpi = 300)
  }
  
  # Then try to combine
  tryCatch({
    p_combined <- grid.arrange(grobs = example_plots, ncol = ncol)
    ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Example_Trajectories_4Categories_Combined.png", p_combined, 
           width = 14, height = nrow * 5, dpi = 300)
  }, error = function(e) {
    cat("Note: Combined plot failed, but individual plots saved successfully\n")
  })
}
```

# ============================================
# VISUALIZATION 4: TOP 5 TRAJECTORIES - FIXED
# ============================================

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)
library(patchwork)

# ============================================
# FUNCTION: PLOT ALL PROTEINS IN A CATEGORY
# ============================================

plot_category_trajectories <- function(protein_list, means_data, burnout_traj, 
                                       category_name, color, n_proteins = NULL) {
  
  # Get proteins (all or specified number)
  if(is.null(n_proteins)) {
    proteins_to_plot <- protein_list$protein
  } else {
    proteins_to_plot <- head(protein_list$protein, n_proteins)
  }
  
  if(length(proteins_to_plot) == 0) {
    cat("No proteins to plot for", category_name, "\n")
    return(NULL)
  }
  
  cat("Plotting", length(proteins_to_plot), "proteins for", category_name, "\n")
  
  # Prepare protein data
  plot_data <- means_data %>%
    select(Timepoint, all_of(proteins_to_plot)) %>%
    pivot_longer(cols = all_of(proteins_to_plot),
                 names_to = "Protein",
                 values_to = "Value") %>%
    group_by(Protein) %>%
    mutate(Value_norm = (Value - min(Value, na.rm = TRUE)) / 
                        (max(Value, na.rm = TRUE) - min(Value, na.rm = TRUE))) %>%
    ungroup()
  
  # Prepare burnout data
  burnout_data <- data.frame(
    Timepoint = burnout_traj$Timepoint,
    Protein = "Burnout",
    Value_norm = (burnout_traj$mean_burnout - min(burnout_traj$mean_burnout)) /
                 (max(burnout_traj$mean_burnout) - min(burnout_traj$mean_burnout))
  )
  
  # Combine
  plot_data_combined <- rbind(
    plot_data[, c("Timepoint", "Protein", "Value_norm")], 
    burnout_data
  )
  
  # Create overlay plot (all proteins on one plot)
  p_overlay <- ggplot(plot_data_combined, 
                      aes(x = Timepoint, y = Value_norm, 
                          color = Protein, group = Protein)) +
    geom_line(aes(linewidth = Protein == "Burnout", 
                  alpha = Protein == "Burnout")) +
    geom_point(data = filter(plot_data_combined, Protein == "Burnout"),
               size = 3) +
    scale_linewidth_manual(values = c("TRUE" = 2.5, "FALSE" = 0.8), guide = "none") +
    scale_alpha_manual(values = c("TRUE" = 1, "FALSE" = 0.6), guide = "none") +
    scale_color_manual(
      values = c("Burnout" = "#F44336", 
                 setNames(rep(color, length(proteins_to_plot)), proteins_to_plot)),
      breaks = "Burnout"
    ) +
    theme_minimal(base_size = 14) +
    theme(
      legend.position = "bottom",
      legend.title = element_blank(),
      panel.grid.minor = element_blank(),
      plot.title = element_text(face = "bold", size = 16),
      plot.subtitle = element_text(size = 12)
    ) + theme_bw()+
    labs(
      title = category_name,
      subtitle = paste(length(proteins_to_plot), "proteins (normalized 0-1) + Burnout trajectory"),
      x = "Timepoint",
      y = "Normalized Expression (0-1)"
    )
  
  # Create faceted plot (individual panels for each protein)
  n_cols <- ceiling(sqrt(length(proteins_to_plot)))
  
  p_faceted <- ggplot(plot_data, 
                      aes(x = Timepoint, y = Value_norm, group = Protein)) +
    geom_line(color = color, linewidth = 1) +
    geom_point(color = color, size = 2) +
    # Add burnout trajectory to each panel
    geom_line(data = burnout_data,
              aes(x = Timepoint, y = Value_norm),
              color = "#F44336", linewidth = 1.5, linetype = "dashed") +
    facet_wrap(~ Protein, ncol = n_cols, scales = "free_y") +
    theme_minimal(base_size = 10) +
    theme(
      strip.text = element_text(size = 7, face = "bold"),
      axis.text = element_text(size = 7),
      panel.spacing = unit(0.3, "lines"),
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(size = 10)
    ) +
    labs(
      title = paste(category_name, "- Individual Trajectories"),
      subtitle = "Black = Protein | Red dashed = Burnout",
      x = "Timepoint",
      y = "Normalized Expression"
    ) + theme_bw()
  
  return(list(overlay = p_overlay, faceted = p_faceted))
}

# ============================================
# PLOT CATEGORY 2 (58 PROTEINS)
# ============================================

cat("\n=== Plotting Category 2 (58 proteins) ===\n\n")

if(exists("cat2_traj") && nrow(cat2_traj) > 0) {
  
  cat2_plots <- plot_category_trajectories(
    protein_list = cat2_traj,
    means_data = means_by_timepoint,
    burnout_traj = burnout_trajectory,
    category_name = "Category 2: Protein ↓ as Burnout ↓",
    color = "#FF6F00",
    n_proteins = NULL  # Plot all proteins
  )
  
  if(!is.null(cat2_plots)) {
    
    # Save overlay plot
    print(cat2_plots$overlay)
    ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Olink_Cat4_Cat2/Cat2_Overlay_58_Proteins.png", 
           cat2_plots$overlay, 
           width = 10, 
           height = 8, 
           dpi = 300)
     ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Olink_Cat4_Cat2/Cat2_Overlay_58_Proteins.pdf", 
           cat2_plots$overlay, 
           width = 10, 
           height = 8, 
           dpi = 300)
    cat("✓ Cat2_Overlay_58_Proteins.png\n")
    
    # Save faceted plot
    print(cat2_plots$faceted)
    ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Olink_Cat4_Cat2/Cat2_Faceted_58_Proteins.png", 
           cat2_plots$faceted, 
           width = 16, 
           height = 20, 
           dpi = 300)
      ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Olink_Cat4_Cat2/Cat2_Faceted_58_Proteins.pdf", 
           cat2_plots$faceted, 
           width = 16, 
           height = 20, 
           dpi = 300)
    cat("✓ Cat2_Faceted_58_Proteins.png\n\n")
  }
  
} else {
  cat("Category 2 not found or empty\n\n")
}
```

# ============================================
# PLOT CATEGORY 4 (20 PROTEINS)
# ============================================
```{r}
# ============================================
# PLOT CATEGORY 4 (20 PROTEINS)
# ============================================

cat("=== Plotting Category 4 (20 proteins) ===\n\n")

if(exists("cat4_traj") && nrow(cat4_traj) > 0) {
  
  cat4_plots <- plot_category_trajectories(
    protein_list = cat4_traj,
    means_data = means_by_timepoint,
    burnout_traj = burnout_trajectory,
    category_name = "Category 4: Protein ↑ as Burnout ↓",
    color = "#388E3C",
    n_proteins = NULL  # Plot all proteins
  )
  
  if(!is.null(cat4_plots)) {
    
    # Save overlay plot
    print(cat4_plots$overlay)
    ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Olink_Cat4_Cat2/Cat4_Overlay_20_Proteins.png", 
           cat4_plots$overlay, 
           width = 10, 
           height = 8, 
           dpi = 300)
     ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Olink_Cat4_Cat2/Cat4_Overlay_20_Proteins.pdf", 
           cat4_plots$overlay, 
           width = 10, 
           height = 8, 
           dpi = 300)
    cat("✓ Cat4_Overlay_20_Proteins.png\n")
    
    # Save faceted plot
    print(cat4_plots$faceted)
    ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Olink_Cat4_Cat2/Cat4_Faceted_20_Proteins.png", 
           cat4_plots$faceted, 
           width = 12, 
           height = 10, 
           dpi = 300)
      ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Olink_Cat4_Cat2/Cat4_Faceted_20_Proteins.pdf", 
           cat4_plots$faceted, 
           width = 12, 
           height = 10, 
           dpi = 300)
    cat("✓ Cat4_Faceted_20_Proteins.png\n\n")
  }
  
} else {
  cat("Category 4 not found or empty\n\n")
}
```

# ============================================
# ALTERNATIVE: HEATMAP VISUALIZATION
# ============================================

```{r}
means_by_timepoint<-means_by_timepoint %>% filter(Timepoint %in% c("C1", "C2", "C3", "C4", "C5", "C6"))

means_data<-means_by_timepoint
burnout_traj<-burnout_trajectory



```

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

# ============================================
# FIXED: HEATMAP WITH DUPLICATE HANDLING
# ============================================

create_category_heatmap <- function(protein_list, means_data, category_name, color) {
  
  # Get unique proteins only (remove duplicates)
  proteins <- unique(protein_list$protein)
  
  if(length(proteins) == 0) {
    cat("No proteins to plot for", category_name, "\n")
    return(NULL)
  }
  
  cat("Creating heatmap for", length(proteins), "unique proteins\n")
  
  # Check for duplicates in input
  if(length(proteins) < length(protein_list$protein)) {
    cat("  WARNING: Removed", length(protein_list$protein) - length(proteins), "duplicate proteins\n")
  }
  
  # Get data and normalize
  heatmap_data <- means_data %>%
    select(Timepoint, all_of(proteins)) %>%
    pivot_longer(cols = all_of(proteins),
                 names_to = "Protein",
                 values_to = "Value") %>%
    # Remove any duplicate Protein-Timepoint combinations
    distinct(Protein, Timepoint, .keep_all = TRUE) %>%
    group_by(Protein) %>%
    mutate(Z_score = scale(Value)[,1]) %>%
    ungroup()
  
  # Order proteins by their trajectory pattern (use unique proteins only)
  protein_order <- proteins  # Already unique
  
  # Verify no duplicates in protein_order
  if(any(duplicated(protein_order))) {
    cat("  WARNING: Found duplicates in protein_order, removing...\n")
    protein_order <- unique(protein_order)
  }
  
  # Set factor levels
  heatmap_data$Protein <- factor(heatmap_data$Protein, levels = protein_order)
  
  p <- ggplot(heatmap_data, aes(x = Timepoint, y = Protein, fill = Z_score)) +
    geom_tile(color = "white", linewidth = 0.5) +
    scale_fill_gradient2(
      low = "#2166AC", mid = "white", high = "#B2182B",
      midpoint = 0,
      limits = c(-3, 3),
      oob = scales::squish,
      name = "Z-score"
    ) +
    theme_minimal(base_size = 12) +
    theme(
      axis.text.y = element_text(size = 6),
      panel.grid = element_blank(),
      plot.title = element_text(face = "bold"),
      axis.text.x = element_text(size = 10, face = "bold")
    ) +
    labs(
      title = paste(category_name, "- Heatmap"),
      subtitle = paste(length(proteins), "unique proteins"),
      x = "Timepoint",
      y = paste(length(proteins), "Proteins")
    )
  
  return(p)
}

# ============================================
# CHECK AND FIX YOUR DATA FIRST
# ============================================

cat("=== Checking for Duplicates in Categories ===\n\n")

# Check Category 2
if(exists("cat2_traj")) {
  cat("Category 2:\n")
  cat("  Total rows:", nrow(cat2_traj), "\n")
  cat("  Unique proteins:", length(unique(cat2_traj$protein)), "\n")
  
  if(any(duplicated(cat2_traj$protein))) {
    cat("  WARNING: Found", sum(duplicated(cat2_traj$protein)), "duplicates!\n")
    cat("  Removing duplicates...\n")
    
    # Keep only first occurrence of each protein
    cat2_traj <- cat2_traj %>%
      distinct(protein, .keep_all = TRUE)
    
    cat("  After cleaning:", nrow(cat2_traj), "proteins\n")
  }
  cat("\n")
}

# Check Category 4
if(exists("cat4_traj")) {
  cat("Category 4:\n")
  cat("  Total rows:", nrow(cat4_traj), "\n")
  cat("  Unique proteins:", length(unique(cat4_traj$protein)), "\n")
  
  if(any(duplicated(cat4_traj$protein))) {
    cat("  WARNING: Found", sum(duplicated(cat4_traj$protein)), "duplicates!\n")
    cat("  Removing duplicates...\n")
    
    # Keep only first occurrence of each protein (best FDR)
    cat4_traj <- cat4_traj %>%
      arrange(fdr_trend) %>%  # Sort by FDR first
      distinct(protein, .keep_all = TRUE)
    
    cat("  After cleaning:", nrow(cat4_traj), "proteins\n")
  }
  cat("\n")
}

# ============================================
# NOW CREATE HEATMAPS
# ============================================

cat("=== Creating Heatmaps ===\n\n")

# Heatmap for Category 2
if(exists("cat2_traj") && nrow(cat2_traj) > 0) {
  p_heatmap_cat2 <- create_category_heatmap(
    cat2_traj, 
    means_by_timepoint,
    "Category 2: Protein ↓ as Burnout ↓",
    "#FF6F00"
  )
  
  if(!is.null(p_heatmap_cat2)) {
    print(p_heatmap_cat2)
    ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Olink_Cat4_Cat2/Cat2_Heatmap_Proteins.png", 
           p_heatmap_cat2, 
           width = 8, 
           height = max(6, nrow(cat2_traj) * 0.15), 
           dpi = 300,
           limitsize = FALSE)
    ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Olink_Cat4_Cat2/Cat2_Heatmap_Proteins.pdf", 
           p_heatmap_cat2, 
           width = 5, 
           height = max(6, nrow(cat2_traj) * 0.15), 
           dpi = 300,
           limitsize = FALSE)
    cat("✓ Cat2_Heatmap_Proteins.png\n\n")
  }
}

# Heatmap for Category 4
if(exists("cat4_traj") && nrow(cat4_traj) > 0) {
  p_heatmap_cat4 <- create_category_heatmap(
    cat4_traj, 
    means_by_timepoint,
    "Category 4: Protein ↑ as Burnout ↓",
    "#388E3C"
  )
  
  if(!is.null(p_heatmap_cat4)) {
    print(p_heatmap_cat4)
    ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Olink_Cat4_Cat2/Cat4_Heatmap_Proteins.png", 
           p_heatmap_cat4, 
           width = 8, 
           height = max(6, nrow(cat4_traj) * 0.15), 
           dpi = 300,
           limitsize = FALSE)
     ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Olink_Cat4_Cat2/Cat4_Heatmap_Proteins.pdf", 
           p_heatmap_cat4, 
           width = 5, 
           height = max(6, nrow(cat4_traj) * 0.15), 
           dpi = 300,
           limitsize = FALSE)
    cat("✓ Cat4_Heatmap_Proteins.png\n\n")
  }
}

```

# ============================================
# SUMMARY TABLE
# ============================================


```{r}
# ============================================
# SUMMARY TABLE
# ============================================

summary_stats_table <- data.frame(
  Category = c(
    "Category 1: Protein ↑ as Burnout ↑",
    "Category 2: Protein ↓ as Burnout ↓",
    "Category 3: Protein ↓ as Burnout ↑",
    "Category 4: Protein ↑ as Burnout ↓"
  ),
  N_Proteins = c(nrow(cat1_traj), nrow(cat2_traj), nrow(cat3_traj), nrow(cat4_traj)),
  Mean_Slope = c(
    ifelse(nrow(cat1_traj) > 0, mean(cat1_traj$protein_slope), NA),
    ifelse(nrow(cat2_traj) > 0, mean(cat2_traj$protein_slope), NA),
    ifelse(nrow(cat3_traj) > 0, mean(cat3_traj$protein_slope), NA),
    ifelse(nrow(cat4_traj) > 0, mean(cat4_traj$protein_slope), NA)
  ),
  Mean_Correlation = c(
    ifelse(nrow(cat1_traj) > 0, mean(cat1_traj$correlation_with_burnout), NA),
    ifelse(nrow(cat2_traj) > 0, mean(cat2_traj$correlation_with_burnout), NA),
    ifelse(nrow(cat3_traj) > 0, mean(cat3_traj$correlation_with_burnout), NA),
    ifelse(nrow(cat4_traj) > 0, mean(cat4_traj$correlation_with_burnout), NA)
  )
)

cat("\n=== SUMMARY STATISTICS ===\n")
print(summary_stats_table)
write.csv(summary_stats_table, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Summary_Statistics_Trajectories.csv", row.names = FALSE)


cat("\n=== SUMMARY STATISTICS ===\n")
print(summary_stats_table)
write.csv(summary_stats_table, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Summary_Statistics_Trajectories.csv", row.names = FALSE)

# ============================================
# FINAL REPORT
# ============================================

cat("\n\n========================================\n")
cat("TRAJECTORY-BASED 4-CATEGORY ANALYSIS COMPLETE\n")
cat("========================================\n\n")

cat("RESULTS:\n")
cat("  Category 1:", nrow(cat1_traj), "proteins\n")
cat("  Category 2:", nrow(cat2_traj), "proteins\n")
cat("  Category 3:", nrow(cat3_traj), "proteins\n")
cat("  Category 4:", nrow(cat4_traj), "proteins\n\n")

cat("FILES SAVED:\n")
cat("  ✓ All CSV files\n")
cat("  ✓ Individual trajectory plots\n")
cat("  ✓ Summary visualizations\n\n")

cat("========================================\n\n")

```
###Lmer Positive and negative models 
##graphics
```{r}
full_dataT<- read.csv("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/fullData.csv")
rownames(full_dataT)<-full_dataT$X


dex<- read.csv("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/MetaData/burnoutEx.csv")
rownames(dex)<-dex$X


colnames(full_dataT)
library(tidyr)
library(dplyr)

# Example: wide data››¢
# Columns: ID, time, age, sex, protein1, protein2, ..., protein5000
##not used
df_long <- full_dataT %>%
  pivot_longer(
    cols = 9:5435,
    names_to = "protein",
    values_to = "value"
  )
#####
listNeg<- read.csv("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/NegList.csv")
rownames(listNeg)<-listNeg$X


listNeg$cor<-ifelse(listNeg$estimate<0, -1, 1)


listPos<- read.csv("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/PosList.csv")
listPos$cor<-ifelse(listPos$estimate<0,-1,1)

data<- rbind(listNeg,listPos)
# Inspect
head(df_long)

##now from full data select only expression for those neg and pos 
library(dplyr)

met<-full_dataT[1:8]
```


```{r}
ExpNeg <- dex%>%
  select(any_of(listNeg$protein))

library(tibble)
library(tidyr)

dtN <- ExpNeg %>%
  rownames_to_column(var = "rowname") %>%
  separate(rowname, into = c("ID", "timepoint"), sep = "-")
rownames(dtN)<-rownames(ExpNeg)

dtNmet<- merge(met,dtN, by="row.names")
dtNmet<-dtNmet %>% filter(Time %in% c("C1", "C2", "C3", "C4", "C5", "C6"))
write.csv(dtNmet, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/dtNmetRemoveNA.csv")

ExpPos <-dex %>%
  select(any_of(listPos$protein))

dtP <- ExpPos %>%
  rownames_to_column(var = "rowname") %>%
  separate(rowname, into = c("ID", "timepoint"), sep = "-")
rownames(dtP)<-rownames(ExpPos)

dtPmet<- merge(met,dtP, by="row.names")
colnames(dtPmet)
unique(dtPmet$Time)

dtPmet<-dtPmet %>% filter(Time %in% c("C1", "C2", "C3", "C4", "C5", "C6"))

write.csv(dtPmet, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/dtPmetRemoveNA.csv")
```

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

dx<-dtPmet # used remove NA

##second Neg

dx<-dtNmet
# ============================================
# METHOD 1: FACETED PLOT FOR MANY PROTEINS
# ============================================

# Select first 400 proteins (columns 4 to 403 assuming first 3 are ID, Timepoint, Burnout)
# Adjust column indices based on your actual data structure

# Get protein column names (assuming proteins start at column 6)
protein_cols <- colnames(dx)[12:408]  # First 400 proteins for positive

protein_cols <- colnames(dx)[12:50]  # First 400 proteins for negative


# Reshape to long format
protein_long <- dx %>%
  select(ID, Time, all_of(protein_cols)) %>%
  pivot_longer(
    cols = all_of(protein_cols),
    names_to = "Protein",
    values_to = "Expression"
  )

cat("Data reshaped:\n")
cat("  Total rows:", nrow(protein_long), "\n")
cat("  Proteins:", length(unique(protein_long$Protein)), "\n\n")

# Create faceted plot (small multiples)
# For 400 proteins, this creates a grid of small plots

p_faceted <- ggplot(protein_long, 
                    aes(x = Time, y = Expression, group = ID)) +
  geom_line(alpha = 0.2, color = "gray70", linewidth = 0.5) +
  stat_summary(aes(group = 1), fun = mean, geom = "line", 
               color = "red", linewidth = 0.8) +
  facet_wrap(~ Protein, scales = "free_y", ncol = 20) +  # 20 columns
  theme_minimal(base_size = 15) +
  theme(
    strip.text = element_text(size = 16),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16)
  ) + theme_bw()+
  labs(
    title = "Trajectories for 39 Proteins (T1-T6)",  
    subtitle = "Red line = mean trajectory, gray lines = individual subjects",
    x = "Timepoint",
    y = "Protein Expression"
  ) #400 Proteins (T1-T6)

print(p_faceted)



ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Faceted_39_Proteins_Neg.png", p_faceted, 
       width = 40, height = 10, dpi = 300, limitsize = FALSE) # width = 45, height = 50, dpi = 300, limitsize = FALSE)
ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Faceted_39_Proteins_Neg.pdf", p_faceted, 
       width = 40, height = 10, dpi = 300, limitsize = FALSE) #width = 55, height = 65, dpi = 300, limitsize = FALSE

cat("Saved: Faceted_400_Proteins.png (large file!)\n\n")
cat("Saved: Faceted_39_Proteins.png (large file!)\n\n")

```
#pvalue line plot
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

# ============================================
# STEP 1: PREPARE DATA
# ============================================

dx<-dtPmet # used remove NA

protein_cols <- colnames(dx)[12:408]

##second Neg

dx<-dtNmet # used remove NA
# ============================================
# METHOD 1: FACETED PLOT FOR MANY PROTEINS
# ============================================

# Select first 400 proteins (columns 4 to 403 assuming first 3 are ID, Timepoint, Burnout)
# Adjust column indices based on your actual data structure

protein_cols <- colnames(dx)[12:50]  # First 400 proteins for negative
```


```{r}
cat("=== Preparing Data ===\n")
cat("Proteins to analyze:", length(protein_cols), "\n\n")

# Handle duplicates by averaging
dx_clean <- dx %>%
  group_by(ID, Time) %>%
  summarise(across(all_of(protein_cols), ~mean(.x, na.rm = TRUE)), 
            .groups = 'drop')

# Reshape to long format
protein_long <- dx_clean %>%
  select(ID, Time, all_of(protein_cols)) %>%
  pivot_longer(
    cols = all_of(protein_cols),
    names_to = "Protein",
    values_to = "Expression"
  )

cat("Data reshaped:\n")
cat("  Total rows:", nrow(protein_long), "\n")
cat("  Proteins:", length(unique(protein_long$Protein)), "\n")
cat("  Timepoints:", paste(unique(protein_long$Time), collapse = ", "), "\n\n")

# ============================================
# STEP 2: RUN WILCOXON TESTS
# ============================================

cat("=== Running Wilcoxon Tests ===\n")

# Determine reference timepoint
ref_time <- sort(unique(dx_clean$Time))[1]
cat("Reference timepoint:", ref_time, "\n\n")

# Function to test one protein
wilcox_one_protein <- function(protein_name, data, ref_time) {
  
  results <- data.frame()
  other_times <- setdiff(unique(data$Time), ref_time)
  
  for(comp_time in other_times) {
    
    # Get paired data
    ref_data <- data %>% 
      filter(Time == ref_time) %>%
      select(ID, Expression)
    
    comp_data <- data %>%
      filter(Time == comp_time) %>%
      select(ID, Expression)
    
    # Merge
    merged <- inner_join(ref_data, comp_data, by = "ID", 
                        suffix = c("_ref", "_comp"))
    
    # Clean
    merged_clean <- merged %>%
      filter(!is.na(Expression_ref), !is.na(Expression_comp))
    
    if(nrow(merged_clean) >= 5) {
      
      test <- tryCatch({
        wilcox.test(merged_clean$Expression_ref,
                   merged_clean$Expression_comp,
                   paired = TRUE,
                   exact = FALSE)
      }, error = function(e) NULL)
      
      if(!is.null(test)) {
        results <- rbind(results, data.frame(
          Protein = protein_name,
          Time = comp_time,
          P_value = test$p.value,
          N_pairs = nrow(merged_clean)
        ))
      }
    }
  }
  
  return(results)
}

# Run tests for all proteins
cat("Testing", length(protein_cols), "proteins...\n")

all_wilcox <- data.frame()
pb <- txtProgressBar(min = 0, max = length(protein_cols), style = 3)

for(i in 1:length(protein_cols)) {
  prot_data <- protein_long %>% filter(Protein == protein_cols[i])
  result <- wilcox_one_protein(protein_cols[i], prot_data, ref_time)
  if(nrow(result) > 0) {
    all_wilcox <- rbind(all_wilcox, result)
  }
  setTxtProgressBar(pb, i)
}

close(pb)

cat("\n\nWilcoxon tests completed!\n")

# FDR correction
all_wilcox <- all_wilcox %>%
  group_by(Time) %>%
  mutate(FDR = p.adjust(P_value, method = "BH")) %>%
  ungroup()

# Add significance labels
all_wilcox <- all_wilcox %>%
  mutate(
    Sig_label = case_when(
      FDR < 0.05 ~ "***",
      FDR < 0.1 ~ "**",
      FDR < 0.2 ~ "*",
      TRUE ~ ""
    ),
    Significant = FDR < 0.2
  )

cat("Significant results (FDR < 0.05):", sum(all_wilcox$FDR < 0.2), "\n\n")
```


```{r}
# Save results
write.csv(all_wilcox, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/POS/Wilcoxon_Results_All_Proteins.csv", row.names = FALSE)

write.csv(all_wilcox, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/NEG/Wilcoxon_Results_All_Proteins.csv", row.names = FALSE)
```

# ============================================
# STEP 3: PREPARE PLOT ANNOTATIONS
# ============================================

```{r}
# ============================================
# STEP 3: PREPARE PLOT ANNOTATIONS
# ============================================

# Calculate mean expression for positioning annotations
mean_expr <- protein_long %>%
  group_by(Protein, Time) %>%
  summarise(Mean_Expression = mean(Expression, na.rm = TRUE),
            Max_Expression = max(Expression, na.rm = TRUE),
            .groups = 'drop')

# Merge with Wilcoxon results
plot_annotations <- all_wilcox %>%
  left_join(mean_expr, by = c("Protein", "Time")) %>%
  filter(Significant) %>%  # Only show significant
  mutate(
    # Position slightly above the max value
    Y_position = Max_Expression * 1.05
  )

cat("Annotations prepared for", nrow(plot_annotations), "significant comparisons\n\n")

# ============================================
# STEP 4: CREATE PLOT WITH SIGNIFICANCE
# ============================================

cat("=== Creating Faceted Plot with Wilcoxon Results ===\n\n")

p_faceted <- ggplot(protein_long, 
                    aes(x = Time, y = Expression, group = ID)) +
  # Individual trajectories
  geom_line(alpha = 0.2, color = "gray70", linewidth = 0.3) +
  
  # Mean trajectory
  stat_summary(aes(group = 1), fun = mean, geom = "line", 
               color = "red", linewidth = 0.8) +
  stat_summary(aes(group = 1), fun = mean, geom = "point", 
               color = "red", size = 1.5) +
  
  # Add significance annotations
  geom_text(data = plot_annotations,
            aes(x = Time, y = Y_position, label = Sig_label, group = NULL),
            color = "blue", size = 3, fontface = "bold",
            inherit.aes = FALSE) +
  
  # Facets
  facet_wrap(~ Protein, scales = "free_y", ncol = 20) +
  
  # Theme
  theme_minimal(base_size = 10) +
  theme(
    strip.text = element_text(size = 6, face = "bold"),
    axis.text = element_text(size = 8),
    axis.title = element_text(size = 12, face = "bold"),
    panel.spacing = unit(0.1, "lines")
  ) +
  
  # Labels
  labs(
    title = "Protein Trajectories with Wilcoxon Test Results",
    subtitle = paste0("Red = mean trajectory | Blue asterisks = significant vs ", ref_time, 
                     " (* FDR<0.2, ** FDR<0.1, *** FDR<0.05)"),
    x = "Timepoint",
    y = "Protein Expression"
  )

print(p_faceted)
```


```{r}
ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/POS/Faceted_Proteins_with_Wilcoxon.png", p_faceted, 
       width = 40, height = 40, dpi = 300, limitsize = FALSE)
ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/POS/Faceted_Proteins_with_Wilcoxon.pdf", p_faceted, 
       width = 40, height = 40, limitsize = FALSE)

cat("✓ Plot saved: Faceted_Proteins_with_Wilcoxon.png\n\n")


ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/NEG/Faceted_Proteins_with_Wilcoxon.png", p_faceted, 
       width = 40, height = 40, dpi = 300, limitsize = FALSE)
ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/NEG/Faceted_Proteins_with_Wilcoxon.pdf", p_faceted, 
       width = 40, height = 40, limitsize = FALSE)

cat("✓ Plot saved: Faceted_Proteins_with_Wilcoxon.png\n\n")
```

# ============================================
# STEP 5: ALTERNATIVE - COLOR-CODED PLOT
# ============================================
```{r}
# ============================================
# STEP 5: ALTERNATIVE - COLOR-CODED PLOT
# ============================================

cat("=== Creating Alternative Version (Color-coded significance) ===\n\n")

# Add significance info to main data
protein_long_sig <- protein_long %>%
  left_join(
    all_wilcox %>% select(Protein, Time, Significant),
    by = c("Protein", "Time")
  ) %>%
  mutate(Significant = replace_na(Significant, FALSE))

# Calculate mean with significance
mean_data_sig <- protein_long_sig %>%
  group_by(Protein, Time, Significant) %>%
  summarise(Mean_Expression = mean(Expression, na.rm = TRUE),
            .groups = 'drop')

p_faceted_color <- ggplot() +
  # Individual trajectories
  geom_line(data = protein_long,
            aes(x = Time, y = Expression, group = ID),
            alpha = 0.2, color = "gray70", linewidth = 0.3) +
  
  # Mean trajectory - colored by significance
  geom_line(data = mean_data_sig,
            aes(x = Time, y = Mean_Expression, 
                group = Protein, color = Significant),
            linewidth = 1) +
  geom_point(data = mean_data_sig,
             aes(x = Time, y = Mean_Expression, 
                 color = Significant, fill = Significant),
             size = 2, shape = 21, stroke = 1) +
  
  scale_color_manual(
    values = c("FALSE" = "red", "TRUE" = "blue"),
    labels = c("Not significant", paste0("Significant vs ", ref_time))
  ) +
  scale_fill_manual(
    values = c("FALSE" = "red", "TRUE" = "blue"),
    labels = c("Not significant", paste0("Significant vs ", ref_time))
  ) +
  
  facet_wrap(~ Protein, scales = "free_y", ncol = 20) +
  
  theme_minimal(base_size = 10) +
  theme(
    strip.text = element_text(size = 6, face = "bold"),
    axis.text = element_text(size = 8),
    legend.position = "bottom",
    legend.text = element_text(size = 10),
    panel.spacing = unit(0.1, "lines")
  ) +
  
  labs(
    title = "Protein Trajectories with Wilcoxon Test Results",
    subtitle = paste0("Blue points = significantly different from ", ref_time, " (FDR < 0.05)"),
    x = "Timepoint",
    y = "Protein Expression",
    color = "Wilcoxon Test",
    fill = "Wilcoxon Test"
  )

print(p_faceted_color)
```


```{r}
ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/POS/Faceted_Proteins_ColorCoded.png", p_faceted_color, 
       width = 40, height = 40, dpi = 300, limitsize = FALSE)
ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/POS/Faceted_Proteins_ColorCoded.pdf", p_faceted_color, 
       width = 40, height = 40, dpi = 300, limitsize = FALSE)


ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/NEG/Faceted_Proteins_ColorCoded.png", p_faceted_color, 
       width = 40, height = 40, dpi = 300, limitsize = FALSE)
ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/NEG/Faceted_Proteins_ColorCoded.pdf", p_faceted_color, 
       width = 40, height = 40, dpi = 300, limitsize = FALSE)
cat("✓ Alternative plot saved: Faceted_Proteins_ColorCoded.png\n\n")
```

# ============================================
# STEP 6: SUMMARY PLOT - HEATMAP OF SIGNIFICANCE
# ============================================
```{r}
# ============================================
# STEP 6: SUMMARY PLOT - HEATMAP OF SIGNIFICANCE
# ============================================

cat("=== Creating Summary Heatmap ===\n\n")

# Create significance matrix
sig_matrix <- all_wilcox %>%
  mutate(Sig_numeric = -log10(FDR)) %>%
  select(Protein, Time, Sig_numeric) %>%
  pivot_wider(names_from = Time, values_from = Sig_numeric, values_fill = 0)

sig_matrix_long <- sig_matrix %>%
  pivot_longer(cols = -Protein, names_to = "Time", values_to = "neg_log_FDR")

p_heatmap <- ggplot(sig_matrix_long, 
                    aes(x = Time, y = Protein, fill = neg_log_FDR)) +
  geom_tile(color = "white", linewidth = 0.1) +
  scale_fill_gradient2(
    low = "white", mid = "yellow", high = "red",
    midpoint = 1.3,  # -log10(0.05)
    name = "-log10(FDR)",
    breaks = c(0, 1.3, 2, 3),
    labels = c("NS", "0.2", "0.1", "0.05")
  ) +
  theme_minimal(base_size = 10) +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid = element_blank()
  ) +
  labs(
    title = "Heatmap: Wilcoxon Test Significance Across All Proteins",
    subtitle = paste0("Comparing each timepoint to ", ref_time),
    x = "Timepoint",
    y = paste(length(protein_cols), "Proteins")
  )

print(p_heatmap)

ggsave("Heatmap_Wilcoxon_Significance.png", p_heatmap, 
       width = 8, height = 20, dpi = 300)

cat("✓ Heatmap saved: Heatmap_Wilcoxon_Significance.png\n\n")
```


```{r}
# ============================================
# STEP 7: SUMMARY STATISTICS
# ============================================

cat("=== Summary Statistics ===\n\n")

summary_stats <- all_wilcox %>%
  group_by(Time) %>%
  summarise(
    N_proteins = n(),
    N_sig_p05 = sum(P_value < 0.05),
    N_sig_FDR1 = sum(FDR < 0.1),
    N_sig_FDR2 = sum(FDR < 0.2),
    Median_pvalue = median(P_value),
    .groups = 'drop'
  )

cat("Significance summary by timepoint:\n")
print(summary_stats)

write.csv(summary_stats, "Wilcoxon_Summary_by_Timepoint.csv", row.names = FALSE)

# ============================================
# FINAL REPORT
# ============================================

cat("\n========================================\n")
cat("ANALYSIS COMPLETE\n")
cat("========================================\n\n")

cat("PROTEINS ANALYZED:", length(protein_cols), "\n")
cat("REFERENCE TIMEPOINT:", ref_time, "\n")
cat("COMPARISONS:", paste(setdiff(unique(protein_long$Time), ref_time), collapse = ", "), "\n\n")

cat("SIGNIFICANT RESULTS (FDR < 0.2):\n")
for(i in 1:nrow(summary_stats)) {
  cat("  ", summary_stats$Time[i], ": ", 
      summary_stats$N_sig_FDR05[i], " proteins\n", sep = "")
}

cat("\nFILES CREATED:\n")
cat("  ✓ Faceted_Proteins_with_Wilcoxon.png (asterisks)\n")
cat("  ✓ Faceted_Proteins_ColorCoded.png (color-coded)\n")
cat("  ✓ Heatmap_Wilcoxon_Significance.png (overview)\n")
cat("  ✓ Wilcoxon_Results_All_Proteins.csv (detailed results)\n")
cat("  ✓ Wilcoxon_Summary_by_Timepoint.csv (summary)\n\n")

cat("========================================\n\n")
```
# ============================================
# PLOT ONLY PROTEINS WITH SIGNIFICANT CHANGES
# ============================================
```{r}
# ============================================
# PLOT ONLY PROTEINS WITH SIGNIFICANT CHANGES
# ============================================

cat("=== Creating Plot for Significant Proteins Only ===\n\n")

# Get proteins with at least one significant change
sig_proteins <- unique(all_wilcox$Protein[all_wilcox$FDR < 0.2])

cat("Proteins with significant changes:", length(sig_proteins), "\n\n")

if(length(sig_proteins) > 0) {
  
  # Filter data
  protein_long_sig_only <- protein_long %>%
    filter(Protein %in% sig_proteins)
  
  plot_annotations_sig <- plot_annotations %>%
    filter(Protein %in% sig_proteins)
  
  # Create plot
  p_sig_only <- ggplot(protein_long_sig_only, 
                       aes(x = Time, y = Expression, group = ID)) +
    geom_line(alpha = 0.2, color = "gray70", linewidth = 0.3) +
    stat_summary(aes(group = 1), fun = mean, geom = "line", 
                 color = "red", linewidth = 1) +
    stat_summary(aes(group = 1), fun = mean, geom = "point", 
                 color = "red", size = 2) +
    geom_text(data = plot_annotations_sig,
              aes(x = Time, y = Y_position, label = Sig_label, group = NULL),
              color = "blue", size = 4, fontface = "bold",
              inherit.aes = FALSE) +
    facet_wrap(~ Protein, scales = "free_y", ncol = 10) +
    theme_minimal(base_size = 12) +
    theme(
      strip.text = element_text(size = 10, face = "bold"),
      axis.text = element_text(size = 10)
    ) +
    labs(
      title = paste("Significant Proteins Only (n =", length(sig_proteins), ")"),
      subtitle = paste0("Proteins with FDR < 0.2 vs ", ref_time),
      x = "Timepoint",
      y = "Expression"
    )
  
  print(p_sig_only)
  
  ggsave("Significant_Proteins_Only.png", p_sig_only, 
         width = 20, height = ceiling(length(sig_proteins)/10) * 3, 
         dpi = 300, limitsize = FALSE)
    ggsave("Significant_Proteins_Only.pdf", p_sig_only, 
         width = 20, height = ceiling(length(sig_proteins)/10) * 3, 
         dpi = 300, limitsize = FALSE)
  
  cat("✓ Significant proteins plot saved\n")
}
```
##baseline T1
# ============================================
# STEP 1: PREPARE DATA
# ============================================
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)


# ============================================
# STEP 1: PREPARE DATA
# ============================================

cat("=== Data Preparation ===\n\n")

# Handle duplicates
dx_clean <- dx %>%
  group_by(ID, Time) %>%
  summarise(across(all_of(protein_cols), ~mean(.x, na.rm = TRUE)), 
            .groups = 'drop')

# Check available timepoints
available_times <- sort(unique(dx_clean$Time))
cat("Available timepoints:", paste(available_times, collapse = ", "), "\n")

# EXPLICITLY SET C1 AS BASELINE
ref_time <- "C1"

# If C1 doesn't exist, find the correct name
if(!ref_time %in% available_times) {
  cat("\nWARNING: 'C1' not found!\n")
  cat("Looking for alternative baseline names...\n")
  
  # Try common alternatives
  if("T1" %in% available_times) {
    ref_time <- "T1"
    cat("Using T1 as baseline\n")
  } else if("1" %in% available_times) {
    ref_time <- "1"
    cat("Using 1 as baseline\n")
  } else {
    ref_time <- available_times[1]
    cat("Using first timepoint (", ref_time, ") as baseline\n")
  }
}

cat("\n*** BASELINE SET TO:", ref_time, "***\n\n")

# Reshape to long
protein_long <- dx_clean %>%
  select(ID, Time, all_of(protein_cols)) %>%
  pivot_longer(
    cols = all_of(protein_cols),
    names_to = "Protein",
    values_to = "Expression"
  )

# ============================================
# STEP 2: WILCOXON TESTS - C1 AS REFERENCE
# ============================================

cat("=== Running Wilcoxon Tests ===\n")
cat("Reference (baseline):", ref_time, "\n")

other_times <- setdiff(unique(dx_clean$Time), ref_time)
cat("Comparing to:", paste(other_times, collapse = ", "), "\n\n")

# Function to test one protein against C1
wilcox_vs_baseline <- function(protein_name, data, baseline) {
  
  results <- data.frame()
  comparison_times <- setdiff(unique(data$Time), baseline)
  
  # Get baseline data once
  baseline_data <- data %>% 
    filter(Time == baseline) %>%
    select(ID, Expression) %>%
    rename(Expression_baseline = Expression)
  
  for(comp_time in comparison_times) {
    
    # Get comparison timepoint data
    comp_data <- data %>%
      filter(Time == comp_time) %>%
      select(ID, Expression) %>%
      rename(Expression_comp = Expression)
    
    # Merge on ID (paired test)
    merged <- inner_join(baseline_data, comp_data, by = "ID") %>%
      filter(!is.na(Expression_baseline), !is.na(Expression_comp))
    
    if(nrow(merged) >= 5) {
      
      test <- tryCatch({
        wilcox.test(merged$Expression_baseline,
                   merged$Expression_comp,
                   paired = TRUE,
                   exact = FALSE)
      }, error = function(e) NULL)
      
      if(!is.null(test)) {
        
        # Calculate effect (median difference)
        median_baseline <- median(merged$Expression_baseline, na.rm = TRUE)
        median_comp <- median(merged$Expression_comp, na.rm = TRUE)
        median_diff <- median_comp - median_baseline
        
        results <- rbind(results, data.frame(
          Protein = protein_name,
          Baseline = baseline,
          Comparison = comp_time,
          Comparison_label = paste0(baseline, "_vs_", comp_time),
          N_pairs = nrow(merged),
          Median_baseline = median_baseline,
          Median_comparison = median_comp,
          Median_difference = median_diff,
          P_value = test$p.value,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
  
  return(results)
}

# Run for all proteins
cat("Testing", length(protein_cols), "proteins against baseline", ref_time, "...\n")

all_wilcox <- data.frame()
pb <- txtProgressBar(min = 0, max = length(protein_cols), style = 3)

for(i in 1:length(protein_cols)) {
  prot_data <- protein_long %>% filter(Protein == protein_cols[i])
  result <- wilcox_vs_baseline(protein_cols[i], prot_data, ref_time)
  if(!is.null(result) && nrow(result) > 0) {
    all_wilcox <- rbind(all_wilcox, result)
  }
  setTxtProgressBar(pb, i)
}

close(pb)

cat("\n\nTests completed!\n")
cat("Total comparisons:", nrow(all_wilcox), "\n\n")

# FDR correction
all_wilcox <- all_wilcox %>%
  group_by(Comparison) %>%
  mutate(FDR = p.adjust(P_value, method = "BH")) %>%
  ungroup()

# Global FDR
all_wilcox$FDR_global <- p.adjust(all_wilcox$P_value, method = "BH")

# Significance labels
all_wilcox <- all_wilcox %>%
  mutate(
    Sig_label = case_when(
      FDR < 0.05 ~ "***",
      FDR < 0.1 ~ "**",
      FDR < 0.2 ~ "*",
      TRUE ~ ""
    ),
    Significant = FDR < 0.2,
    Direction = case_when(
      Median_difference > 0 ~ "Increased vs baseline",
      Median_difference < 0 ~ "Decreased vs baseline",
      TRUE ~ "No change"
    )
  )

# Summary
cat("=== SUMMARY (vs baseline", ref_time, ") ===\n\n")

summary_by_time <- all_wilcox %>%
  group_by(Comparison) %>%
  summarise(
    N_proteins = n(),
    N_sig = sum(Significant),
    N_increased = sum(Significant & Median_difference > 0),
    N_decreased = sum(Significant & Median_difference < 0),
    .groups = 'drop'
  )

print(summary_by_time)
cat("\n")

# Save results
write.csv(all_wilcox, 
          paste0("Wilcoxon_vs_", ref_time, "_All_Results.csv"), 
          row.names = FALSE)

write.csv(summary_by_time, 
          paste0("Wilcoxon_vs_", ref_time, "_Summary.csv"), 
          row.names = FALSE)

# ============================================
# STEP 3: CREATE PLOT WITH CLEAR BASELINE LABEL
# ============================================

cat("=== Creating Plot ===\n\n")

# Prepare annotations
mean_expr <- protein_long %>%
  group_by(Protein, Time) %>%
  summarise(
    Mean_Expression = mean(Expression, na.rm = TRUE),
    Max_Expression = max(Expression, na.rm = TRUE),
    .groups = 'drop'
  )

plot_annotations <- all_wilcox %>%
  filter(Significant) %>%
  left_join(mean_expr, by = c("Protein", "Comparison" = "Time")) %>%
  mutate(Y_position = Max_Expression * 1.05)

# Create plot
p_faceted <- ggplot(protein_long, 
                    aes(x = Time, y = Expression, group = ID)) +
  
  # Highlight baseline timepoint
  annotate("rect", 
           xmin = as.numeric(factor(ref_time, levels = sort(unique(protein_long$Time)))) - 0.4,
           xmax = as.numeric(factor(ref_time, levels = sort(unique(protein_long$Time)))) + 0.4,
           ymin = -Inf, ymax = Inf,
           fill = "yellow", alpha = 0.1) +
  
  # Individual trajectories
  geom_line(alpha = 0.2, color = "gray70", linewidth = 0.3) +
  
  # Mean trajectory
  stat_summary(aes(group = 1), fun = mean, geom = "line", 
               color = "red", linewidth = 0.8) +
  stat_summary(aes(group = 1), fun = mean, geom = "point", 
               color = "red", size = 1.5) +
  
  # Significance annotations
  geom_text(data = plot_annotations,
            aes(x = Comparison, y = Y_position, label = Sig_label, group = NULL),
            color = "blue", size = 3, fontface = "bold",
            inherit.aes = FALSE) +
  
  facet_wrap(~ Protein, scales = "free_y", ncol = 20) +
  
  theme_minimal(base_size = 10) +
  theme(
    strip.text = element_text(size = 6, face = "bold"),
    axis.text = element_text(size = 8),
    panel.spacing = unit(0.1, "lines")
  ) +
  
  labs(
    title = paste0("Protein Trajectories with Significance vs Baseline (", ref_time, ")"),
    subtitle = paste0("Yellow background = baseline (", ref_time, 
                     ") | Blue asterisks = significant vs baseline (* FDR<0.2, ** FDR<0.1, *** FDR<0.05)"),
    x = "Timepoint",
    y = "Protein Expression"
  ) + theme_bw()

print(p_faceted)

ggsave(paste0("Proteins_vs_", ref_time, "_baseline.png"), p_faceted, 
       width = 30, height = 10, dpi = 300, limitsize = FALSE) #height = 40
ggsave(paste0("Proteins_vs_", ref_time, "_baseline.pdf"), p_faceted, 
       width = 30, height = 10, dpi = 300, limitsize = FALSE) #height = 40

cat("✓ Plot saved with", ref_time, "as baseline\n\n")

# ============================================
# VERIFICATION TABLE
# ============================================

cat("=== VERIFICATION ===\n\n")

cat("BASELINE TIMEPOINT:", ref_time, "\n\n")

cat("EXAMPLE RESULTS (first protein):\n")
example <- all_wilcox %>%
  filter(Protein == protein_cols[1]) %>%
  select(Protein, Baseline, Comparison, N_pairs, Median_difference, P_value, FDR, Sig_label)

print(example)
cat("\n")

cat("INTERPRETATION:\n")
cat("  - 'Baseline' column shows", ref_time, "for all tests\n")
cat("  - 'Comparison' shows which timepoint is being compared to", ref_time, "\n")
cat("  - Asterisks (*) appear only when that timepoint differs from", ref_time, "\n\n")

cat("========================================\n\n")
```


#PValue calculation
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)

library(dplyr)
library(tidyr)
library(ggplot2)

# ============================================
# STEP 1: DIAGNOSE AND FIX DUPLICATES
# ============================================

cat("=== Diagnosing Data Structure ===\n\n")

# Get protein columns
protein_cols <- colnames(dx)[12:408]

# Check for duplicates in original data
duplicates_check <- dx %>%
  group_by(ID, Time) %>%
  summarise(n = n(), .groups = "drop") %>%
  filter(n > 1)

if(nrow(duplicates_check) > 0) {
  cat("WARNING: Found", nrow(duplicates_check), "duplicate ID-Time combinations\n")
  cat("First few duplicates:\n")
  print(head(duplicates_check))
  cat("\nAveraging duplicates...\n\n")
  
  # Average duplicates
  dx_clean <- dx %>%
    group_by(ID, Time) %>%
    summarise(
      across(all_of(protein_cols), ~mean(.x, na.rm = TRUE)),
      .groups = 'drop'
    )
} else {
  cat("✓ No duplicates found\n\n")
  dx_clean <- dx %>%
    select(ID, Time, all_of(protein_cols))
}

cat("Clean data dimensions:", dim(dx_clean), "\n")
cat("Unique ID-Time combinations:", nrow(dx_clean), "\n")
cat("Timepoints:", paste(unique(dx_clean$Time), collapse = ", "), "\n\n")

# ============================================
# STEP 2: RESHAPE TO LONG FORMAT
# ============================================

protein_long <- dx_clean %>%
  select(ID, Time, all_of(protein_cols)) %>%
  pivot_longer(
    cols = all_of(protein_cols),
    names_to = "Protein",
    values_to = "Expression"
  )

cat("Long format created:\n")
cat("  Total rows:", nrow(protein_long), "\n")
cat("  Proteins:", length(unique(protein_long$Protein)), "\n")
cat("  Subjects:", length(unique(protein_long$ID)), "\n\n")

# ============================================
# STEP 3: WILCOXON TEST - FIXED VERSION
# ============================================

cat("=== Running Wilcoxon Signed-Rank Tests ===\n\n")

# Function to perform Wilcoxon test for one protein - FIXED
wilcox_test_protein <- function(protein_name, data_long) {
  
  # Get data for this protein
  protein_data <- data_long %>%
    filter(Protein == protein_name) %>%
    select(ID, Time, Expression) %>%
    # Remove NAs before pivoting
    filter(!is.na(Expression)) %>%
    # Average any remaining duplicates
    group_by(ID, Time) %>%
    summarise(Expression = mean(Expression, na.rm = TRUE), .groups = 'drop') %>%
    # Pivot to wide
    pivot_wider(
      names_from = Time, 
      values_from = Expression
    )
  
  # Check what columns we have
  available_cols <- colnames(protein_data)
  
  # Determine reference column name (C1 or T1 or 1, etc.)
  ref_col <- NULL
  if("C1" %in% available_cols) {
    ref_col <- "C1"
    comp_cols <- c("C2", "C3", "C4", "C5", "C6")
  } else if("T1" %in% available_cols) {
    ref_col <- "T1"
    comp_cols <- c("T2", "T3", "T4", "T5", "T6")
  } else if("1" %in% available_cols) {
    ref_col <- "1"
    comp_cols <- c("2", "3", "4", "5", "6")
  } else {
    # Take first column as reference
    time_cols <- setdiff(available_cols, "ID")
    if(length(time_cols) > 0) {
      ref_col <- time_cols[1]
      comp_cols <- time_cols[-1]
    } else {
      return(NULL)
    }
  }
  
  # Check if reference column exists
  if(is.null(ref_col) || !ref_col %in% available_cols) {
    return(NULL)
  }
  
  # Get reference values
  ref_values <- protein_data[[ref_col]]
  
  # Initialize results
  results <- data.frame(
    Protein = character(),
    Comparison = character(),
    N_pairs = numeric(),
    Median_Ref = numeric(),
    Median_Comp = numeric(),
    Median_Diff = numeric(),
    Statistic = numeric(),
    P_value = numeric(),
    stringsAsFactors = FALSE
  )
  
  # Test reference vs each other timepoint
  for(comp_col in comp_cols) {
    
    if(comp_col %in% available_cols) {
      
      comp_values <- protein_data[[comp_col]]
      
      # Remove pairs with NAs
      valid_pairs <- !is.na(ref_values) & !is.na(comp_values)
      ref_clean <- ref_values[valid_pairs]
      comp_clean <- comp_values[valid_pairs]
      
      if(length(ref_clean) >= 5) {  # Need at least 5 pairs
        
        tryCatch({
          # Perform paired Wilcoxon test
          test_result <- wilcox.test(ref_clean, comp_clean, 
                                     paired = TRUE, 
                                     exact = FALSE)
          
          # Calculate medians
          median_ref <- median(ref_clean, na.rm = TRUE)
          median_comp <- median(comp_clean, na.rm = TRUE)
          median_diff <- median_comp - median_ref
          
          # Store results
          results <- rbind(results, data.frame(
            Protein = protein_name,
            Comparison = paste0(ref_col, "_vs_", comp_col),
            N_pairs = length(ref_clean),
            Median_Ref = median_ref,
            Median_Comp = median_comp,
            Median_Diff = median_diff,
            Statistic = as.numeric(test_result$statistic),
            P_value = test_result$p.value,
            stringsAsFactors = FALSE
          ))
        }, error = function(e) {
          # Skip this comparison if test fails
          NULL
        })
      }
    }
  }
  
  return(results)
}

# Run for all proteins with progress bar
cat("Testing", length(protein_cols), "proteins...\n")

all_wilcox_results <- data.frame()

pb <- txtProgressBar(min = 0, max = length(protein_cols), style = 3)

for(i in 1:length(protein_cols)) {
  protein_results <- wilcox_test_protein(protein_cols[i], protein_long)
  
  if(!is.null(protein_results) && nrow(protein_results) > 0) {
    all_wilcox_results <- rbind(all_wilcox_results, protein_results)
  }
  
  setTxtProgressBar(pb, i)
}

close(pb)

cat("\n\nTests completed!\n")
cat("Total tests performed:", nrow(all_wilcox_results), "\n")
cat("Unique comparisons:", paste(unique(all_wilcox_results$Comparison), collapse = ", "), "\n\n")

# Check if we have results
if(nrow(all_wilcox_results) == 0) {
  stop("No Wilcoxon tests were successfully completed. Check your data structure.")
}

# ============================================
# STEP 4: MULTIPLE TESTING CORRECTION
# ============================================

cat("=== Applying Multiple Testing Correction ===\n\n")

# FDR correction within each comparison
all_wilcox_results <- all_wilcox_results %>%
  group_by(Comparison) %>%
  mutate(FDR = p.adjust(P_value, method = "BH")) %>%
  ungroup()

# Global FDR
all_wilcox_results$FDR_global <- p.adjust(all_wilcox_results$P_value, method = "BH")

# Summary by comparison
summary_by_comparison <- all_wilcox_results %>%
  group_by(Comparison) %>%
  summarise(
    N_tests = n(),
    Sig_p005 = sum(P_value < 0.05, na.rm = TRUE),
    Sig_p001 = sum(P_value < 0.01, na.rm = TRUE),
    Sig_FDR005 = sum(FDR < 0.1, na.rm = TRUE),
    Sig_FDR01 = sum(FDR < 0.2, na.rm = TRUE),
    Median_pvalue = median(P_value, na.rm = TRUE),
    Mean_N_pairs = mean(N_pairs, na.rm = TRUE),
    .groups = 'drop'
  )

cat("Summary by comparison:\n")
print(summary_by_comparison)
cat("\n")

# Save results
write.csv(all_wilcox_results, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Wilcoxon_All_Results.csv", row.names = FALSE)
write.csv(summary_by_comparison, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Wilcoxon_Summary.csv", row.names = FALSE)

cat("Files saved:\n")
cat("  ✓ Wilcoxon_All_Results.csv\n")
cat("  ✓ Wilcoxon_Summary.csv\n\n")

# ============================================
# STEP 5: EXTRACT SIGNIFICANT PROTEINS
# ============================================

cat("=== Significant Proteins (FDR < 0.2) ===\n\n")

sig_proteins <- all_wilcox_results %>%
  filter(FDR < 0.2) %>%
  arrange(Comparison, P_value)

if(nrow(sig_proteins) > 0) {
  
  cat("Total significant results:", nrow(sig_proteins), "\n\n")
  
  # By comparison
  sig_by_comp <- sig_proteins %>%
    group_by(Comparison) %>%
    summarise(
      N_sig = n(),
      N_increased = sum(Median_Diff > 0),
      N_decreased = sum(Median_Diff < 0),
      .groups = 'drop'
    )
  
  cat("Breakdown by comparison:\n")
  print(sig_by_comp)
  cat("\n")
  
  # Save
  write.csv(sig_proteins, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Wilcoxon_Significant_FDR05.csv", row.names = FALSE)
  
  # By direction
  sig_increased <- sig_proteins %>% filter(Median_Diff > 0)
  sig_decreased <- sig_proteins %>% filter(Median_Diff < 0)
  
  write.csv(sig_increased, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Wilcoxon_INCREASED.csv", row.names = FALSE)
  write.csv(sig_decreased, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Wilcoxon_DECREASED.csv", row.names = FALSE)
  
  cat("Saved:\n")
  cat("  ✓ Wilcoxon_Significant_FDR05.csv\n")
  cat("  ✓ Wilcoxon_INCREASED.csv (", nrow(sig_increased), "results )\n")
  cat("  ✓ Wilcoxon_DECREASED.csv (", nrow(sig_decreased), "results )\n\n")
  
} else {
  cat("No significant proteins found at FDR < 0.05\n")
  cat("Try relaxing threshold or check if comparisons have enough power\n\n")
}

# ============================================
# VISUALIZATION 1: BAR CHART
# ============================================

cat("=== Creating Visualizations ===\n\n")

# Summary for plotting
plot_summary <- all_wilcox_results %>%
  mutate(
    Significance = case_when(
      FDR < 0.1 ~ "FDR < 0.1",
      FDR < 0.2 ~ "FDR < 0.2",
      P_value < 0.05 ~ "p < 0.05 (nominal)",
      TRUE ~ "Not Significant"
    ),
    Direction = ifelse(Median_Diff > 0, "Increased", "Decreased")
  ) %>%
  filter(FDR < 0.2) %>%
  group_by(Comparison, Direction) %>%
  summarise(Count = n(), .groups = 'drop')

if(nrow(plot_summary) > 0) {
  
  p1 <- ggplot(plot_summary, 
               aes(x = Comparison, y = Count, fill = Direction)) +
    geom_bar(stat = "identity", position = "dodge", 
             color = "black", linewidth = 0.5) +
    geom_text(aes(label = Count), 
              position = position_dodge(width = 0.9), 
              vjust = -0.5, size = 4, fontface = "bold") +
    scale_fill_manual(values = c("Increased" = "#E53935", 
                                  "Decreased" = "#1E88E5")) +
    theme_minimal(base_size = 14) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
      legend.position = "bottom"
    ) +
    labs(
      title = "Significant Proteins by Comparison",
      subtitle = "Wilcoxon signed-rank test (FDR < 0.2)",
      x = "Comparison (Reference vs Timepoint)",
      y = "Number of Significant Proteins",
      fill = "Direction"
    ) +
    ylim(0, max(plot_summary$Count) * 1.15)
  
  print(p1)
  ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Wilcoxon_Barplot_Significant.png", p1, 
         width = 10, height = 7, dpi = 300)
  
  cat("✓ Bar plot saved\n")
}

# ============================================
# VISUALIZATION 2: P-VALUE DISTRIBUTION
# ============================================

p2 <- ggplot(all_wilcox_results, aes(x = P_value)) +
  geom_histogram(bins = 50, fill = "steelblue", 
                 color = "black", alpha = 0.7) +
  geom_vline(xintercept = 0.05, color = "red", 
             linetype = "dashed", linewidth = 1) +
  facet_wrap(~ Comparison, scales = "free_y") +
  theme_minimal(base_size = 12) +
  labs(
    title = "Distribution of P-values",
    subtitle = "Red line = p = 0.05",
    x = "P-value",
    y = "Frequency"
  )

print(p2)
ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Wilcoxon_Pvalue_Distribution.png", p2, 
       width = 12, height = 8, dpi = 300)
cat("✓ P-value distribution saved\n")

# ============================================
# VISUALIZATION 3: VOLCANO PLOTS
# ============================================

if(nrow(sig_proteins) > 0) {
  
  comparisons <- unique(all_wilcox_results$Comparison)
  
  for(comp in comparisons) {
    
    comp_data <- all_wilcox_results %>%
      filter(Comparison == comp) %>%
      mutate(
        Sig_category = case_when(
          FDR < 0.1 ~ "FDR < 0.1",
          FDR < 0.2 ~ "FDR < 0.2",
          P_value < 0.05 ~ "p < 0.05",
          TRUE ~ "Not Significant"
        )
      )
    
    p_vol <- ggplot(comp_data, 
                    aes(x = Median_Diff, y = -log10(P_value), 
                        color = Sig_category)) +
      geom_point(alpha = 0.6, size = 2) +
      geom_hline(yintercept = -log10(0.05), 
                 linetype = "dashed", color = "gray50") +
      geom_vline(xintercept = 0, linetype = "solid", color = "black") +
      scale_color_manual(
        values = c(
          "FDR < 0.1" = "#D32F2F",
          "FDR < 0.5" = "#FF6F00",
          "p < 0.05" = "#FDD835",
          "Not Significant" = "#BDBDBD"
        )
      ) +
      theme_minimal(base_size = 12) +
      labs(
        title = paste("Volcano Plot:", comp),
        x = "Median Difference",
        y = "-log10(p-value)",
        color = "Significance"
      )
    
    ggsave(paste0("Volcano_", comp, ".png"), p_vol, 
           width = 8, height = 7, dpi = 300)
  }
  
  cat("✓ Volcano plots saved\n")
}

# ============================================
# TABLE: TOP PROTEINS
# ============================================

cat("\n=== Top 20 Proteins per Comparison ===\n\n")

comparisons <- unique(all_wilcox_results$Comparison)

for(comp in comparisons) {
  
  top20 <- all_wilcox_results %>%
    filter(Comparison == comp) %>%
    arrange(P_value) %>%
    head(20) %>%
    select(Protein, N_pairs, Median_Ref, Median_Comp, 
           Median_Diff, P_value, FDR)
  
  write.csv(top20, paste0("Top20_", gsub("_vs_", "_", comp), ".csv"), 
            row.names = FALSE)
  
  cat(comp, "- Top 5:\n")
  print(head(top20, 5))
  cat("\n")
}

# ============================================
# FINAL SUMMARY
# ============================================

cat("\n========================================\n")
cat("WILCOXON TEST ANALYSIS COMPLETE\n")
cat("========================================\n\n")

cat("SUMMARY:\n")
cat("  Proteins tested:", length(unique(all_wilcox_results$Protein)), "\n")
cat("  Total tests:", nrow(all_wilcox_results), "\n")
cat("  Significant (FDR < 0.05):", nrow(sig_proteins), "\n\n")

if(nrow(sig_proteins) > 0) {
  cat("BY COMPARISON:\n")
  for(i in 1:nrow(sig_by_comp)) {
    cat("  ", sig_by_comp$Comparison[i], ":\n")
    cat("      Total:", sig_by_comp$N_sig[i], "\n")
    cat("      Increased:", sig_by_comp$N_increased[i], "\n")
    cat("      Decreased:", sig_by_comp$N_decreased[i], "\n")
  }
}

cat("\n========================================\n\n")
```



# ============================================
# Heatmap
# ============================================
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

# ============================================
# STEP 1: CHECK AND CLEAN DATA
# ============================================

cat("=== Checking Data Quality ===\n\n")

# Check for missing values
cat("Missing values in Timepoint:", sum(is.na(dx$Time)), "\n")
cat("Missing values in ID:", sum(is.na(dx$ID)), "\n\n")

# Remove rows with missing Timepoint or ID
aggregated_clean <- dx %>%
  filter(!is.na(Time), !is.na(ID))

cat("Rows after removing NAs:", nrow(aggregated_clean), "\n")
cat("Unique Timepoints:", unique(aggregated_clean$Time), "\n\n")

# ============================================
# STEP 2: DEFINE PROTEINS TO PLOT
# ============================================

# Find protein columns (adjust based on your data structure)
# Method 1: By column position
protein_start <- 12 # Adjust this!
protein_end <- min(408, ncol(aggregated_clean))  # First 400 or however many exist

protein_end <- min(50, ncol(aggregated_clean))  # First 39 or however many exist


protein_cols <- colnames(aggregated_clean)[protein_start:protein_end]

cat("Found", length(protein_cols), "protein columns\n")
cat("First protein:", protein_cols[1], "\n")
cat("Last protein:", protein_cols[length(protein_cols)], "\n\n")

# ============================================
# METHOD 1: HEATMAP - FIXED
# ============================================

cat("=== Creating Heatmap ===\n")

# Calculate means with proper NA handling
protein_means <- aggregated_clean %>%
  select(Time, all_of(protein_cols)) %>%
  group_by(Time) %>%
  summarise(across(all_of(protein_cols), ~mean(.x, na.rm = TRUE)),
            .groups = 'drop')

# Check for any remaining issues
cat("Timepoint column has NAs:", any(is.na(protein_means$Time)), "\n")
cat("Timepoints in means:", protein_means$Time, "\n\n")

# Convert to matrix (FIXED - proper handling)
protein_means_clean <- protein_means %>%
  filter(!is.na(Time)) %>%
  as.data.frame()

# Set rownames manually
rownames(protein_means_clean) <- protein_means_clean$Time

# Remove Timepoint column
protein_means_clean <- protein_means_clean %>%
  select(-Time)

# Transpose
protein_matrix <- t(as.matrix(protein_means_clean))

cat("Matrix dimensions:", dim(protein_matrix), "\n")
cat("Column names:", colnames(protein_matrix), "\n")
cat("Any NAs in matrix:", any(is.na(protein_matrix)), "\n\n")

# Scale by row (Z-score normalization)
protein_scaled <- t(scale(t(protein_matrix)))

# Replace any remaining NaNs with 0
protein_scaled[is.nan(protein_scaled)] <- 0
protein_scaled[is.infinite(protein_scaled)] <- 0

# Create heatmap
library(pheatmap)

tryCatch({
  pheatmap(
    protein_scaled,
    cluster_rows = TRUE,
    cluster_cols = FALSE,
    show_rownames = FALSE,
    show_colnames = TRUE,
    color = colorRampPalette(c("blue", "white", "red"))(100),
    main = paste("Heatmap:", length(protein_cols), "Proteins Across Timepoints"),
    fontsize = 10,
    cellwidth = 30,
    cellheight = 1,
    filename = "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/NEG/Heatmap_NEG_Proteins.pdf",
    width = 8,
    height = 20
  )
  cat("✓ Heatmap saved: Heatmap_Proteins.png\n\n")
}, error = function(e) {
  cat("Error creating heatmap:", e$message, "\n")
  cat("Trying alternative method...\n\n")
})
```


```{r}
# ============================================
# METHOD 2: GGPLOT HEATMAP (More robust)
# ============================================

cat("=== Creating ggplot Heatmap ===\n")

# Reshape to long format
protein_long <- aggregated_clean %>%
  select(ID, Time, all_of(protein_cols)) %>%
  pivot_longer(
    cols = all_of(protein_cols),
    names_to = "Protein",
    values_to = "Expression"
  ) %>%
  filter(!is.na(Expression))

cat("Long format created:", nrow(protein_long), "rows\n")

# Calculate means
heatmap_data <- protein_long %>%
  group_by(Protein, Time) %>%
  summarise(Mean_Expression = mean(Expression, na.rm = TRUE),
            .groups = 'drop') %>%
  group_by(Protein) %>%
  mutate(
    Z_score = if(sd(Mean_Expression, na.rm = TRUE) > 0) {
      (Mean_Expression - mean(Mean_Expression, na.rm = TRUE)) / sd(Mean_Expression, na.rm = TRUE)
    } else {
      0
    }
  ) %>%
  ungroup() %>%
  filter(!is.na(Z_score), !is.infinite(Z_score))

# Create heatmap
p_heatmap <- ggplot(heatmap_data, 
                    aes(x = Time, y = Protein, fill = Z_score)) +
  geom_tile(color = NA) +
  scale_fill_gradient2(
    low = "#2166AC", mid = "white", high = "#B2182B",
    midpoint = 0,
    name = "Z-score",
    limits = c(-3, 3),
    oob = scales::squish
  ) +
  theme_minimal(base_size = 10) +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid = element_blank()
  ) +
  labs(
    title = paste("Expression Heatmap:", length(protein_cols), "Proteins"),
    subtitle = "Each row = one protein, colored by z-scored expression",
    x = "Timepoint",
    y = paste(length(protein_cols), "Proteins")
  )

print(p_heatmap)
ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/NEG/Heatmap_ggplot_Proteins.png", p_heatmap, 
       width = 8, height = 20, dpi = 300)
ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/NEG/Heatmap_ggplot_Proteins.pdf", p_heatmap, 
       width = 5, height = 10, dpi = 300)
cat("✓ ggplot heatmap saved: Heatmap_ggplot_Proteins.png\n\n")

# ============================================
# METHOD 3: FACETED PLOT (SUBSET)
# ============================================

cat("=== Creating Faceted Plot (First 100) ===\n")

# Use only first 100 proteins for faceted view
proteins_subset <- protein_cols[1:min(100, length(protein_cols))]

subset_long <- protein_long %>%
  filter(Protein %in% proteins_subset)

p_faceted <- ggplot(subset_long, 
                    aes(x = Time, y = Expression, group = ID)) +
  geom_line(alpha = 0.2, color = "gray70", linewidth = 0.3) +
  stat_summary(aes(group = 1), fun = mean, geom = "line", 
               color = "red", linewidth = 0.8) +
  stat_summary(aes(group = 1), fun = mean, geom = "point", 
               color = "red", size = 1.5) +
  facet_wrap(~ Protein, scales = "free_y", ncol = 10) +
  theme_minimal(base_size = 6) +
  theme(
    strip.text = element_text(size = 5),
    axis.text = element_text(size = 4),
    panel.spacing = unit(0.1, "lines")
  ) +
  labs(
    title = paste("Trajectories for First", length(proteins_subset), "Proteins"),
    subtitle = "Red = mean trajectory, gray = individual subjects",
    x = "Timepoint",
    y = "Expression"
  )

print(p_faceted)
ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/NEG/Faceted_First100_Proteins.png", p_faceted, 
       width = 20, height = 20, dpi = 300)

cat("✓ Faceted plot saved: Faceted_First100_Proteins.png\n\n")

# ============================================
# METHOD 4: MEAN TRAJECTORIES OVERLAY
# ============================================

cat("=== Creating Overlay Plot (Mean Trajectories) ===\n")

# Calculate mean trajectory for each protein
mean_trajectories <- protein_long %>%
  group_by(Protein, Time) %>%
  summarise(Mean_Expression = mean(Expression, na.rm = TRUE),
            .groups = 'drop') %>%
  group_by(Protein) %>%
  mutate(Mean_Scaled = scale(Mean_Expression)[,1]) %>%
  ungroup()

# Select random subset for visibility
set.seed(123)
proteins_to_show <- sample(unique(mean_trajectories$Protein), 
                          min(50, length(unique(mean_trajectories$Protein))))

mean_subset <- mean_trajectories %>%
  filter(Protein %in% proteins_to_show)

p_overlay <- ggplot(mean_subset, 
                    aes(x = Time, y = Mean_Scaled, 
                        group = Protein, color = Protein)) +
  geom_line(alpha = 0.6, linewidth = 0.8) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none") +
  labs(
    title = "Mean Trajectories: 50 Random Proteins",
    subtitle = "Z-scored for comparison",
    x = "Timepoint",
    y = "Z-scored Expression"
  )

print(p_overlay)
ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/NEG/Overlay_50_Proteins.png", p_overlay, 
       width = 10, height = 7, dpi = 300)

cat("✓ Overlay plot saved: Overlay_50_Proteins.png\n\n")

# ============================================
# METHOD 5: SUMMARY STATISTICS
# ============================================

cat("=== Creating Summary Statistics ===\n")

summary_by_timepoint <- protein_long %>%
  group_by(Time, Protein) %>%
  summarise(
    Mean = mean(Expression, na.rm = TRUE),
    SD = sd(Expression, na.rm = TRUE),
    Median = median(Expression, na.rm = TRUE),
    N = sum(!is.na(Expression)),
    .groups = 'drop'
  )

write.csv(summary_by_timepoint, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/NEG/Protein_Summary_by_Timepoint.csv", row.names = FALSE)

cat("✓ Summary statistics saved: Protein_Summary_by_Timepoint.csv\n\n")

# Overall summary
overall_summary <- data.frame(
  Total_Proteins = length(protein_cols),
  Total_Observations = nrow(protein_long),
  Timepoints = paste(unique(aggregated_clean$Time), collapse = ", "),
  Subjects = length(unique(aggregated_clean$ID)),
  Missing_Values = sum(is.na(protein_long$Expression)),
  Missing_Percent = round(sum(is.na(protein_long$Expression)) / nrow(protein_long) * 100, 2)
)

cat("=== Overall Summary ===\n")
print(overall_summary)
write.csv(overall_summary, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/NEG/Overall_Summary.csv", row.names = FALSE)

# ============================================
# METHOD 6: CLUSTERED HEATMAP (ALTERNATIVE)
# ============================================
```
# ============================================
# METHOD 6: CLUSTERED HEATMAP (ALTERNATIVE)
#ComplexHeatmap
# ============================================

```{r}
cat("\n=== Creating Clustered Heatmap (Alternative) ===\n")

# Using ComplexHeatmap if available
if(require(ComplexHeatmap, quietly = TRUE)) {
  
  library(ComplexHeatmap)
  library(circlize)
  
  # Create color function
  col_fun <- colorRamp2(c(-2, 0, 2), c("blue", "white", "red"))
  
  # Create heatmap
  ht <- Heatmap(
    protein_scaled,
    name = "Z-score",
    col = col_fun,
    cluster_rows = TRUE,
    cluster_columns = FALSE,
    show_row_names =TRUE,
    show_column_names = TRUE,
    column_title = paste(length(protein_cols), "Proteins Across Timepoints"),
    row_title = "Proteins",
    heatmap_legend_param = list(title = "Z-score")
  )
  
  png("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/NEG/Heatmap_ComplexHeatmap_Proteins_neg.png", 
      width = 5, height = 10, units = "in", res = 300)
    pdf("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/NEG/Heatmap_ComplexHeatmap_Proteins_Neg.pdf", 
      width = 5, height = 10)
  draw(ht)
  dev.off()
  
  cat("✓ ComplexHeatmap saved: Heatmap_ComplexHeatmap_Proteins.png\n")
  
} else {
  cat("ComplexHeatmap not available, skipping\n")
}

# ============================================
# FINAL SUMMARY
# ============================================

cat("\n\n========================================\n")
cat("VISUALIZATION COMPLETE\n")
cat("========================================\n\n")

cat("FILES CREATED:\n")
cat("  ✓ Heatmap_Proteins.png (pheatmap)\n")
cat("  ✓ Heatmap_ggplot_Proteins.png (ggplot)\n")
cat("  ✓ Faceted_First100_Proteins.png (faceted)\n")
cat("  ✓ Overlay_50_Proteins.png (overlay)\n")
cat("  ✓ Protein_Summary_by_Timepoint.csv (data)\n")
cat("  ✓ Overall_Summary.csv (summary)\n\n")

cat("PROTEINS ANALYZED:", length(protein_cols), "\n")
cat("TIMEPOINTS:", paste(unique(aggregated_clean$Time), collapse = ", "), "\n")
cat("SUBJECTS:", length(unique(aggregated_clean$ID)), "\n\n")

cat("========================================\n\n")
```

```{r}
# ============================================
# COMPLETE WORKFLOW FOR 400 PROTEINS
# ============================================

library(dplyr)
library(tidyr)
library(ggplot2)

# 1. Define which proteins to plot
protein_cols <- colnames(dx)[12:408]  # Adjust indices!

cat("Plotting", length(protein_cols), "proteins\n\n")

# 2. HEATMAP (best overview for many proteins)
protein_means_matrix <- dx %>%
  select(Time, all_of(protein_cols)) %>%
  group_by(Time) %>%
  summarise(across(all_of(protein_cols), ~mean(.x, na.rm = TRUE))) %>%
  column_to_rownames("Time") %>%
  t()

protein_scaled <- t(scale(t(protein_means_matrix)))

library(pheatmap)
pheatmap(
  protein_scaled,
  cluster_rows = TRUE,
  cluster_cols = FALSE,
  show_rownames = FALSE,
  main = "400 Proteins Across Timepoints",
  filename = "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Heatmap_400_Proteins_Final.png",
  width = 8,
  height = 20
)

cat("✓ Created heatmap\n")

# 3. FACETED PLOT (for detailed view)
protein_long <- dx%>%
  select(ID, Time, all_of(protein_cols)) %>%
  pivot_longer(cols = all_of(protein_cols),
               names_to = "Protein",
               values_to = "Expression")

p_faceted <- ggplot(protein_long, 
                    aes(x = Time, y = Expression, group = ID)) +
  geom_line(alpha = 0.1, color = "gray70", linewidth = 0.2) +
  stat_summary(aes(group = 1), fun = mean, geom = "line", 
               color = "red", linewidth = 0.6) +
  facet_wrap(~ Protein, scales = "free_y", ncol = 20) +
  theme_minimal(base_size = 4) +
  theme(
    strip.text = element_text(size = 3),
    axis.text = element_text(size = 3)
  ) +
  labs(title = "400 Protein Trajectories", x = "Time", y = "Expression")

ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/Faceted_400_Proteins_Final.png", p_faceted, 
       width = 40, height = 40, dpi = 300, limitsize = FALSE)

cat("✓ Created faceted plot\n")

# 4. SUMMARY STATISTICS
summary_stats <- protein_long %>%
  group_by(Protein, Time) %>%
  summarise(
    Mean = mean(Expression, na.rm = TRUE),
    SD = sd(Expression, na.rm = TRUE),
    .groups = 'drop'
  )

write.csv(summary_stats, "/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/400_Proteins_Summary_Stats.csv", row.names = FALSE)

cat("✓ Saved summary statistics\n")
cat("\nComplete!\n")
```

#compare witrh 395 altered proteins in meningful changes
```{r}
library(readxl)

dmean<- read.csv("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/result_StreamlitApp/Meaningful_Change1_3/MeanigfulDiff_presistence_burnout/dataExc1C3MeaningfulAcross.csv")

##C1 and C6 values
dlmerC1C6<- read_excel("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/C1C6/pathway_annotation.xlsx", sheet = "list")
meanLmer <- dmean%>%
  select(any_of(dlmerC1C6$preferredName)) #Q13434, Q5H989
```

##C1to C6 values burnout
```{r}
library(readxl)
##C1ToT6
dposLmer<- read_excel ("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/ORA_PosNeg.xlsx", sheet = "Sheet1")

dnegLmer<- read_excel ("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/ORA_PosNeg.xlsx", sheet = "Sheet2")
## merge with lmer positive and negative 

meanLmer <- dmean%>%
  select(any_of(dposLmer$UNIPROT)) #Q13434, Q5H989

meanPLmer <- dmean%>%
  select(any_of(dnegLmer$UNIPROT))  #NA

```

#C1C6valuea burnout
```{r}
dlmer<- read_excel("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/C1C6/pathway_annotation.xlsx", sheet = "list")

meanLmer <- dlmer%>%
  select(any_of(dlmer$preferredName))  #NA

colnames(meanLmer)
```

#Box Plot of selected proteins
ATF4, BDH2, FABP9, and DNAJC21

```{r}
full_data<- read.csv("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/fullData.csv")


rownames(full_dataT)<-full_dataT$X

dex<- read.csv("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/MetaData/burnoutEx.csv")
rownames(dex)<-dex$X

protein_cols <- colnames(dex)[12:5442]

colnames(dex)[5]<- "Time"

dx_clean <- dex %>%
  group_by(ID, Time) %>%
  summarise(across(all_of(protein_cols), ~mean(.x, na.rm = TRUE)), 
            .groups = 'drop')

proteins_subset<- intersect(colnames(dx_clean), c("P18848", "Q9BUT1", "Q0Z7S8", "Q5F1R6"))



# Reshape to long format
protein_lSub<- dx_clean %>%
  select(ID, Time, all_of(proteins_subset)) %>%
  pivot_longer(
    cols = all_of(proteins_subset),
    names_to = "Protein",
    values_to = "Expression"
  )
```

```{r}

# Plot longitudinal trajectories
p_faceted <- ggplot(
  protein_lSub,
  aes(x = Time, y = Expression, group = ID)
) +
  # Individual trajectories
  geom_line(
    alpha = 0.2,
    color = "gray70",
    linewidth = 0.5
  ) +
  # Mean trajectory
  stat_summary(
    aes(group = 1),
    fun = mean,
    geom = "line",
    color = "red",
    linewidth = 0.8
  ) +
  stat_summary(
    aes(group = 1),
    fun = mean,
    geom = "point",
    color = "red",
    size = 1.5
  ) +
  # Facet by protein
  facet_wrap(
    ~ Protein,
    scales = "free_y",
    ncol = 2
  ) +
  # Styling
  theme_minimal(base_size = 10) +
  theme(
    strip.text   = element_text(size = 10),
    axis.text   = element_text(size = 10),
    panel.spacing = unit(0.1, "lines")
  ) + theme_bw()+
  labs(
    title = paste(
      "Longitudinal Trajectories for",
      length(proteins_subset),
      "Selected Proteins"
    ),
    subtitle = "Gray: individual subjects · Red: mean trajectory",
    x = "Timepoint",
    y = "Expression"
  )

print(p_faceted)

```

```{r}
ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/C1C6/MeaningfulC1C6_Proteins.png", p_faceted, 
       width = 5, height = 7, dpi = 300)

ggsave("/Users/nasimbararpour/Library/CloudStorage/Box-Box/NasimBararpour/Lasting_Change_Study/Olink/Datasets/res_lmerlongitudinal/C1C6/MeaningfulC1C6_Proteins.pdf", p_faceted, 
       width = 5, height = 8, dpi = 300)

cat("✓ Faceted plot saved: Faceted_First100_Proteins.png\n\n")
```

